### 2011-11-02 paper.tex
moved existing files from pubtex.

%\documentclass[a4paper,10pt]{article}
\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{../bib/IEEEtran}
%\documentclass[correspondence,9pt]{../bib/IEEEtran}

\usepackage{../bib/mathbf-abbrevs}

\input{defs}

%opening
\title{Coding with the Lattice $A_n$}
\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo}

\begin{document}

\newcommand{\calR}{\mathcal{R}}
\newcommand{\hist}{\operatorname{hist}}

\maketitle

\begin{abstract}
Lattices can be used to construct block codes.  The special family of root lattices has attracted particular attention for this purpose.  We consider the root lattice $A_n$ and derive formulae for the moments of its Voronoi cell.  These formulae enable accurate prediction of the performance of codes constructed from $A_n$.  In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.  These are the densest known sphere packings in dimensions two and three.  Our results automatically include codes constructed using these packings.
\end{abstract}

\begin{keywords}
Lattices, lattice decoding, root lattice, probability of error.
\end{keywords}

\section{Introduction}\label{sec:introduction}

We consider block codes for the Gaussian channel constructed from a subset of points from a Euclidean lattice using \emph{lattice decoding} (as opposed to maximum likelihood decoding)~\cite{Erex2004_lattice_decoding}.  It is known that such codes can reach the capacity of the Gaussian channel~\cite{Buda1989_some_opt_codes_structure,Erez2005,Erex2004_lattice_decoding}.  These capacity results make use of randomly generated lattices that do not have any obvious exploitable structure for efficient decoding and encoding~\cite{Jalden2005_sphere_decoding_complexity}.  For this reason (and many others) the study of specific lattices with known structure is important.  The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention for this purpose~\cite{SPLAG}.  In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.  This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.  Conway and Sloane suggested this type of procedure would be possible in~\cite{Conway1982VoronoiRegions}.  This paper can be seen as a continuation of that work focusing particularly on the lattice $A_n$.

In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.  These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.  Unfortunately, but not surprisingly, the lattice $A_n$ does not produce good codes in large dimensions.  We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.  In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e. \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result without any proof.  We give recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.  
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $n$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In  Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.  We also plot the results of Monte-Carlo simulations.  These simulations corroborate with our analytical results.  %An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

\section{Lattices, lattice codes, and lattice decoding} \label{sec:latt-latt-codes}

A \term{lattice}\index{lattice}, $\Lambda$, is a set of points in $\reals^m$ such that
\[ 
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf$ is an $m \times n$ matrix of rank $n$ called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.  If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. The set of integers $\ints^n$ is a lattice (a generator is the identity matrix) and we call this the~\emph{integer lattice}.

The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^n$ is the subset of $\reals^n$ containing all points nearer (in Euclidean distance) the lattice point at the origin than any other lattice point. The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.  It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\reals^n$.  That is, the Voronoi cell tessellates space when translated by points in $\Lambda$.  To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed. Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.  We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.

The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, center density, thickness, and the normalized second moment (or quantising constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}. The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.  There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.  In general these algorithms are only computationally feasible when the dimension is small (approximately $n \leq 9$).  Even with a complete description of the Voronoi cell it is not necessarily easy to compute the probability of coding error.

The Voronoi cell is linked with the problem of \emph{lattice decoding}.  Given some point $\ybf \in \reals^n$ a \emph{lattice decoder} (or \emph{nearest lattice point algorithm}) returns the lattice point in $\Lambda$ that is nearest to $\ybf$~\cite{Agrell2002}.  Equivalently it returns the lattice point $\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$ contains $\ybf$.  Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small (approximately $n \leq 60$). One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.   %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder. A good overview of these techniques is given by~\cite{Agrell2002}.
Fast nearest point algorithms are known for specific lattices~\citep{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}. For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e. in a number of operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.

%Approximate algorithms for computing the nearest point have also been studied.  A classic example is \term{Babai's nearest plane algorithm}~\citep{Babai1986}\index{Babai's nearest plane algorithm}, which requires $O(n^4)$ arithmetic operations in the worst case where $n$ is the dimension of the lattice and only $O(n^2)$ if the lattice basis is \term{Lov\'asz reduced}~\citep{Lenstra1982}\index{Lov\'asz reduced}. Recently, various approximate techniques for solving the nearest lattice point problem have been motivated by applications to MIMO communications.  An example is the \term{$K$-best algorithm}\index{K-best algorithm}~\citep{Zhan2006_K_best_sphere_decoder} that works similarly to the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} used in coding theory. Yet another example is the \term{fixed sphere decoder}~\citep{Jalden2009_error_prob_fixed_sphere_decoder,Barbero2008_fixed_sphere_decoder}. In this thesis we will make use of Babai's nearest plane algorithm, the sphere decoder and the $K$-best algorithm.  We will not detail the workings of these algorithms as excellent descriptions already exist in the literature cited above.

%For each lattice point $\ybf \in \Lambda$ there exists a Voronoi cell containing those point from $\reals^n$ nearer to $\ybf$ that any other lattice point.  Due to the structure of the lattice the Voronoi cell of each lattice point is the same shape.  Denote by $\vor(\Lambda)$ the Vornoi cell of the lattice point at the origin. NEAREST POINT PROBLEM.
\newcommand{\calX}{\mathcal X}
Lattices can be used to construct \emph{block codes}.  A block code $\calX$ of length $n$ is a finite set of vectors from $\reals^n$.  Each vector in $\calX$ is called a \emph{codeword}.  The number of codewords is denoted by $\abs{\calX}$.  %If each codeword is transmitted with equal probability then the rate of the code is $R = B\frac{1}{n}\log_2\abs{\calX}$ where $B$ is the symbol rate or bandwidth, i.e. the rate in Hertz at which the elements (or symbols) in a codeword are transmitted.  
The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the additive white Gaussian channel the received signal takes the form
\[
\ybf = \xbf_0 + \wbf
\]
where $\ybf \in \reals^n$, $\xbf_0 \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.  If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
\begin{equation}\label{eq:mldecoder}
\hat{\xbf} = \min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
\end{equation}
That is, the receiver computes the codeword in $\calX$ nearest in Euclidean distance to the received signal $\ybf$.  Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
\[
\sum_{\xbf \in \calX} \frac{\prob( \hat{\xbf} = \xbf )}{\abs{\calX}}.
\]
The probability of error is $1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf )$.  %The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.  If the codewords are chosen uniformly randomly on a sphere of radius $\sqrt{P}$ then as $n$ gets large the rate of the code can be made arbitrarily close to the channel capacity while simultaneously the probability of error can be made arbitrarily small.  Another way of stating this is that we can choose close to  $2^{nR/B}$ codewords uniformly randomly on a sphere of radius $\sqrt{P}$ and as $n$ gets large the probability of error decreases to zero.

%A problem with randomly generated codes and maximum likelihood decoding is that there are no computationally efficient methods for encoding and decoding.  For encoding the transmitter must store a list of size $\abs{\calX} \approx 2^{nR/B}$ and for decoding the receiver must compute $\hat{\xbf}$ requiring a search over all $\abs{\calX} \approx 2^{nR/B}$ codewords.  As $n$ gets large this becomes infeasible.  To avoid this computational complexity it is preferable for a code to have some type of `structure' that is exploitable for efficient encoding and decoding.  One way of enforcing structure is to choose codewords from a lattice.  The idea being that encoding and decoding can now be performed efficiently by \emph{lattice decoding}, i.e. by finding a nearest point in a lattice.

A \emph{lattice code} is a block code with codewords given by a finite subset of points from some lattice $\Lambda$ in $\reals^n$.  There are an infinity of ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$ called a \emph{shaping region}.  The lattice points in $\Lambda$ that intersect the shaping region make up the code.  That is, if $S \subset \reals^n$ is a shaping region then our codewords are given by
\[
\calX = S \cap \Lambda.
\]
Possible choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}. A consequence of the finiteness of a lattice code is that the maximum likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to lattice decoding.  Computing a nearest lattice point will not in general return a lattice point from the code (the decoded lattice point might lie outside the shaping region).  A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.  This question was answered in the affirmative by Erez and Zamir~\cite{Erex2004_lattice_decoding}.

Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.  Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).  A result of this is that that observed noise distribution is no longer Gaussian.  However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.  In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
\begin{align}
P_C &= \frac{1}{\abs{\calX}} \sum_{\cbf\in \calX} \prob( \xbf + \wbf \in \vor(\Lambda) + \xbf )  \\
&= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)} e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
\end{align}
The probability of error is $P_E = 1 - P_C$.  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ we obtain
\begin{align}
P_C &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{ 4\sigma^42!} - \dots d\xbf \nonumber \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty \frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf.  \label{eq:summomentproberror}
\end{align}
So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf$ for $m=1,2\dots,M$ for some sufficiently large $M$.

In this paper we focus on lattice codes constructed from the family of lattices called $A_n$. We will discover expressions for
\[
 C_n(m) = \int_{\vor(A_n)} \left(\|\xbf\|^2\right)^m d\xbf.
\]
These can be summed to give arbitrarily accurate approximations for the probability of error.  In this paper we call $C_n(m)$ the $m$th \emph{moment} of the lattice $A_n$.

% We now briefly describe our setting for coding using lattices.  Let $C$ denote an arbitrary set of codewords, each codeword a vector from $\reals^n$.  The rate of this code is $\log_2(\abs{C})$ bits per dimension (or symbol) where $\abs{C}$ is the number of codewords in $C$.  To each codeword $\ybf \in C$ there exists a Voronoi cell, or nearest neighbour region, containing all those points from $\reals^n$ nearer in Euclidean distance to $\ybf$ than any other codeword in $C$.  Denote by ${\mathcal D}(\ybf)$ the Vornoi cell of $\ybf$.  The average probability of correct decoding in the Gaussian channel is
% \[
% P_E(C) = \frac{1}{(\sqrt{2\pi}\sigma)^n \abs{C}} \sum_{\ybf \in C} \int_{{\mathcal D}(\ybf)} e^{\|\xbf - \ybf\|^2 / 2\sigma^2 } d\xbf.
% \] 
% When $n$ is large it is known that randomly generated codes approach channel capacity.  However, random codes are not practical when $n$ is large as decoding and encoding operations require enumeration over each codeword in $C$ and for a fixed rate the number of codewords grow exponentially in $n$.  For this reason, code with structure are prefered.

% One way to prescribe structure is to take codewords from a \emph{lattice}.

% A code is constructed from a lattice $\Lambda$ by taking a finite subset of lattice points.  Typically this is achieved using a region of finite volume $R \subset \reals^n$, called the \emph{shaping region}.  The codewords are all those lattice points that intersect with $R$, i.e. our code is $C = \Lambda \cap R$.  An uncountable number of shaping regions are possible  but spheres, sphereical shells and hypercubes are common.  Another possiblity is to choose as $R$, the Voronoi cell $\vor{\Lambda'}$ where $\Lambda'$ is a sublattice of $\Lambda$.  A specific advantage of this approach is that encoding and decoding can be efficiently performed if there is a fast algorithm for computing a nearest point in $\Lambda$~\cite{Conway1982FastQuantDec}.

% A caveat with this construction is that the Voronoi cells of the codewords in $C$ are no longer nessecarily similar to the Voronoi cell of $\Lambda$.  This is a results of taking only a finite set of lattice points.  Infact some of the codewords will have unbounded Voronoi cells.  The problem is that now

% Given a lattice code $C = \Lambda \cap R$ and a recieved codeword $\ybf = \xbf + \wbf$ where $\xbf \in C$ and $\wbf$ is a vector of independent and identically distributioned Gaussian random variables with variance $\sigma^2$, the maximum likelihood decoder returns the lattice point in $C$ that is nearest to $\ybf$ in Euclidean distance.  A difficult

\section{The main result}\label{sec:main-result}

We now state our main result.  The $m$th moment $C_n(m)$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{C_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)} 
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $C_m(n)$ in terms of $n$.  The first four such formula are:
\begin{align*}
C_0(n) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
C_1(n) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
C_2(n) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
C_3(n) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
%C_4(n) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have tabulated these formula for $m=0$ to $40$.  We derive these results in Section~\ref{sec:powers-eucl-norm}.  First we need some properties of the lattice $A_n$.

\section{The lattice $A_n$}\label{sec:lattice-a_n}
Let $H$ be the hyperplane orthogonal to the all ones vector of length $n+1$, denoted by $\onebf$, that is
\[
\onebf = \left[ \begin{array}{cccc} 1 & 1 & \cdots & 1 \end{array} \right]^\prime,
\]
where superscript $^\prime$ indicates the transpose.  Any vector in $H$ has the property that the sum (and therefore the mean) of its elements is zero and for this reason $H$ is often referred to as the \emph{zero-sum plane} or the \emph{zero-mean plane}.  
The lattice $A_n$ is the intersection of the integer lattice $\ints^{n+1}$ with $H$, that is
\begin{equation} 
\label{eq:An_sub_Zn}
  A_{n} = \ints^{n+1} \cap H = \big\{ \xbf \in \ints^{n+1} \mid \xbf\cdot\onebf = 0  \big\}.
\end{equation}
Equivalently, $A_n$ consists of all of those points in $\ints^{n+1}$ with coordinate sum equal to zero.
% A generator matrix for $A_n$ is the $(n+1)\times n$ matrix
% \[
% \left[ \begin{array}{rrrrrrr}
% 1 & 0 & 0 &  \cdots & 0 & 0 \\
% -1 & 1 & 0 &  \cdots & 0 & 0 \\
% 0 & -1 & 1 &  \cdots & 0 & 0 \\
% 0 & 0 & -1 & \cdots & 0 & 0 \\
% \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
% 0 & 0 & 0 & \cdots & -1 & 1 \\
% 0 & 0 & 0 & \cdots & 0 & -1  \\
% \end{array} \right].
% \] 
The lattice has determinant $n+1$ so the volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$.  The lattice has $n(n+1)$ minimal vectors, each of norm $2$, so the packing radius is $\frac{1}{\sqrt{2}}$~\cite[p. 108]{SPLAG}.

The Voronoi cell of $A_n$ is closely related to the $n+1$ dimensional hypercube $\vor(\ints^{n+1})$ as the next theorem will show.  This result has appeared previously~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}, but we repeat it here so that this paper is self contained.  We denote by 
\[
\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
\]
the projection matrix orthogonal to $\onebf$ (i.e. into $H$) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.  Given a set $S$ of vectors from $\reals^{n+1}$ we write $\Qbf S$  to denote the set with elements $\Qbf \sbf$ for all $\sbf \in S$, i.e. the set containing the projection of the vectors from $S$.

\begin{lemma} \label{lem:QVorZnsubsetVorAn}
The projection of $\vor(\ints^{n+1})$ into H is a subset of $\vor(A_{n}) \cap H$.  That is,
\[
\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_{n}).
\]
\end{lemma}
\begin{IEEEproof}
Let $\ybf \in \vor(\ints^{n+1})$.  Decompose $\ybf$ into orthogonal components so that $\ybf = \Qbf \ybf + t \onebf$ for some $t \in \reals$.  Then $\Qbf\ybf \in \Qbf\vor(\ints^{n+1})$.  Assume that $\Qbf\ybf \notin \vor(A_n)$.  Then there exists some $\xbf \in A_n$ such that
\begin{align*}
\|\xbf - \Qbf\ybf\|^2 < \|\zerobf - \Qbf\ybf\|^2 & \Rightarrow \|\xbf - \ybf + t\onebf\|^2 < \|\ybf - t\onebf\|^2 \\
%& \Rightarrow \|\xbf - \ybf\|^2 + 2t(\xbf - \ybf)'\onebf + t^2\|\onebf\|^2 < \|\ybf\|^2 - 2t\ybf'\onebf + t^2\|\onebf\|^2 \\
& \Rightarrow \|\xbf - \ybf\|^2 + 2t\xbf'\onebf < \|\ybf\|^2.
\end{align*}
By definition \eqref{eq:An_sub_Zn} $\xbf'\onebf = 0$ so $\|\xbf - \ybf\|^2 < \|\ybf\|^2$.  This violates that $\ybf \in \vor(\ints^{n+1})$ and hence $\Qbf\ybf \in \vor(A_n)$.\footnote{This proof can be generalised to show that for any lattice $L$ and hyperplane $P$ such that $P\cap L$ is also a lattice it is true that $p\vor(L) \subseteq \vor(L \cap P) \cap P$ where $p$ indicates the orthogonal projection into $P$~\cite[Lemma~2.1]{McKilliam2010thesis}.}
\end{IEEEproof}

\begin{theorem}  \label{thm:VorAn=QVorZn1}
The projection of $\vor(\ints^{n+1})$ into H is equal to $H \cap \vor(A_{n})$. That is,
\[
H \cap \vor(A_n) = \Qbf\vor(\ints^{n+1}).
\]
\end{theorem}
\begin{proof}
The $n$-volume of $H \cap \vor(A_n)$ is given by the square root of the determinant of $A_n$, that is 
\[
\sqrt{\det{A_n}} = \sqrt{n + 1}.
\]  
From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to $\sqrt{n + 1}$ also. It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq H \cap \vor(A_n)$, so, because the volumes are the same, and because $H \cap \vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $H \cap \vor(A_n) = \Qbf\vor(\ints^{n+1})$.
\end{proof}

\section{Integrating a function over $\vor(A_n)$}\label{sec:integr-funct-over}

Consider a function $f : \reals^{n+1} \mapsto \reals$.  We would like to be able to solve integrals of the form
\begin{equation}\label{eq:vorfint}
\int_{\vor(A_n)} f(\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1})} f(\xbf) d\xbf.
\end{equation}
It is not immediately clear how an integral over $\vor(A_n)$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
\[
\int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
\]
This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.  To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.  Then $\Qbf\xbf = \Qbf(\xbf + k \onebf)$ for all $k \in \reals$ such that $\xbf +  k \onebf \in \vor(\ints^{n+1})$, i.e. for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.  The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by 
\[
\|\onebf\|(1 - x_{\text{max}} + x_{\text{min}}) = \sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}}).
\] 
The integral~\eqref{eq:vorfint} can now be written as
\[
\int_{\vor(A_n)} f(\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
\]
The primary advantage of this integral is that the bounds of are given by the hypercube $\vor(\ints^{n+1})$.

Let us now restrict $f(\xbf)$ so that it depends only on the magnitude $\|\xbf\|$, for example $f(\xbf) = \|\xbf\|^{2m}$ could be a power of the Euclidean norm of $\xbf$.  Now $f(\xbf)$ is invariant to permutation of $\xbf$.  Let $\xbf$ be such that $x_1$ is the maximum element and $x_2$ is the minimum element.  Our integral is now equal to
\[
\frac{n (n+1)}{\sqrt{n+1}} \int^{1/2}_{-1/2} \int^{x_1}_{-1/2}  \int^{x_1}_{x_2} \cdots \int^{x_1}_{x_2} \frac{f(\Qbf\xbf)}{1 - x_{1} + x_{2}}  dx_{n+1} \, \dots \, dx_2 \, dx_1.
\]
The factor $n(n+1)$ arises because there are $n(n+1)$ ways to place two elements (i.e. $x_1$ and $x_2$) into $n+1$ positions.

We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables gives
\[
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} \frac{f(\Qbf\xbf )}{1 - t}  dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
\]
and letting $w_{i-2} = x_i - y + 1/2 + t$ for $i = 3,\dots,n+1$ gives
\begin{equation}\label{eq:intwewilluse}
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{t}_{0} \cdots \int^{t}_{0} \frac{f(\Qbf\xbf )}{1 - t}  dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
\end{equation}
Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector $[t, 0, w_1, w_2, \dots, w_{n-1}]^\prime$.  Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.  Interestingly $\wbf$ does not contain $y$ so the term inside the integral does not depend on $y$.  This is the integral we will use to compute the moments of $A_n$.

\begin{example}\emph{\textbf{(The volume of the Voronoi cell)}
In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting $f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1$ we obtain,
\begin{align*}
C_0 &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
 &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy. \\
&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1}.
\end{align*}
as required.  Here $\beta(x,a,b) =  \int_{0}^{x}  t^{a-1}(1 - t)^{b-1} \,dt$ is the incomplete beta function~\cite{Pearson_tables_of_beta_functions} and we have used the identity $\int^{1}_{0} \beta(y, n, 0) \, dy = \frac{1}{n}$.}
\end{example}

% \subsection{The second moment of the Voronoi cell}

% Setting 
% \begin{align*}
% f(\Qbf\wbf) = \|\Qbf\wbf\|^2  &= \|\wbf\|^2 - \frac{(\wbf^\prime \onebf)^2}{n+1} \\
% &= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t + \sum_{i=1}^{n+1}w_i \right)^2 \\
% &= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n+1}w_i + \left(\sum_{i=1}^{n+1}w_i \right)^2 \right)\\
% &= c + d 
% \end{align*}
% where
% \[
% c = \frac{n}{n+1} t^2,\qquad \text{and} \qquad d = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
% \]
% where $A = \sum_{i=1}^{n-1}w_i^2$ and $B = \sum_{i=1}^{n-1}w_i$.  Now the second moment is given by
% \[
% C_1 = n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{c + d}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy.
% \]
% Integrating $A$, $B$ and $B^2$ over the $w_1,\dots,w_{n-1}$ gives,
% \begin{align*}
% &\int^{t}_{0} \cdots \int^{t}_{0} A \,dw_{n-1}\dots dw_1 = \frac{n-1}{3}t^{n+1},\\
% &\int^{t}_{0} \cdots \int^{t}_{0} B \,dw_{n-1}\dots dw_1 = \frac{n-1}{2}t^n, \,\,\,\, \text{and} \\
% &\int^{t}_{0} \cdots \int^{t}_{0} B^2 \,dw_{n-1}\dots dw_1 = \frac{(n-1)(3n - 2)}{12} t^{n+1}
% \end{align*}
% So integrating $d$ over the $w_1,\dots,w_{n-1}$ gives,
% \[
% \int^{t}_{0} \cdots \int^{t}_{0} d \,dw_{n-1}\dots dw_1 =  \frac{7(n-1)(n-2)}{12(n+1)}t^{n+1}.
% \]
% Also,
% \[
% \int^{t}_{0} \cdots \int^{t}_{0} c \,dw_{n-1}\dots dw_1 =  \frac{n}{n+1}t^{n+1}.
% \]
% So,
% \begin{align*}
% C_1 &=  \sqrt{n+1}\frac{n(12n + 7(n-1)(n-2))}{12(n+1)} \int^{1}_{0} \int^{y}_{0}\frac{t^{n+1}}{1 - t}  \, dt  \, dy \\
% &= \sqrt{n+1}\frac{12n + 7(n-1)(n-2)}{12(n+1)}
% \end{align*}

%\newcommand{\calV}{\mathcal{V}}
\section{The moments of $A_n$}\label{sec:powers-eucl-norm}

We now derive expressions for the $C_n(m)$.  Setting $f(\Qbf\xbf) = \left(\|\Qbf\wbf\|^{2}\right)^m$ in~\eqref{eq:intwewilluse} we obtain, 
\[
\frac{C_m(n)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
\]
Now $\|\Qbf\wbf\|^2 = \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2$ and recalling that $\wbf = [t,0,w_1,\dots,w_{n-1}]'$ we can write
\begin{align*}
\|\Qbf\wbf\|^2 &= \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2 \\
&= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t + \sum_{i=1}^{n+1}w_i \right)^2 \\
&= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n+1}w_i + \left(\sum_{i=1}^{n+1}w_i \right)^2 \right)\\
&= c + d, 
\end{align*}
say, where
\[
c = \left(\frac{n}{n+1} \right) t^2,\qquad \text{and} \qquad d = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
\]
and where $A = \sum_{i=1}^{n-1}w_i^2$ and $B = \sum_{i=1}^{n-1}w_i$.  Now,
\[
 \frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(c+d)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
\]
and by expanding the binomial $(c+d)^m$ we get
\[
\frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} c^{m-k} \int_0^t \cdots \int_0^t \, d^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
\]
Expanding $d^k$ as a trinomial gives
\begin{align*}
d^k &= \sum_{k_1+k_2+k_3=k} \frac{k! A^{k_1} B^{2k_3 + k_2}}{k_1! k_2! k_3!} \left(\frac{-1}{n+1}\right)^{k_2+k_3}2^{k_2}t^{k_2}  \\
&=  \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{k!  A^{a} B^{2k - 2a - b}}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}2^{b}t^{b}
\end{align*}
where the second line follows by setting $k_1 = a$, $k_2 = b$ and $k_3 = k - a - b$.  In Appendix~\ref{sec:mult-type-integr} we show that the integral of $A^{a} B^{2k - 2a - b}$ over $w_1,\dots w_{n-1}$ is
\[
\int_0^t \cdots \int_0^t A^{a} B^{2k-2a-b} dw_{n-1} \, \dots \, dw_1 = t^{n-1+2k - b} G(n-1,a,2k - 2a - b).
\]
where $G(n,k,m)$ satisfies the recursion given by~\eqref{eq:theGrecursion}.  So, let $D$ satisfy
\begin{align*}
D &= t^{1 - n - 2k}\int_0^t \cdots \int_0^t d^k dw_{n-1} \, \dots \, dw_1 \\
&= \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{2^{b} k!G(n-1,a,2k - 2a - b)}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}.
\end{align*}
Now $c^{m-k} = \left(\frac{n}{n+1} \right)^{m-k} t^{2(m-k)}$ and
\begin{align*}
\frac{C_m(n)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} D \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy. \\
&= \frac{1}{n+2m} \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} D.
\end{align*}
This expression is equivalent to that from~(\ref{eq:theCmformula}).

\section{Results and simulations}\label{sec:results-simulations}

We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3$ and $A_4$.  We assume that codes are constructed from these lattices using a \emph{Voronoi constellation}~\cite{Conway1983VoronoiCodes} of size $2^n$.  This corresponds to a rate of $1$ bit per dimension (or symbol).  The shaping region of these constellations is $2\vor(A_n)$, i.e. the Voronoi cell of $A_n$ scaled by a factor of $2$.  The average power of these constellations can be approximated by four times the second moment $4C_1(m)$.  This approximation is conservative as Voronoi constellations can have average power smaller than the second moment of the shaping region~\cite{Conway1983VoronoiCodes}.  The approximation becomes more accurate as the code rate and correspondingly the volume of the shaping region grows.  %In proving that lattice codes using lattice decoding can achieve capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a common dither between transmitter and receiver. This is to ensure that the average power is \emph{exactly} the second moment of the shaping region.  For this reason we think that this approximation is valid.  
The SNR is then given by
\[
\text{SNR} = \frac{4C_1(n)}{n \sigma^2} = \frac{n+3}{3\sqrt{n+1}\sigma^2}
\]   
where $\sigma^2$ is the noise variance.  %As we have assumed that 1 bit is transmitted per symbol the SNR at channel capacity is equal to one.

Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $C_n(m)$ and~\eqref{eq:summomentproberror} (solid line).  The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.  At most 196 moments where needed for Figure~\ref{fig:peplots}.  We also display the probability of error computed approximately by Monte-Carlo simulation (dots).  %We have used the fast nearest point algorithm for $A_n$ described in~\cite{McKilliam2009CoxeterLattices} and~\cite[p. 448]{SPLAG}.  
The simulations are iterated until 500 error events occur.

We have also plotted approximations for the probability of error for codes constructed from the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.  The approximation are made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.  The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.  The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.  By the union bound the probability of error satisfies 
\newcommand{\erfc}{\operatorname{erfc}}
\newcommand{\erf}{\operatorname{erf}}
\[
P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
\]
where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.  The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.

For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  The union bound on the probability of error is
\[
P_E \leq 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
\]
The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.  This was computed numerically by Conway and Sloane~\cite[p. 61]{SPLAG}.  The relationship between SNR and $\sigma^2$ is $\text{SNR} = 4 \times 0.065771 \sigma^{-2}$.

\begin{figure*}[tp]
	\centering 
		\includegraphics{plots/peplots-1.mps}
		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$, $A_4$, $E_8$ and the Leech lattice $\Lambda_{24}$.}
		\label{fig:peplots}
\end{figure*}

\section{Conclusion}

Recursive formulae for the moments of the Voronoi cell of the lattice $A_n$ were found.  These enable accurate prediction of the performance of codes constructed from $A_n$ when \emph{lattice decoding} is used.  In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice and our results include codes constructed using these packings as a special case.  The lattice $A_n$ does not produce good codes in large dimension, but we are hopeful that the techniques developed here can be applied to other lattices.  In particular similar techniques might apply to the family Coxeter lattices, which contains the root lattice $E_8$.

\appendix

\subsection{A multinomial type integral over a hypercube}\label{sec:mult-type-integr}

We need to evaluate integrals of the form
\[
F(n,c,d) = \int^{t}_{0} \cdots \int^{t}_{0} \left(\sum_{j=1}^{n} x_j^2\right)^c \left( \sum_{i=1}^{n} x_i \right)^d \,dx_1\cdots dx_{n}.
 \]
We shall find a recursion describing this integral.  Write
\[
F(n,c,d) = \int^t_0 \cdots \int^t_0 \left(x_n^2 + \sum_{j=1}^{n-1} x_j^2\right)^c \left( x_n + \sum_{i=1}^{n-1} x_i \right)^d \,dx_1\cdots dx_n.
\]
Expanding the two binomials gives
\begin{align*}
F(n,c,d) &=  \int^t_0 \cdots \int^t_0 \sum_{c'=0}^{c} \binom{c}{c'} x_n^{2c'} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'}  \sum_{d'=0}^{d}\binom{d}{d'} x_n^{d'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n} \\
&=  \sum_{c'=0}^{c} \sum_{d'=0}^d \int^t_0 \cdots \int^t_0 \binom{c}{c'} \binom{d}{d'} x_n^{2c'+d'} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n}. 
\end{align*}
Integrating the $x_n$ term gives
\[
F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1} \int^t_0 \cdots \int^t_0 \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
\]
Note that
\[
F(n-1,c-c',d-d')  = \int^t_0 \cdots \int^t_0 \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
\]
So $F(n,c,d)$ satisfies the recursion
\[
F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'}}{2c'+d'+1}F(n-1,c-c',d-d')
\]
with the initial conditions
\[
F(1,c,d) = \frac{t^{2c+d+1}}{2c+d+1} \qquad \text{and} \qquad F(n,0,0) = t^n.
\]
The $F(n,c,d)$ can be written as $t^{n+2c+d}G(n,c,d)$ where $G(n,c,d)$ is rational.  To see this write
\begin{align*}
F(n,c,d) &=  \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1}t^{n-1+2(c-c')+d-d'}G(n-1,c-c',d-d') \\
&=  t^{n+2c+d}\sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1} \\
&=  t^{n+2c+d}G(n,c,d).
\end{align*}
Now $G(n,c,d)$ is the rational number satisfying the recursion
\[
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
\]
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]

\subsection{Solving this recursion for fixed $d$ and $c$}\label{sec:solv-this-recurs}
\newcommand{\calG}{\mathcal G}

For fixed $d$ and $c$ this recursion can be solved explicitly.  Write
\[
G(n,c,d) = G(n-1,c,d) + \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
\]
where the sum $\sum_{ (c',d') \neq (0,0)}$ is over all $0 \leq c' \leq c$ and $0 \leq d' \leq d$ except when both $d$ and $c$ are zero.  Denote by $\calG(z,c,d)$ the $z$-transform of $G(n,c,d)$.  Taking the $z$-transform of both sides in the equation above gives
\[
\calG(z,c,d) = \frac{z^{-1}}{1-z^{-1}} \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{\calG(z,c-c',d-d')}{2c'+d'+1}.
\]
So the $z$-transform $\calG(z,c,d)$ satisfies this recursive equation.  The initial condition is $\calG(z,0,0) = \frac{z^{-1}}{1 - z^{-1}}$.  By inverting this $z$-transform and using the resultant expressions in~\eqref{eq:theCmformula} we obtain formulae in $n$ for the moment $C_n(m)$.  This procedure was used to generate the formula described in Section~\ref{sec:main-result}.

% \section{Conclusion}
% \begin{itemize}
% \item We can use these techniques to compute other things about the Voronoi cell, i.e. the average $L_p$ norms.
% \item Not restricted to Gaussian noise.  Any noise pdf that can be described by the moments will work.
% \item Can other lattices be done?
% \end{itemize}

\small
\bibliography{../bib/bib}

% \section{Notable references}
% \begin{itemize}
% \item Conway and Sloane compute the second moment of all the root lattices and there duals~\cite{Conway1982VoronoiRegions}.  This is really \emph{the} paper on this topic.  The approach is to break the Voronoi cells into simplicies and compute the second moment of the simplicies.

% \begin{quotation}
% In the second application, the same Euclidean code
% $y_1,\dots ,y_n$ is used as a code for the Gaussian channel. Now
% the Voronoi regions are the decoding regions: all points $x$
% in the interior of $V(y_i)$ are decoded as $y_i$. If the codewords
% are equally likely and all the Voronoi regions $V(y_i)$ are
% congruent to a polytope P, the probability of correct decoding is proportional to
% \[
% \int_P e^{x^\prime x} dx.
% \]
% The description of the Voronoi regions given in Section III
% thus makes it possible to calculate this probability exactly
% for many Euclidean codes. These results will be described
% elsewhere.
% \end{quotation}

% But it doesn't appear to be published elsewhere.  I'm not really sure how you would go about doing it either, because I think integrating $e^{x^\prime x}$ over an arbitray simplex is hard.

% \item Baldoni \emph{et. al.}~\cite{Baldoni_how_to_integrate_poly_over_simplex_2008} claim to show that in general integrating a polynomial under a simplex is NP-hard.  However, if you fix the degree of the polynomial then it can be done in polynomial time.

% \end{itemize}

\end{document}

### 2011-11-02 wrongnorm.tex
moved existing files from pubtex.

\documentclass[a4paper,10pt]{article}

\usepackage{../bib/mathbf-abbrevs}

\input{defs}

%opening
\title{Formula for the moments of the lattice $A_n$}
\author{Robby McKilliam}

\begin{document}

\newcommand{\calR}{\mathcal{R}}
\newcommand{\calV}{\mathcal{V}}

Let $\calV$ be the Voronoi cell of the lattice $A_n$.  The probability of correct decoding in Gaussian noise is
\begin{align*}
\prob(\sigma^2) &= \prob( \xbf \in \calV ) = \frac{1}{(\sqrt{2\pi}\sigma)^d}\int_\calV e^{-\|\xbf\|^2 / 2\sigma^2} d\xbf \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^d}\int_\calV 1 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{ 4\sigma^42!} - \frac{\left(\|\xbf\|^2\right)^3}{ 8\sigma^63!} + \dots d\xbf \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^d} \sum_{k=0}^\infty \frac{(-1)^k}{2^k\sigma^{2k}k!} \int_\calV \left(\|\xbf\|^2\right)^k
\end{align*}
where $\sigma^2$ is the noise variance.  The probability of error is $1 - \prob(\sigma^2)$.  So, the aim is to find expressions for
\[
 C_k = \int_\calV \left(\|\xbf\|^2\right)^k d\xbf
\]
These can be summed to give arbitrarily accurate expressions for the probability of error.  A problem could be that you need many terms to get reasonable accuracy when the dimension $n$ is large and when $\sigma^2$ is small.  You could use the covering radius of $A_n$ to get a (very conservative) bound on the number of terms required.  I expect the required number of terms to be quite large.  % Here is a bound on the number of terms

% \begin{proposition}
% Let $\epsilon > 0$ and let $d > 0$ be the smallest integer such that
% \[
% \sqrt{n+1}\frac{R^{2d}}{\sigma^{2d}d!} < \epsilon
% \]
% where
% \[
% R = \sqrt{ \frac{a(n+1 - a)}{n+1}  }
% \] 
% is the covering radius of $A_n$ and $a = \floor{(n+1)/2}$.
% Then,
% \[
% \abs{\prob(\sigma^2) -  \sum_{k=0}^d \frac{(-1)^k}{\sigma^{2k}k!} C_k} < \epsilon.
% \]
% That is, using the first $d+1$ an accuracy of $\epsilon$. 
% \end{proposition}

% This bound is \emph{very} conservative.

Unfortunately, I have computed formulae for are the $k$-norms,
\[
V_k = \int_\calV \|\xbf\|_k^k d\xbf = \int_\calV \left(\sum_{i=1}^{n}x_i^k\right) d\xbf.
\]
I computed the wrong thing!  Fortunately, the techniques I used to get the $V_k$ can be reused to get the $C_k$, but I need to work through a fair bit of new algebra and write the corresponding computer program.

In the meantime, if anyone can think of an application for the $V_k$, i.e. the average of the $k$-norm over the Voronoi cell, then that would be great :)

I'm still very interested in why/how the formula for the $V_k$ simplify as much as they do.  I'm hopeful that similar simplifications will occur for the $C_k$.

%\bibliography{../bib/bib}

\end{document}

### 2011-11-02 build.sh

@@ -1,0 +1,8 @@
+cd plots
+mpost -interaction=nonstopmode peplots.mp
+cd ..
+
+pdflatex paper.tex
+bibtex paper
+pdflatex paper.tex
+pdflatex paper.tex

### 2011-11-02 plots/approxmoments.scala

@@ -1,0 +1,52 @@
+/**
+* Brute force an for powers of the Euclidean norm of An
+*/
+import pubsim.lattices.util.PowerOfEuclideanNorm
+import pubsim.lattices.An.AnFastSelect
+
+for( m <- List(1,2,3,4,5) ){
+  val mfile = new java.io.FileWriter("m" + m + "approx")
+  println(m + " moment test")
+  for( n <- 1 to 40 ) {
+    val lattice = new AnFastSelect(n)
+    val latticeprops = new PowerOfEuclideanNorm(lattice,m)
+    latticeprops.uniformlyDistributed(100000)
+    println(n + "\t" + latticeprops.moment() )
+    mfile.write(n + "\t" +  latticeprops.moment().toString.replace('E', 'e') + "\n")
+  }
+  mfile.close
+  println
+}
+
+//~ println("4th moment test")
+//~ for( n <- 1 to 40 ) {
+  //~ val lattice = new AnFastSelect(n)
+  //~ val latticeprops = new MomentApproximator(lattice,4)
+  //~ latticeprops.uniformlyDistributed(100000)
+  //~ //val moment = n*(-10 + 33*n + 22*n*n + 3*n*n*n)/(240.0*(1 + n)*(1 + n)*(1 + n))
+  //~ println(n + "\t" + latticeprops.moment() )
+//~ }
+//~ println
+//~ println("6th moment test")
+//~ for( n <- 1 to 40 ) {
+  //~ val lattice = new AnFastSelect(n)
+  //~ val latticeprops = new MomentApproximator(lattice,6)
+  //~ latticeprops.uniformlyDistributed(100000)
+  //~ println(n + "\t" + latticeprops.moment() )
+//~ }
+//~ println
+//~ println("8th moment test")
+//~ for( n <- 1 to 40 ) {
+  //~ val lattice = new AnFastSelect(n)
+  //~ val latticeprops = new MomentApproximator(lattice,8)
+  //~ latticeprops.uniformlyDistributed(100000)
+  //~ println(n + "\t" + latticeprops.moment() )
+//~ }
+//~ println
+//~ println("10th moment test")
+//~ for( n <- 1 to 40 ) {
+  //~ val lattice = new AnFastSelect(n)
+  //~ val latticeprops = new MomentApproximator(lattice,10)
+  //~ latticeprops.uniformlyDistributed(100000)
+  //~ println(n + "\t" + latticeprops.moment() )
+//~ }

### 2011-11-02 plots/approxpe.scala

@@ -1,0 +1,19 @@
+/**
+* Brute force an approximation for probabilty of correct decoding with the lattice An
+*/
+import pubsim.lattices.util.ProbabilityOfCodingError
+import pubsim.lattices.An.AnFastSelect
+import pubsim.distributions.GaussianNoise
+
+val n = 8
+val lattice = new AnFastSelect(n)
+val mfile = new java.io.FileWriter("n" + n + "approxpc")
+
+println("var \t prob correct" )
+
+for( v <- (0 to 60).map(1/scala.math.pow(15.0/14.0,_)) ){
+    val probc = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),1000).probError;
+    println(v + "\t" + probc )
+    mfile.write(v + "\t" +  probc.toString.replace('E', 'e') + "\n")
+ }
+  mfile.close

### 2011-11-02 plots/e8sim.scala

@@ -1,0 +1,100 @@
+/**
+* Probability of error plots for the lattice Zn.
+*/
+import java.math.BigInteger
+import pubsim.BigRational
+import pubsim.lattices.E8
+import pubsim.Util
+import pubsim.lattices.util.ProbabilityOfCodingError
+import pubsim.distributions.GaussianNoise
+
+/*
+{
+println
+println("Running Monte-Carlo trials with the E8 lattice")
+
+val lattice = new E8
+val startdb = 5.5
+val stepdb = 0.5
+val tope = 1e-6
+
+val sm = lattice.secondMoment
+
+var pe = 1.0
+var dbnow = startdb 
+val efile = new java.io.FileWriter("pemontecarlo_E8")
+
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val snr = scala.math.pow(10.0,dbnow/10.0)
+  val v = 4.0*sm/snr/8
+  pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),200).probError
+  println(dbnow + "\t" + pe)
+  efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+  dbnow = dbnow + stepdb
+}
+efile.close
+}
+*/
+
+
+{
+val varstart = 1.0/6.0
+val varstep = 9.5/10.0
+val tope = 1e-12
+
+var pe = 1.0
+var varnow = varstart 
+val mfile = new java.io.FileWriter("E8UnionBound")
+
+val tol = new BigRational(BigInteger.ONE, BigInteger.TEN.pow(25));
+
+println
+println("Caculating union lower bound on the error probabilty of E8")
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val r2sdev = new BigRational(scala.math.sqrt(varnow*4.0) ,20)
+  val snr = 929.0/3240.0/varnow
+  val snrdb = 10*scala.math.log10(snr)
+  pe = 240*(1 - pubsim.Util.erf(r2sdev.reciprocal, tol))
+  varnow = varnow*varstep
+  println(snrdb + "\t" + pe)
[…]

### 2011-11-02 plots/leechsim.scala

@@ -1,0 +1,74 @@
+/**
+* Probability of error plots for the lattice Zn.
+*/
+import java.math.BigInteger
+import pubsim.BigRational
+import pubsim.lattices.leech.Leech
+import pubsim.Util
+import pubsim.lattices.util.ProbabilityOfCodingError
+import pubsim.distributions.GaussianNoise
+import pubsim.lattices.GeneralLatticeAndNearestPointAlgorithm
+
+{
+println
+println("Running Monte-Carlo trials with the Eleech lattice")
+
+val lattice = new Leech
+val startdb = 5.5
+val stepdb = 0.5
+val tope = 1e-4
+
+val sm = lattice.secondMoment
+val decoder = new GeneralLatticeAndNearestPointAlgorithm(lattice.getGeneratorMatrix)
+
+var pe = 1.0
+var dbnow = startdb 
+val efile = new java.io.FileWriter("pemontecarlo_Leech_new")
+
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val snr = scala.math.pow(10.0,dbnow/10.0)
+  val v = 4.0*sm/snr/24
+  pe = new ProbabilityOfCodingError(decoder,new GaussianNoise(0,v),100).probError
+  println(dbnow + "\t" + pe)
+  efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+  dbnow = dbnow + stepdb
+}
+efile.close
+}
+
+
+/*
+{
+val varstart = 1.0/2.0
+val varstep = 9.0/10.0
+val tope = 5e-11
+val petol = 1e-16
+
+var pe = 1.0
+var varnow = varstart 
+val mfile = new java.io.FileWriter("moment1")
+
+val tol = new BigRational(BigInteger.ONE, BigInteger.TEN.pow(15));
+
+println
+println("Caculating 'exact' error probabilty using the error function")
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val r2sdev = new BigRational(scala.math.sqrt(varnow*8.0) ,25)
+  val snr = 2.0/3.0/varnow
+  val snrdb = 10*scala.math.log10(snr)
[…]

### 2011-11-02 plots/montecarlo.scala

@@ -1,0 +1,36 @@
+/**
+* Probability of error plots for the lattice An.
+*/
+import pubsim.lattices.util.ProbabilityOfCodingError
+import pubsim.lattices.An.AnFastSelect
+import pubsim.distributions.GaussianNoise
+
+val ns = List(2,3,4)
+
+val startdb = 5.5
+val stepdb = 0.5
+val tope = 1e-6
+
+
+for(n <- ns){
+  val lattice = new AnFastSelect(n)
+  var pe = 1.0
+  var dbnow = startdb 
+  val mfile = new java.io.FileWriter("pemontecarlo" + n)
+
+  println
+  println("Now running n = " + n)
+  println("SNRdB \t Prob Error")
+  while(pe > tope){
+    //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+    val snr = scala.math.pow(10.0,dbnow/10.0)
+    val v = (3.0+n)/3.0/scala.math.sqrt(n+1)/snr
+    pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),500).probError
+    println(dbnow + "\t" + pe)
+    mfile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+    dbnow = dbnow + stepdb
+  }
+
+  mfile.close
+
+}

### 2011-11-02 plots/montecarlo.sh

@@ -1,0 +1,1 @@
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar montecarlo.scala

### 2011-11-02 plots/proberrorsim.scala

@@ -1,0 +1,42 @@
+/**
+* Probability of error plots for the lattice An.
+*/
+import pubsim.lattices.util.PowerOfEuclideanNorm
+import pubsim.lattices.An.AnFastSelect
+import pubsim.lattices.An.MomentComputer
+import pubsim.BigRational
+
+val ns = List(2,3,4)
+//val lattice = new AnFastSelect(n)
+val varstart = new BigRational(1,2)
+val varstep = new BigRational(9,10)
+val tope = 1e-11
+val petol = 1e-16
+
+print("Reading existing moment data, this might take a moment....")
+val mc = new MomentComputer("momentstore")
+println("finished")
+
+for(n <- ns){
+  var pe = 1.0
+  var varnow = varstart 
+  val mfile = new java.io.FileWriter("moment" + n)
+
+  println
+  println("Now running n = " + n)
+  println("SNRdB \t Prob Error")
+  while(pe > tope){
+    //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+    val snr = (3.0+n)/3.0/scala.math.sqrt(n+1)/varnow.doubleValue
+    val snrdb = 10*scala.math.log10(snr)
+    pe = mc.ProbError(n,varnow,petol)
+    varnow = varnow.multiply(varstep)
+    println(snrdb + "\t" + pe + "\t" + mc.numberMomentsLastUsed)
+    mfile.write(snrdb.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+  }
+
+  mfile.close
+
+}
+
+mc.save

### 2011-11-02 plots/runsim.sh

@@ -1,0 +1,1 @@
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar proberrorsim.scala

### 2011-11-02 plots/testmoments.sh

@@ -1,0 +1,1 @@
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar approxmoments.scala

### 2011-11-02 plots/znsim.scala

@@ -1,0 +1,71 @@
+/**
+* Probability of error plots for the lattice Zn.
+*/
+import java.math.BigInteger
+import pubsim.BigRational
+import pubsim.lattices.Zn
+import pubsim.Util
+import pubsim.lattices.util.ProbabilityOfCodingError
+import pubsim.distributions.GaussianNoise
+
+/*
+{
+println
+println("Running Monte-Carlo trials with the Zn lattice")
+
+val lattice = new Zn(1)
+val startdb = 0.0
+val stepdb = 0.5
+val tope = 1e-6
+
+var pe = 1.0
+var dbnow = startdb 
+val efile = new java.io.FileWriter("pemontecarlo1")
+
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val snr = scala.math.pow(10.0,dbnow/10.0)
+  val v = 2.0/3.0/snr
+  pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),1000).probError
+  println(dbnow + "\t" + pe)
+  efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+  dbnow = dbnow + stepdb
+}
+efile.close
+}
+*/
+
+
+{
+val varstart = 1.0/2.0
+val varstep = 9.0/10.0
+val tope = 5e-11
+val petol = 1e-16
+
+var pe = 1.0
+var varnow = varstart 
+val mfile = new java.io.FileWriter("moment1")
+
+val tol = new BigRational(BigInteger.ONE, BigInteger.TEN.pow(15));
+
+println
+println("Caculating 'exact' error probabilty using the error function")
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val r2sdev = new BigRational(scala.math.sqrt(varnow*8.0) ,25)
+  val snr = 2.0/3.0/varnow
+  val snrdb = 10*scala.math.log10(snr)
+  pe = 1.0 - pubsim.Util.erf(r2sdev.reciprocal, tol)
+  varnow = varnow*varstep
+  println(snrdb + "\t" + pe)
+  mfile.write(snrdb.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+}
+mfile.close
+}
+
+
+

### 2011-11-02 paper.tex
Added bib and other missing bits.
Added Emanuele's changes.

@@
 
-\usepackage{../bib/mathbf-abbrevs}
+\usepackage{mathbf-abbrevs}
 
@@
 %opening
-\title{Coding with the Lattice $A_n$}
+\title{On the error performance of the $A_n$ lattices}
 \author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo}
@@
 \begin{abstract}
-Lattices can be used to construct block codes.
-The special family of root lattices has attracted particular attention for this purpose.
-We consider the root lattice $A_n$ and derive formulae for the moments of its Voronoi cell.
-These formulae enable accurate prediction of the performance of codes constructed from $A_n$.
-In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
-These are the densest known sphere packings in dimensions two and three.
-Our results automatically include codes constructed using these packings.
+We consider the root lattice $A_n$ and derive explicit formulae for
+the moments of its Voronoi cell.  We then show that these formulae
+enable accurate prediction of the error probability performance of
+lattice codes constructed from $A_n$.
 \end{abstract}
@@
 \begin{keywords}
-Lattices, lattice decoding, root lattice, probability of error.
+Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
 \end{keywords}
@@
 
-We consider block codes for the Gaussian channel constructed from a subset of points from a Euclidean lattice using \emph{lattice decoding} (as opposed to maximum likelihood decoding)~\cite{Erex2004_lattice_decoding}.
-It is known that such codes can reach the capacity of the Gaussian channel~\cite{Buda1989_some_opt_codes_structure,Erez2005,Erex2004_lattice_decoding}.
-These capacity results make use of randomly generated lattices that do not have any obvious exploitable structure for efficient decoding and encoding~\cite{Jalden2005_sphere_decoding_complexity}.
-For this reason (and many others) the study of specific lattices with known structure is important.
-The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention for this purpose~\cite{SPLAG}.
-In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
-This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.
-Conway and Sloane suggested this type of procedure would be possible in~\cite{Conway1982VoronoiRegions}.
-This paper can be seen as a continuation of that work focusing particularly on the lattice $A_n$.
+The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have
+attracted particular attention as structured codes for the AWGN
+channel~\cite{SPLAG}. The highly symmetric structure of these lattices
+provides the grounds for extremely efficient decoding algorithms.  In
+this paper we consider codes constructed from the root lattice $A_n$
+and derive formulae for accurately predicting the performance of these
+codes.  This is achieved by deriving formulae for the \emph{moments}
+of the \emph{Voronoi cell} of $A_n$.  Conway and Sloane suggested this
+approach to compute the quantizing constants (second order moments) of
+the root lattices,~\cite{Conway1982VoronoiRegions}.  In this paper we
+extend their technique to compute the moments of any order for lattice
+$A_n$.
 
-In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
-These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
-Unfortunately, but not surprisingly, the lattice $A_n$ does not produce good codes in large dimensions.
-We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
+In two dimensions $A_2$ is the hexagonal lattice and in three
+dimensions $A_3$ is the face-centered cubic lattice.  These are the
+densest known sphere packings in dimensions two and three and our
+results automatically include low dimensional codes constructed using
+these packings.  In general, the lattice $A_n$ does not produce
+asymptotically good codes in large dimensions, but still offer a
+significant coding gain in small dimensions.  For these cases, we
+provide a error probability expression which can be computed to any
+degree of accuracy at any finite signal-to-noise ratio.
+%We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
 In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.
 
-This paper is organised as follows.
-In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e.
-\emph{lattice codes}.
-We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.
-Section~\ref{sec:main-result} states the main result without any proof.
-We give recursive formula to compute the moments of the Voronoi cell of $A_n$.
-Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.
-An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
-%This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
-The proof is short and we give it again here so that this paper is self contained.
-In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $n$-dimensional hypercube.
-These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.
-In  Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
-We also plot the results of Monte-Carlo simulations.
-These simulations corroborate with our analytical results.
+This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices
+and codes constructed from them, i.e., \emph{lattice codes}.  We
+describe lattice decoding and show how the probability of coding error
+can be expressed in terms of the moments of the \emph{Voronoi cell} of
+the lattice used.  Section~\ref{sec:main-result} states the main
+result (without proof), which gives recursive formula to compute the
+moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n}
+describes the lattice $A_n$ and some of its properties.  An important
+property for our purposes is that the Voronoi cell of $A_n$ is
+precisely the projection of a $n+1$-dimensional hypercube orthogonal
+to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+ %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
+In Section~\ref{sec:integr-funct-over} we use this property to show
+how integrals over the Voronoi cell of $A_n$ can be expressed as
+integrals over the $n$-dimensional hypercube. These integrals are
+solvable and we use them to obtain the moments of the Voronoi cell in
+Section~\ref{sec:powers-eucl-norm}.  In
+Section~\ref{sec:results-simulations} we plot the probability of error
+versus signal to noise ratio for codes constructed from the lattices
+$A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.  We also plot
+the results of Monte-Carlo to  corroborate our analytical results.
 %An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.
@@
 
-A \term{lattice}\index{lattice}, $\Lambda$, is a set of points in $\reals^m$ such that
+A \term{lattice}\index{lattice}, $\Lambda$, is a set of points in
+$\reals^m$ such that
 \[
@@
  \]
- where $\Bbf$ is an $m \times n$ matrix of rank $n$ called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
-If the generator is square, i.e.
-$m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}.
-If $\Bbf$ has more rows than columns, i.e.
-$m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
-The set of integers $\ints^n$ is a lattice (a generator is the identity matrix) and we call this the~\emph{integer lattice}.
+ where $\Bbf$ is an $m \times n$ matrix of rank $n$ called the
+\term{generator matrix} or \term{basis matrix} or simply generator or
+basis.  If the generator is square, i.e. $m = n$, then the lattice
+points span $\reals^n$ and we say that the lattice is \term{full
+rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the
+lattice points lie in a $n$-dimensional subspace of $\reals^m$. The
+set of integers $\ints^n$ is a lattice (a generator is the identity
+matrix) and we call this the~\emph{integer lattice}.
 
-The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^n$ is the subset of $\reals^n$ containing all points nearer (in Euclidean distance) the lattice point at the origin than any other lattice point.
-The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
-It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\reals^n$.
-That is, the Voronoi cell tessellates space when translated by points in $\Lambda$.
-To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed.
-Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.
-We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.
+The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice
+$\Lambda$ in $\reals^n$ is the subset of $\reals^n$ containing all
+points nearer (in Euclidean distance) the lattice point at the origin
+than any other lattice point. The Voronoi cell is an $n$-dimensional
+convex polytope that is symmetric about the origin.  It is convenient
+to modify this definition of the Voronoi cell slightly so that the
+union of translated Voronoi cells $\cup_{\xbf \in
+\Lambda}\vor(\Lambda) + \xbf$ is equal to  $\reals^n$.  That is, the
+Voronoi cell tessellates space when translated by points in $\Lambda$.
+ To ensure this we require that if a face of $\vor(\Lambda)$ is open,
+then its opposing face is closed. Specifically, if $\xbf \in
+\vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf
+\notin \vor(\Lambda)$.  We wont specifically define which opposing
+face is open and which is closed as the results that follow hold for
+any choice of open and closed opposing faces.
 
-The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, center density, thickness, and the normalized second moment (or quantising constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
-The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.
-There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
-In general these algorithms are only computationally feasible when the dimension is small (approximately $n \leq 9$).
-Even with a complete description of the Voronoi cell it is not necessarily easy to compute the probability of coding error.
+The Voronoi cell encodes many interesting lattice properties such as
+the packing radius, covering radius, kissing number, center density,
+thickness, and the normalized second moment (or quantising
+constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}. The error
+probability of a lattice code can also be evaluated from the Voronoi
+cell as we will see.  There exist algorithms to completely describe
+the Voronoi cell of an arbitrary
+lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
+ In general these algorithms are only computationally feasible when
+the dimension is small (approximately $n \leq 9$).  Even with a
+complete description of the Voronoi cell it is not generally easy to
+compute the probability of error of a lattice in an additive white
+Gaussian noise (AWGN) channel.
 
-The Voronoi cell is linked with the problem of \emph{lattice decoding}.
-Given some point $\ybf \in \reals^n$ a \emph{lattice decoder} (or \emph{nearest lattice point algorithm}) returns the lattice point in $\Lambda$ that is nearest to $\ybf$~\cite{Agrell2002}.
-Equivalently it returns the lattice point $\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$ contains $\ybf$.
-Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}.
-Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small (approximately $n \leq 60$).
-One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.
+The Voronoi cell is linked with the problem of \emph{lattice
+decoding}.  Given some point $\ybf \in \reals^n$ a \emph{lattice
+decoder} (or \emph{nearest lattice point algorithm}) returns the
+lattice point in $\Lambda$ that is nearest to
+$\ybf$~\cite{Agrell2002}.  Equivalently it returns the lattice point
+$\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$
+contains $\ybf$.  Computationally lattice decoding is known to be
+NP-hard under certain conditions when the lattice itself, or rather a
+basis thereof, is considered as an additional input
+parameter~\cite{micciancio_hardness_2001,
+Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist
+that can compute the nearest lattice point in reasonable time if the
+dimension is small (approximately $n \leq 60$). One such algorithm is
+the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.
  %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder. 
 A good overview of these techniques is given by~\cite{Agrell2002}.
-Fast nearest point algorithms are known for specific lattices~\citep{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
-For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e.
-in a number of operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.
+Fast nearest point algorithms are known for specific
+lattices~\cite{Conway1982FastQuantDec, McKilliam2008,
+McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}. For
+example, the root lattices $D_n$ and $A_n$ and their dual lattices
+$D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e. in a number of
+operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.
 
@@
 \newcommand{\calX}{\mathcal X}
-Lattices can be used to construct \emph{block codes}.
-A block code $\calX$ of length $n$ is a finite set of vectors from $\reals^n$.
-Each vector in $\calX$ is called a \emph{codeword}.
-The number of codewords is denoted by $\abs{\calX}$.
-%If each codeword is transmitted with equal probability then the rate of the code is $R = B\frac{1}{n}\log_2\abs{\calX}$ where $B$ is the symbol rate or bandwidth, i.e.
-the rate in Hertz at which the elements (or symbols) in a codeword are transmitted.
-The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the additive white Gaussian channel the received signal takes the form
+Lattices can be used to construct \emph{signal space codes}.  A signal
+space code $\calX$ of dimension $n$ is a finite set of vectors (or
+points) in the Euclidean space $\reals^n$.  Each vector in $\calX$ is
+called a \emph{codeword} and represents a particular signal.  The
+number of codewords is denoted by $\abs{\calX}$.  If each codeword is
+transmitted with equal probability then the rate of the code is $R =
+\frac{1}{n}\log_2\abs{\calX}$ bits per codeword.
+The average power of the code is given by $P =
+\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the AWGN
+channel the received signal takes the form
 \[
-\ybf = \xbf_0 + \wbf
+\ybf = \xbf + \wbf
 \]
-where $\ybf \in \reals^n$, $\xbf_0 \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
-If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
+where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of
+independent and identically distributed Gaussian random variables with
+variance $\sigma^2$.  If the receiver employs maximum likelihood
+decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
 \begin{equation}\label{eq:mldecoder}
-\hat{\xbf} = \min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
+\hat{\xbf} = \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
 \end{equation}
-That is, the receiver computes the codeword in $\calX$ nearest in Euclidean distance to the received signal $\ybf$.
-Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
+That is, the receiver computes the codeword in $\calX$ nearest in
+Euclidean distance to the received signal $\ybf$.  Assuming that each
+codeword is transmitted with equal probability then the probability of
+correct decoding is
 \[
-\sum_{\xbf \in \calX} \frac{\prob( \hat{\xbf} = \xbf )}{\abs{\calX}}.
+P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ).
 \]
-The probability of error is $1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf )$.
+and the probability of error is
+\[
+P_E = 1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob(
+\hat{\xbf} = \xbf ).
+\]
 %The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.
@@
 
-A \emph{lattice code} is a block code with codewords given by a finite subset of points from some lattice $\Lambda$ in $\reals^n$.
-There are an infinity of ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$ called a \emph{shaping region}.
-The lattice points in $\Lambda$ that intersect the shaping region make up the code.
-That is, if $S \subset \reals^n$ is a shaping region then our codewords are given by
+A \emph{lattice code} is a signal space code with codewords taken from
+a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
+There are infinitely many ways to choose a finite subset from a
+lattice, but common approaches make use of a bounded subset of
+$\reals^n$, called a \emph{shaping region} $S \subset \reals^n$ and
 \[
@@
 \]
-Possible choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
-A consequence of the finiteness of a lattice code is that the maximum likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to lattice decoding.
-Computing a nearest lattice point will not in general return a lattice point from the code (the decoded lattice point might lie outside the shaping region).
-A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.
+Possible choices of shaping region are $n$-dimensional spheres,
+spherical shells, hypercubes or the Voronoi cell of a
+\emph{sublattice} of
+$\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
+A consequence of the finiteness of a lattice code is that the maximum
+likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to
+lattice decoding.  Computing a nearest lattice point will not in
+general return a lattice point from the code (the decoded lattice
+point might lie outside the shaping region).
+ %A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.
 This question was answered in the affirmative by Erez and Zamir~\cite{Erex2004_lattice_decoding}.
@@
 Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.
-Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
+Assuming that each codeword $\xbf \in \calX$ has an equal probability
+of being transmitted and that the receiver employs lattice decoding,
+then the average probability of correct decoding in the AWGN channel
+is
+%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
 A result of this is that that observed noise distribution is no longer Gaussian.
@@
 \begin{align}
-P_C &= \frac{1}{\abs{\calX}} \sum_{\cbf\in \calX} \prob( \xbf + \wbf \in \vor(\Lambda) + \xbf )  \\
+P_C &\leq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
+ \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
+\vor(\Lambda) + \xbf )  \\
 &= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
-&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)} e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
+&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
+e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
 \end{align}
-The probability of error is $P_E = 1 - P_C$.  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ we obtain
+The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the
+upper bound is asymptotically tight for large constellations (the
+contribution of the signal set points on near the bounding region
+boarder can be assumed negligible).  By expanding $e^x = 1  + x +
+\frac{x^2}{2} + \dots$ we obtain
 \begin{align}
-P_C &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{ 4\sigma^42!} - \dots d\xbf \nonumber \\
-&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty \frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf.  \label{eq:summomentproberror}
+P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
+- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
+4\sigma^42!} - \dots d\xbf \nonumber \\
+&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
+\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
+d\xbf.  \label{eq:summomentproberror}
 \end{align}
-So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf$ for $m=1,2\dots,M$ for some sufficiently large $M$.
+So, to obtain arbitrarily accurate approximations to the probability
+of error it is enough to know the values of $\int_{\vor(\Lambda)}
+\|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
+The number of terms required will increase as the noise variance gets
+smaller. This implies that the bound with a fixed number of terms is
+not asymptotically tight but is very accurate up to a finite
+signal-to-noise ratio.
 
-In this paper we focus on lattice codes constructed from the family of lattices called $A_n$. We will discover expressions for
+In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
 \[
- C_n(m) = \int_{\vor(A_n)} \left(\|\xbf\|^2\right)^m d\xbf.
+ C_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
 \]
-These can be summed to give arbitrarily accurate approximations for the probability of error.  In this paper we call $C_n(m)$ the $m$th \emph{moment} of the lattice $A_n$.
+These can be summed in (\ref{eq:summomentproberror}) to give arbitrarily accurate approximations for the probability of error.
 
@@
 \end{align*}
-We have tabulated these formula for $m=0$ to $40$.  We derive these results in Section~\ref{sec:powers-eucl-norm}.  First we need some properties of the lattice $A_n$.
+We have explicitly tabulated these formula for $m=0$ to $40$.
+For larger $m$s direct evaluation for specific $n$ from the recursive formula is preferable.
+We will derive these results in Section~\ref{sec:powers-eucl-norm}, but we first need some properties of the lattice $A_n$.
 
@@
 \small
-\bibliography{../bib/bib}
+\bibliography{bib}
 

### 2011-11-02 paper.tex
removed hard newlines.

@@
 \documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal]{../bib/IEEEtran}
+%\documentclass[journal]{IEEEtran}
 %\documentclass[correspondence,9pt]{../bib/IEEEtran}
@@
 \begin{abstract}
-We consider the root lattice $A_n$ and derive explicit formulae for
-the moments of its Voronoi cell.  We then show that these formulae
-enable accurate prediction of the error probability performance of
-lattice codes constructed from $A_n$.
+We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.
+We then show that these formulae enable accurate prediction of the error probability performance of lattice codes constructed from $A_n$.
 \end{abstract}
@@
 
-The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have
-attracted particular attention as structured codes for the AWGN
-channel~\cite{SPLAG}. The highly symmetric structure of these lattices
-provides the grounds for extremely efficient decoding algorithms.  In
-this paper we consider codes constructed from the root lattice $A_n$
-and derive formulae for accurately predicting the performance of these
-codes.  This is achieved by deriving formulae for the \emph{moments}
-of the \emph{Voronoi cell} of $A_n$.  Conway and Sloane suggested this
-approach to compute the quantizing constants (second order moments) of
-the root lattices,~\cite{Conway1982VoronoiRegions}.  In this paper we
-extend their technique to compute the moments of any order for lattice
-$A_n$.
+The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
+The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms.
+In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
+This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.
+Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices,~\cite{Conway1982VoronoiRegions}.
+In this paper we extend their technique to compute the moments of any order for lattice $A_n$.
 
-In two dimensions $A_2$ is the hexagonal lattice and in three
-dimensions $A_3$ is the face-centered cubic lattice.  These are the
-densest known sphere packings in dimensions two and three and our
-results automatically include low dimensional codes constructed using
-these packings.  In general, the lattice $A_n$ does not produce
-asymptotically good codes in large dimensions, but still offer a
-significant coding gain in small dimensions.  For these cases, we
-provide a error probability expression which can be computed to any
-degree of accuracy at any finite signal-to-noise ratio.
+In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
+These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
+In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but still offer a significant coding gain in small dimensions.
+For these cases, we provide a error probability expression which can be computed to any degree of accuracy at any finite signal-to-noise ratio.
 %We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
@@
 
-This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices
-and codes constructed from them, i.e., \emph{lattice codes}.  We
-describe lattice decoding and show how the probability of coding error
-can be expressed in terms of the moments of the \emph{Voronoi cell} of
-the lattice used.  Section~\ref{sec:main-result} states the main
-result (without proof), which gives recursive formula to compute the
-moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n}
-describes the lattice $A_n$ and some of its properties.  An important
-property for our purposes is that the Voronoi cell of $A_n$ is
-precisely the projection of a $n+1$-dimensional hypercube orthogonal
-to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+This paper is organised as follows.
+In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.
+We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.
+Section~\ref{sec:main-result} states the main result (without proof), which gives recursive formula to compute the moments of the Voronoi cell of $A_n$.
+Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.
+An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n+1$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
-In Section~\ref{sec:integr-funct-over} we use this property to show
-how integrals over the Voronoi cell of $A_n$ can be expressed as
-integrals over the $n$-dimensional hypercube. These integrals are
-solvable and we use them to obtain the moments of the Voronoi cell in
-Section~\ref{sec:powers-eucl-norm}.  In
-Section~\ref{sec:results-simulations} we plot the probability of error
-versus signal to noise ratio for codes constructed from the lattices
-$A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.  We also plot
-the results of Monte-Carlo to  corroborate our analytical results.
+In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $n$-dimensional hypercube.
+These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.
+In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
+We also plot the results of Monte-Carlo to  corroborate our analytical results.
 %An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.
@@
 
-A \term{lattice}\index{lattice}, $\Lambda$, is a set of points in
-$\reals^m$ such that
+A \term{lattice}\index{lattice}, $\Lambda$, is a set of points in $\reals^m$ such that
 \[
@@
  \] 
- where $\Bbf$ is an $m \times n$ matrix of rank $n$ called the
-\term{generator matrix} or \term{basis matrix} or simply generator or
-basis.  If the generator is square, i.e. $m = n$, then the lattice
-points span $\reals^n$ and we say that the lattice is \term{full
-rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the
-lattice points lie in a $n$-dimensional subspace of $\reals^m$. The
-set of integers $\ints^n$ is a lattice (a generator is the identity
-matrix) and we call this the~\emph{integer lattice}.
+where $\Bbf$ is an $m \times n$ matrix of rank $n$ called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+If the generator is square, i.e.
+$m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}.
+If $\Bbf$ has more rows than columns, i.e.
+$m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
+The set of integers $\ints^n$ is a lattice (a generator is the identity matrix) and we call this the~\emph{integer lattice}.
 
-The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice
-$\Lambda$ in $\reals^n$ is the subset of $\reals^n$ containing all
-points nearer (in Euclidean distance) the lattice point at the origin
-than any other lattice point. The Voronoi cell is an $n$-dimensional
-convex polytope that is symmetric about the origin.  It is convenient
-to modify this definition of the Voronoi cell slightly so that the
-union of translated Voronoi cells $\cup_{\xbf \in
-\Lambda}\vor(\Lambda) + \xbf$ is equal to  $\reals^n$.  That is, the
-Voronoi cell tessellates space when translated by points in $\Lambda$.
- To ensure this we require that if a face of $\vor(\Lambda)$ is open,
-then its opposing face is closed. Specifically, if $\xbf \in
-\vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf
-\notin \vor(\Lambda)$.  We wont specifically define which opposing
-face is open and which is closed as the results that follow hold for
-any choice of open and closed opposing faces.
+The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^n$ is the subset of $\reals^n$ containing all points nearer (in Euclidean distance) the lattice point at the origin than any other lattice point.
+The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
+It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\reals^n$.
+That is, the Voronoi cell tessellates space when translated by points in $\Lambda$.
+To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed.
+Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.
+We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.
 
-The Voronoi cell encodes many interesting lattice properties such as
-the packing radius, covering radius, kissing number, center density,
-thickness, and the normalized second moment (or quantising
-constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}. The error
-probability of a lattice code can also be evaluated from the Voronoi
-cell as we will see.  There exist algorithms to completely describe
-the Voronoi cell of an arbitrary
-lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
- In general these algorithms are only computationally feasible when
-the dimension is small (approximately $n \leq 9$).  Even with a
-complete description of the Voronoi cell it is not generally easy to
-compute the probability of error of a lattice in an additive white
-Gaussian noise (AWGN) channel.
+The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, center density, thickness, and the normalized second moment (or quantising constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
+The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.
+There exist algorithms to completely describe the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
+In general these algorithms are only computationally feasible when the dimension is small (approximately $n \leq 9$).
+Even with a complete description of the Voronoi cell it is not generally easy to compute the probability of error of a lattice in an additive white Gaussian noise (AWGN) channel.
 
-The Voronoi cell is linked with the problem of \emph{lattice
-decoding}.  Given some point $\ybf \in \reals^n$ a \emph{lattice
-decoder} (or \emph{nearest lattice point algorithm}) returns the
-lattice point in $\Lambda$ that is nearest to
-$\ybf$~\cite{Agrell2002}.  Equivalently it returns the lattice point
-$\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$
-contains $\ybf$.  Computationally lattice decoding is known to be
-NP-hard under certain conditions when the lattice itself, or rather a
-basis thereof, is considered as an additional input
-parameter~\cite{micciancio_hardness_2001,
-Jalden2005_sphere_decoding_complexity}. Nevertheless, algorithms exist
-that can compute the nearest lattice point in reasonable time if the
-dimension is small (approximately $n \leq 60$). One such algorithm is
-the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.
+The Voronoi cell is linked with the problem of \emph{lattice decoding}.
+Given some point $\ybf \in \reals^n$ a \emph{lattice decoder} (or \emph{nearest lattice point algorithm}) returns the lattice point in $\Lambda$ that is nearest to $\ybf$~\cite{Agrell2002}.
+Equivalently it returns the lattice point $\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$ contains $\ybf$.
+Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}.
+Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small (approximately $n \leq 60$).
+One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.
  %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder. 
 A good overview of these techniques is given by~\cite{Agrell2002}.
-Fast nearest point algorithms are known for specific
-lattices~\cite{Conway1982FastQuantDec, McKilliam2008,
-McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}. For
-example, the root lattices $D_n$ and $A_n$ and their dual lattices
-$D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e. in a number of
-operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.
+Fast nearest point algorithms are known for specific lattices~\cite{Conway1982FastQuantDec, McKilliam2008,McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
+For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e.
+in a number of operations of order $O(n)$~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.
 
@@
 \newcommand{\calX}{\mathcal X}
-Lattices can be used to construct \emph{signal space codes}.  A signal
-space code $\calX$ of dimension $n$ is a finite set of vectors (or
-points) in the Euclidean space $\reals^n$.  Each vector in $\calX$ is
-called a \emph{codeword} and represents a particular signal.  The
-number of codewords is denoted by $\abs{\calX}$.  If each codeword is
-transmitted with equal probability then the rate of the code is $R =
-\frac{1}{n}\log_2\abs{\calX}$ bits per codeword.
-The average power of the code is given by $P =
-\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the AWGN
-channel the received signal takes the form
+Lattices can be used to construct \emph{signal space codes}.
+A signal space code $\calX$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.
+Each vector in $\calX$ is called a \emph{codeword} and represents a particular signal.
+The number of codewords is denoted by $\abs{\calX}$.
+If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calX}$ bits per codeword.
+The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.
+In the AWGN channel the received signal takes the form
 \[
@@
 \]
-where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of
-independent and identically distributed Gaussian random variables with
-variance $\sigma^2$.  If the receiver employs maximum likelihood
-decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
+where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
+If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
 \begin{equation}\label{eq:mldecoder}
@@
 \end{equation}
-That is, the receiver computes the codeword in $\calX$ nearest in
-Euclidean distance to the received signal $\ybf$.  Assuming that each
-codeword is transmitted with equal probability then the probability of
-correct decoding is
+That is, the receiver computes the codeword in $\calX$ nearest in Euclidean distance to the received signal $\ybf$.
+Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
 \[
@@
 
-A \emph{lattice code} is a signal space code with codewords taken from
-a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
-There are infinitely many ways to choose a finite subset from a
-lattice, but common approaches make use of a bounded subset of
-$\reals^n$, called a \emph{shaping region} $S \subset \reals^n$ and
+A \emph{lattice code} is a signal space code with codewords taken from a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
+There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$ and
 \[
@@
 \]
-Possible choices of shaping region are $n$-dimensional spheres,
-spherical shells, hypercubes or the Voronoi cell of a
-\emph{sublattice} of
-$\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
-A consequence of the finiteness of a lattice code is that the maximum
-likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to
-lattice decoding.  Computing a nearest lattice point will not in
-general return a lattice point from the code (the decoded lattice
-point might lie outside the shaping region).
+Possible choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
+A consequence of the finiteness of a lattice code is that the maximum likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to lattice decoding.
+Computing a nearest lattice point will not in general return a lattice point from the code (the decoded lattice point might lie outside the shaping region).
  %A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.
@@
 Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.
-Assuming that each codeword $\xbf \in \calX$ has an equal probability
-of being transmitted and that the receiver employs lattice decoding,
-then the average probability of correct decoding in the AWGN channel
-is
+Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
 %\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
@@
 \end{align}
-The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the
-upper bound is asymptotically tight for large constellations (the
-contribution of the signal set points on near the bounding region
-boarder can be assumed negligible).  By expanding $e^x = 1  + x +
-\frac{x^2}{2} + \dots$ we obtain
+The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the contribution of the signal set points on near the bounding region boarder can be assumed negligible).
+By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ we obtain
 \begin{align}
@@
 \end{align}
-So, to obtain arbitrarily accurate approximations to the probability
-of error it is enough to know the values of $\int_{\vor(\Lambda)}
-\|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
-The number of terms required will increase as the noise variance gets
-smaller. This implies that the bound with a fixed number of terms is
-not asymptotically tight but is very accurate up to a finite
-signal-to-noise ratio.
+So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
+The number of terms required will increase as the noise variance gets smaller.
+This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
 

### 2011-11-02 paper.tex
minor edits

@@
 Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices,~\cite{Conway1982VoronoiRegions}.
-In this paper we extend their technique to compute the moments of any order for lattice $A_n$.
+In this paper we extend their technique to compute the moments of any order for $A_n$.
 
@@
 These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
-In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but still offer a significant coding gain in small dimensions.
-For these cases, we provide a error probability expression which can be computed to any degree of accuracy at any finite signal-to-noise ratio.
+In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but does offer a coding gain in small dimensions.
+For these cases, we provide an error probability expression that can be computed to any degree of accuracy at any finite signal-to-noise ratio.
 %We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
@@
 In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
-We also plot the results of Monte-Carlo to  corroborate our analytical results.
+We also plot the results of Monte-Carlo simulations that support our analytical results.
 %An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.
@@
 
-A \term{lattice}\index{lattice}, $\Lambda$, is a set of points in $\reals^m$ such that
+A \term{lattice}\index{lattice}, $\Lambda$, is a discrete set of points in $\reals^m$ such that
 \[
@@
  \] 
-where $\Bbf$ is an $m \times n$ matrix of rank $n$ called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+where $\Bbf \in \reals^{m\times n}$ is an $m \times n$ matrix of rank $n$ called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
 If the generator is square, i.e.
@@
  \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
-\vor(\Lambda) + \xbf )  \\
+\vor(\Lambda) + \xbf )  \nonumber \\
 &= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
@@
 \end{align}
-The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the contribution of the signal set points on near the bounding region boarder can be assumed negligible).
+The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).
 By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ we obtain
@@
 
-We now state our main result.  The $m$th moment $C_n(m)$ satisfies
+We now state our main result.  The $m$th moment $C_n(m)$ of the lattice $A_n$ satisfies
 \begin{equation}\label{eq:theCmformula}
@@
 We have explicitly tabulated these formula for $m=0$ to $40$.
-For larger $m$s direct evaluation for specific $n$ from the recursive formula is preferable.
-We will derive these results in Section~\ref{sec:powers-eucl-norm}, but we first need some properties of the lattice $A_n$.
+For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.
+We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.
 

### 2011-11-10 plots/proberrorsim.scala
Added to moment store.  Now goes up to n = 5.

@@ -6,12 +6,12 @@ import pubsim.lattices.An.AnFastSelect
 import pubsim.lattices.An.MomentComputer
 import pubsim.BigRational
 
-val ns = List(2,3,4)
+val ns = List(2,3,4,5)
 //val lattice = new AnFastSelect(n)
 val varstart = new BigRational(1,2)
 val varstep = new BigRational(9,10)
-val tope = 1e-11
-val petol = 1e-16
+val tope = 1e-12
+val petol = 1e-17
 
 print("Reading existing moment data, this might take a moment....")
 val mc = new MomentComputer("momentstore")

### 2011-11-11 paper.tex
Ramanan: Polished lattice/voronoi cell definitions for consistency

--- 2011-11-02 paper.tex
@@
 \documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal]{IEEEtran}
+%\documentclass[journal]{../bib/IEEEtran}
 %\documentclass[correspondence,9pt]{../bib/IEEEtran}
@@
 \usepackage{mathbf-abbrevs}
+\usepackage{amsmath}
+\DeclareMathOperator{\argmin}{\arg\!\min}
 
@@
 %opening
-\title{On the error performance of the $A_n$ lattices}
+\title{Coding with the Lattice $A_n$}
 \author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo}
@@
 \begin{abstract}
-We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.
-We then show that these formulae enable accurate prediction of the error probability performance of lattice codes constructed from $A_n$.
+Lattices can be used to construct block codes.
+The special family of root lattices has attracted particular attention for this purpose.
+We consider the root lattice $A_n$ and derive formulae for the moments of its Voronoi cell.
+These formulae enable accurate prediction of the performance of codes constructed from $A_n$.
+In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
+These are the densest known sphere packings in dimensions two and three.
+Our results automatically include codes constructed using these packings.
 \end{abstract}
@@
 \begin{keywords}
-Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
+Lattices, lattice decoding, root lattice, probability of error.
 \end{keywords}
@@
 
-The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
-The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms.
+We consider block codes for the Gaussian channel constructed from a subset of points from a Euclidean lattice using \emph{lattice decoding} (as opposed to maximum likelihood decoding)~\cite{Erex2004_lattice_decoding}.
+It is known that such codes can reach the capacity of the Gaussian channel~\cite{Buda1989_some_opt_codes_structure,Erez2005,Erex2004_lattice_decoding}.
+These capacity results make use of randomly generated lattices that do not have any obvious exploitable structure for efficient decoding and encoding~\cite{Jalden2005_sphere_decoding_complexity}.
+For this reason (and many others) the study of specific lattices with known structure is important.
+The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention for this purpose~\cite{SPLAG}.
 In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
 This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.
-Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices,~\cite{Conway1982VoronoiRegions}.
-In this paper we extend their technique to compute the moments of any order for lattice $A_n$.
+Conway and Sloane suggested this type of procedure would be possible in~\cite{Conway1982VoronoiRegions}.
+This paper can be seen as a continuation of that work focusing particularly on the lattice $A_n$.
 
@@
 These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
-In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but still offer a significant coding gain in small dimensions.
-For these cases, we provide a error probability expression which can be computed to any degree of accuracy at any finite signal-to-noise ratio.
-%We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
+Unfortunately, but not surprisingly, the lattice $A_n$ does not produce good codes in large dimensions.
+We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
 In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.
@@
 This paper is organised as follows.
-In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.
+In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e.
+\emph{lattice codes}.
 We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.
-Section~\ref{sec:main-result} states the main result (without proof), which gives recursive formula to compute the moments of the Voronoi cell of $A_n$.
+Section~\ref{sec:main-result} states the main result without any proof.
+We give a recursive formula to compute the moments of the Voronoi cell of $A_n$.
 Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.
-An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n+1$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
- %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
+An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+%This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+The proof is short and we give it again here so that this paper is self contained.
 In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $n$-dimensional hypercube.
@@
 In  Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
-We also plot the results of Monte-Carlo to  corroborate our analytical results.
+We also plot the results of Monte-Carlo simulations.
+These simulations corroborate with our analytical results.
 %An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.
@@
 
-A \term{lattice}\index{lattice}, $\Lambda$, is a set of points in $\reals^m$ such that
+A \term{lattice}\index{lattice}, $\Lambda$, is a countable subset of points in $\reals^m$ such that
 \[
@@
  \]
-where $\Bbf$ is an $m \times n$ matrix of rank $n$ called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+ for a certain rank-$n$ $m \times n$ matrix $\Bbf$ of real elements, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+Since the generator matrix has rank $n$, the lattice $\Lambda$ is termed $n$-dimensional.
 If the generator is square, i.e.
@@
 $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
-The set of integers $\ints^n$ is a lattice (a generator is the identity matrix) and we call this the~\emph{integer lattice}.
+For any lattice $\Lambda$ with an $m \times n$ generator matrix, we also define $\mathcal{S}_{\Lambda}$ to be the sub-space spanned by the columns of the generator matrix.
+The set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
 
-The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ in $\reals^n$ is the subset of $\reals^n$ containing all points nearer (in Euclidean distance) the lattice point at the origin than any other lattice point.
-The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
-It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\reals^n$.
+The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\mathcal{S}_{\Lambda}$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point.
+It can be shown that the Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
+It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.
 That is, the Voronoi cell tessellates space when translated by points in $\Lambda$.
@@
 The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.
-There exist algorithms to completely describe the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
+There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
 In general these algorithms are only computationally feasible when the dimension is small (approximately $n \leq 9$).
-Even with a complete description of the Voronoi cell it is not generally easy to compute the probability of error of a lattice in an additive white Gaussian noise (AWGN) channel.
+Even with a complete description of the Voronoi cell it is not necessarily easy to compute the probability of coding error.
 
@@
 A good overview of these techniques is given by~\cite{Agrell2002}.
-Fast nearest point algorithms are known for specific lattices~\cite{Conway1982FastQuantDec, McKilliam2008,McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
+Fast nearest point algorithms are known for specific lattices~\citep{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
 For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e.
@@
 \newcommand{\calX}{\mathcal X}
-Lattices can be used to construct \emph{signal space codes}.
-A signal space code $\calX$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.
-Each vector in $\calX$ is called a \emph{codeword} and represents a particular signal.
+Lattices can be used to construct \emph{block codes}.
+A block code $\calX$ of length $n$ is a finite set of vectors from $\reals^n$.
+Each vector in $\calX$ is called a \emph{codeword}.
 The number of codewords is denoted by $\abs{\calX}$.
-If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calX}$ bits per codeword.
-The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.
-In the AWGN channel the received signal takes the form
+%If each codeword is transmitted with equal probability then the rate of the code is $R = B\frac{1}{n}\log_2\abs{\calX}$ where $B$ is the symbol rate or bandwidth, i.e.
+the rate in Hertz at which the elements (or symbols) in a codeword are transmitted.
+The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.  In the additive white Gaussian channel the received signal takes the form
 \[
-\ybf = \xbf + \wbf
+\ybf = \xbf_0 + \wbf
 \]
-where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
+where $\ybf \in \reals^n$, $\xbf_0 \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
 If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
 \begin{equation}\label{eq:mldecoder}
-\hat{\xbf} = \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
+\hat{\xbf} = \argmin_{\xbf \in \calX} \| \ybf - \xbf \|^2.
 \end{equation}
@@
 \[
-P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ).
+\sum_{\xbf \in \calX} \frac{\prob( \hat{\xbf} = \xbf )}{\abs{\calX}}.
 \]
-and the probability of error is
-\[
-P_E = 1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob(
-\hat{\xbf} = \xbf ).
-\]
+The probability of error is $1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf )$.
 %The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.
@@
 
-A \emph{lattice code} is a signal space code with codewords taken from a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
-There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$ and
+A \emph{lattice code} is a block code with codewords given by a finite subset of points from some lattice $\Lambda$ in $\reals^n$.
+There are innumerable ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$ called a \emph{shaping region}.
+The lattice points in $\Lambda$ that intersect the shaping region make up the code.
+That is, if $S \subset \reals^n$ is a shaping region then our codewords are given by
 \[
@@
 Computing a nearest lattice point will not in general return a lattice point from the code (the decoded lattice point might lie outside the shaping region).
- %A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.
+A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.
 This question was answered in the affirmative by Erez and Zamir~\cite{Erex2004_lattice_decoding}.
@@
 Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.
-Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
-%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
+Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
 A result of this is that that observed noise distribution is no longer Gaussian.
@@
 \begin{align}
-P_C &\leq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
- \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
-\vor(\Lambda) + \xbf )  \\
+P_C &= \frac{1}{\abs{\calX}} \sum_{\cbf\in \calX} \prob( \xbf + \wbf \in \vor(\Lambda) + \xbf )  \\
 &= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
-&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
-e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
+&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)} e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
 \end{align}
-The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the contribution of the signal set points on near the bounding region boarder can be assumed negligible).
-By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ we obtain
+The probability of error is $P_E = 1 - P_C$.  By expanding the exponential in terms of its McLaurin's series, we obtain
+%Removed $e^x = 1  + x + \frac{x^2}{2} + \dots$ and simply say "McLaurin's series"
 \begin{align}
-P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
-- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
-4\sigma^42!} - \dots d\xbf \nonumber \\
-&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
-\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
-d\xbf.  \label{eq:summomentproberror}
+P_C &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{ 4\sigma^42!} - \dots d\xbf \nonumber \\
+&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty \frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf.  \label{eq:summomentproberror}
 \end{align}
-So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
-The number of terms required will increase as the noise variance gets smaller.
-This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
+Hence it is enough to know the values of $\int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf$ for $m=1,2\dots,M$ for some sufficiently large $M$, in order to obtain arbitrarily accurate approximations to the probability of error.
 
-In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
+In this paper we focus on lattice codes constructed from the family of lattices called $A_n$. We will discover expressions for
 \[
- C_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
+ C_n(m) = \int_{\vor(A_n)} \left(\|\xbf\|^2\right)^m d\xbf.
 \]
-These can be summed in (\ref{eq:summomentproberror}) to give arbitrarily accurate approximations for the probability of error.
+These can be summed to give arbitrarily accurate approximations for the probability of error.  In this paper we call $C_n(m)$ the $m^{\text{th}}$ \emph{moment} of the lattice $A_n$.
 
@@
 \end{align*}
-We have explicitly tabulated these formula for $m=0$ to $40$.
-For larger $m$s direct evaluation for specific $n$ from the recursive formula is preferable.
-We will derive these results in Section~\ref{sec:powers-eucl-norm}, but we first need some properties of the lattice $A_n$.
+We have tabulated these formula for $m=0$ to $40$.  We derive these results in Section~\ref{sec:powers-eucl-norm}.  First we need some properties of the lattice $A_n$.
 
@@
 Any vector in $H$ has the property that the sum (and therefore the mean) of its elements is zero and for this reason $H$ is often referred to as the \emph{zero-sum plane} or the \emph{zero-mean plane}.
-The lattice $A_n$ is the intersection of the integer lattice $\ints^{n+1}$ with $H$, that is
+The lattice $A_n$ is the intersection of the integer lattice $\ints^{n+1}$ with the zero-sum plane, that is
 \begin{equation}

### 2011-11-11 paper.tex
Added to paper, merges Emmanuele's and Ramanan's edits.  This needs me to go over it next week with Vaughan.  It's good, but I can see how it is still hard for people.

--- 2011-11-02 paper.tex
@@
 
-A \term{lattice}\index{lattice}, $\Lambda$, is a discrete set of points in $\reals^m$ such that
+A \term{lattice}\index{lattice}, $\Lambda$, is a countable subset of points in $\reals^m$ such that
 \[
@@
  \]
-where $\Bbf \in \reals^{m\times n}$ is an $m \times n$ matrix of rank $n$ called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+ for a certain rank-$n$ $m \times n$ matrix $\Bbf$ of real elements, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+Since the generator matrix has rank $n$, the lattice $\Lambda$ is termed $n$-dimensional.
 If the generator is square, i.e.
@@
 \]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $C_m(n)$ in terms of $n$.  The first four such formula are:
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $C_m(n)$ in terms of $n$.  The first five such formula are:
 \begin{align*}
@@
 C_2(n) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
-C_3(n) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
-%C_4(n) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
+C_3(n) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
+C_4(n) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
 \end{align*}
@@
 % \]
-The lattice has determinant $n+1$ so the volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$.
-The lattice has $n(n+1)$ minimal vectors, each of norm $2$, so the packing radius is $\frac{1}{\sqrt{2}}$~\cite[p.
-108]{SPLAG}.
+The volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$~\cite[p. 108]{SPLAG}.
 
@@
 \begin{lemma} \label{lem:QVorZnsubsetVorAn}
-The projection of $\vor(\ints^{n+1})$ into H is a subset of $\vor(A_{n}) \cap H$.  That is,
+The projection of $\vor(\ints^{n+1})$ into H is a subset of $\vor(A_{n})$.  That is,
 \[
@@
 By definition \eqref{eq:An_sub_Zn} $\xbf'\onebf = 0$ so $\|\xbf - \ybf\|^2 < \|\ybf\|^2$.
-This violates that $\ybf \in \vor(\ints^{n+1})$ and hence $\Qbf\ybf \in \vor(A_n)$.\footnote{This proof can be generalised to show that for any lattice $L$ and hyperplane $P$ such that $P\cap L$ is also a lattice it is true that $p\vor(L) \subseteq \vor(L \cap P) \cap P$ where $p$ indicates the orthogonal projection into $P$~\cite[Lemma~2.1]{McKilliam2010thesis}.}
+This violates that $\ybf \in \vor(\ints^{n+1})$ and hence $\Qbf\ybf \in \vor(A_n)$.\footnote{This proof can be generalised to show that for any lattice $L$ and hyperplane $P$ such that $P\cap L$ is also a lattice it is true that $p\vor(L) \subseteq \vor(L \cap P)$ where $p$ indicates the orthogonal projection into $P$~\cite[Lemma~2.1]{McKilliam2010thesis}.}
 \end{IEEEproof}
@@
 \begin{theorem}  \label{thm:VorAn=QVorZn1}
-The projection of $\vor(\ints^{n+1})$ into H is equal to $H \cap \vor(A_{n})$. That is,
+The projection of $\vor(\ints^{n+1})$ into H is equal to $\vor(A_{n})$. That is,
 \[
-H \cap \vor(A_n) = \Qbf\vor(\ints^{n+1}).
+\vor(A_n) = \Qbf\vor(\ints^{n+1}).
 \]
@@
 \begin{proof}
-The $n$-volume of $H \cap \vor(A_n)$ is given by the square root of the determinant of $A_n$, that is 
-\[
-\sqrt{\det{A_n}} = \sqrt{n + 1}.
-\]  
+The $n$-volume of $\vor(A_n)$ is given $\sqrt{n + 1}$.
 From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to $\sqrt{n + 1}$ also.
-It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq H \cap \vor(A_n)$, so, because the volumes are the same, and because $H \cap \vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $H \cap \vor(A_n) = \Qbf\vor(\ints^{n+1})$.
+It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
 \end{proof}
@@
 It is not immediately clear how an integral over $\vor(A_n)$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
-\[
+\begin{equation}\label{eq:unormalised}
 \int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
-\]
+\end{equation}
 This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.
 To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.
-Then $\Qbf\xbf = \Qbf(\xbf + k \onebf)$ for all $k \in \reals$ such that $\xbf +  k \onebf \in \vor(\ints^{n+1})$, i.e.
-for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.
+Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.
 The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by
@@
 \]
-The integral~\eqref{eq:vorfint} can now be written as
+The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
 \[
@@
 \]
-The primary advantage of this integral is that the bounds of are given by the hypercube $\vor(\ints^{n+1})$.
+The primary advantage of this integral is that the bounds are given by the hypercube $\vor(\ints^{n+1})$.
 
@@
 \]
-and where $A = \sum_{i=1}^{n-1}w_i^2$ and $B = \sum_{i=1}^{n-1}w_i$.  Now,
+and where,
+\[
+ A = \sum_{i=1}^{n-1}w_i^2 \qquad \text{and} \qquad B = \sum_{i=1}^{n-1}w_i.
+\]
+Now,
 \[
@@
 In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice and our results include codes constructed using these packings as a special case.
-The lattice $A_n$ does not produce good codes in large dimension, but we are hopeful that the techniques developed here can be applied to other lattices.
+%The lattice $A_n$ does not produce good codes in large dimension, but we are hopeful that the techniques developed here can be applied to other lattices.
 In particular similar techniques might apply to the family Coxeter lattices, which contains the root lattice $E_8$.

### 2011-11-11 paper.tex
Added back Emanuele's abstract, title and keywords.

@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal]{../bib/IEEEtran}
-%\documentclass[correspondence,9pt]{../bib/IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal]{IEEEtran}
+%\documentclass[correspondence,9pt]{IEEEtran}
 
@@
 
+Given a lattice with inradius $\rho$ and Voronoi cell having volume $V$ we define the SNR as
+\[
+\frac{n^2 \sigma^2 V_n}{(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
+\]
+where $V_n$ is the volume of the $n$ dimensional hypersphere and the logarithms are base $2$.
+
 Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $C_n(m)$ and~\eqref{eq:summomentproberror} (solid line).
@@
 		\includegraphics{plots/peplots-1.mps}
-		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$, $A_4$, $E_8$ and the Leech lattice $\Lambda_{24}$.}
+		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$, $A_4$, and $A_5$ and $E_8$ and the Leech lattice $\Lambda_{24}$.}
 		\label{fig:peplots}

### 2011-11-12 paper.tex
Fixed some notational inconsistencies.  This would have made things hard to follow.  This really needs another read, and I need to fix the simulations.

@@
 The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
-The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms.
+The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms\cite{,Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
 In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
@@
 These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.
-In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
+In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$, $A_4$ and $A_5$.
 We also plot the results of Monte-Carlo simulations that support our analytical results.
@@
 
-A \term{lattice}\index{lattice}, $\Lambda$, is a countable subset of points in $\reals^m$ such that
+A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
 \[
@@
  \]
- for a certain rank-$n$ $m \times n$ matrix $\Bbf$ of real elements, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+ where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
 Since the generator matrix has rank $n$, the lattice $\Lambda$ is termed $n$-dimensional.
@@
 $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
-For any lattice $\Lambda$ with an $m \times n$ generator matrix, we also define $\mathcal{S}_{\Lambda}$ to be the sub-space spanned by the columns of the generator matrix.
+For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the sub-space spanned by the columns of the generator matrix.
 The set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
@@
 The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).
-By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ we obtain
+By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its MacLauren series we obtain
 \begin{align}
@@
 \[
-\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
+\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{\onebf^\prime \onebf} = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
 \]
 the projection matrix orthogonal to $\onebf$ (i.e.
-into $H$) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.
+into the zero-sum plane) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.
 Given a set $S$ of vectors from $\reals^{n+1}$ we write $\Qbf S$  to denote the set with elements $\Qbf \sbf$ for all $\sbf \in S$, i.e.
@@
 \begin{lemma} \label{lem:QVorZnsubsetVorAn}
-The projection of $\vor(\ints^{n+1})$ into H is a subset of $\vor(A_{n})$.  That is,
+The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is a subset of $\vor(A_{n})$.  That is,
 \[
@@
 \begin{theorem}  \label{thm:VorAn=QVorZn1}
-The projection of $\vor(\ints^{n+1})$ into H is equal to $\vor(A_{n})$. That is,
+The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is equal to $\vor(A_{n})$. That is,
 \[
@@
 \begin{proof}
-The $n$-volume of $\vor(A_n)$ is given $\sqrt{n + 1}$.
-From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to $\sqrt{n + 1}$ also.
+The $n$-volume of $\vor(A_n)$ is given $\sqrt{n + 1}$.  From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to 
+\[
+\sum_{i = 1}^{n+1} \frac{\dotprod{\onebf}{\ebf_i}}{\|\onebf\|} =  \sum_{i = 1}^{n+1} \frac{1}{\sqrt{n+1}} = \sqrt{n + 1}
+\] 
+also.
 It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
@@
 \end{equation}
-Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector $[t, 0, w_1, w_2, \dots, w_{n-1}]^\prime$.
+Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector
+\[
+\wbf = [t, 0, w_1, w_2, \dots, w_{n-1}]^\prime.
+\]
 Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.
@@
 &= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n+1}w_i + \left(\sum_{i=1}^{n+1}w_i \right)^2 \right)\\
-&= c + d,
+&= C + D,
 \end{align*}
@@
 \[
-c = \left(\frac{n}{n+1} \right) t^2,\qquad \text{and} \qquad d = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
+C = \left(\frac{n}{n+1} \right) t^2 \qquad \text{and} \qquad D = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
 \]
@@
 \[
- \frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(c+d)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
+ \frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(C+D)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
 \]
-and by expanding the binomial $(c+d)^m$ we get
+and by expanding the binomial $(C+D)^m$ we get
 \[
-\frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} c^{m-k} \int_0^t \cdots \int_0^t \, d^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
+\frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} C^{m-k} \int_0^t \cdots \int_0^t \, D^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
 \]
-Expanding $d^k$ as a trinomial gives
+Expanding $D^k$ as a trinomial gives
 \begin{align*}
-d^k &= \sum_{k_1+k_2+k_3=k} \frac{k! A^{k_1} B^{2k_3 + k_2}}{k_1! k_2! k_3!} \left(\frac{-1}{n+1}\right)^{k_2+k_3}2^{k_2}t^{k_2}  \\
+D^k &= \sum_{k_1+k_2+k_3=k} \frac{k! A^{k_1} B^{2k_3 + k_2}}{k_1! k_2! k_3!} \left(\frac{-1}{n+1}\right)^{k_2+k_3}2^{k_2}t^{k_2}  \\
 &=  \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{k!  A^{a} B^{2k - 2a - b}}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}2^{b}t^{b}
@@
 \]
-where $G(n,k,m)$ satisfies the recursion given by~\eqref{eq:theGrecursion}.  So, let $D$ satisfy
+where $G(n,c,d)$ satisfies the recursion given by~\eqref{eq:theGrecursion}.  So, let $P$ satisfy
 \begin{align*}
-D &= t^{1 - n - 2k}\int_0^t \cdots \int_0^t d^k dw_{n-1} \, \dots \, dw_1 \\
+P &= t^{1 - n - 2k}\int_0^t \cdots \int_0^t D^k dw_{n-1} \, \dots \, dw_1 \\
 &= \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{2^{b} k!G(n-1,a,2k - 2a - b)}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}.
 \end{align*}
-Now $c^{m-k} = \left(\frac{n}{n+1} \right)^{m-k} t^{2(m-k)}$ and
+Now $C^{m-k} = \left(\frac{n}{n+1} \right)^{m-k} t^{2(m-k)}$ and
 \begin{align*}
-\frac{C_m(n)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} D \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy. \\
-&= \frac{1}{n+2m} \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} D.
+\frac{C_m(n)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy \\
+&= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \beta(y,n+2m,0) dy \\
+&= \frac{1}{n+2m} \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P.
 \end{align*}
@@
 \[
-F(n,c,d) = \int^{t}_{0} \cdots \int^{t}_{0} \left(\sum_{j=1}^{n} x_j^2\right)^c \left( \sum_{i=1}^{n} x_i \right)^d \,dx_1\cdots dx_{n}.
+F(n-1,a,2k - 2a - b) = \int^{t}_{0} \cdots \int^{t}_{0} A^a B^{2k - 2a - b} \,dw_{n-1}\cdots dw_{1},
 \]
-We shall find a recursion describing this integral.  Write
+or equivalently, integrals of the form
+\[
+F(n,c,d) = \int^{t}_{0} \cdots \int^{t}_{0} \left(\sum_{j=1}^{n} x_j^2\right)^c \left( \sum_{i=1}^{n} x_i \right)^d \,dx_1\cdots dx_{n}
+\]
+where $n,c$ and $d$ are integers.  We shall find a recursion describing this integral.  Write
 \[

### 2011-11-15 paper.tex
minor edits to plots.
modified the plot data so that things are more visible.

@@
 %\documentclass[a4paper,10pt]{article}
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal]{IEEEtran}
+\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+%\documentclass[journal]{IEEEtran}
 %\documentclass[correspondence,9pt]{IEEEtran}
@@
 \title{On the error performance of the $A_n$ lattices}
-\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo}
+\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson}
 
@@
 
-Consider a function $f : \reals^{n+1} \mapsto \reals$.  We would like to be able to solve integrals of the form
+We would like to be able to integrate functions over the Voronoi cell of $A_n$.
+The definition of $A_n$ above places it in the $n$ dimensional zero-sum plane, lying in $\reals^{n+1}$.
+The Voronoi cell is a subset of the zero-sum plane and has zero $n+1$-dimensional volume.
+By an appropriate change of variables it would be possible to write the Voronoi cell in an $n$-dimensional coordinate system, and then integrate.
+However, we find the following approach simpler.
+Given a set $S$ let $S \times \onebf$ denote the set of elements that can be writen as $x + y$ where $x \in \vor(A_n)$ and $y = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
+Consider a function $f : \reals^{n+1} \mapsto \reals$.
+We would like to be able to solve integrals of the form
 \begin{equation}\label{eq:vorfint}
-\int_{\vor(A_n)} f(\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1})} f(\xbf) d\xbf.
+\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
 \end{equation}
-It is not immediately clear how an integral over $\vor(A_n)$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
+It is not immediately clear how an integral over $\vor(A_n) \times \onebf$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
 \begin{equation}\label{eq:unormalised}
@@
 \[
-\frac{n^2 \sigma^2 V_n}{(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
+\frac{V_n}{\sigma^2(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
 \]

### 2011-11-15 plots/e8sim.scala

@@ -97,4 +97,3 @@ while(pe > tope){
 mfile.close
 }
 
-

### 2011-11-15 paper.tex
More edits to An paper.

@@
 the moments of its Voronoi cell.  We then show that these formulae
-enable accurate prediction of the error probability performance of
+enable accurate prediction of the error probability of
 lattice codes constructed from $A_n$.
@@
 The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
-The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms\cite{,Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
+The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
 In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
@@
 These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.
-In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$, $A_4$ and $A_5$.
+In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
 We also plot the results of Monte-Carlo simulations that support our analytical results.
@@
 
-The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, center density, thickness, and the normalized second moment (or quantising constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
+The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, minimal vectors, center density, thickness, and the normalized second moment (or quantizing constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
 The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.
@@
  %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder. 
-A good overview of these techniques is given by~\cite{Agrell2002}.
+A good overview of these techniques is given by Agrell~et.~al.~\cite{Agrell2002}.
 Fast nearest point algorithms are known for specific lattices~\cite{Conway1982FastQuantDec, McKilliam2008,McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
@@
 \[
-P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ).
+P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ),
 \]
@@
 The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).
-By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its MacLauren series we obtain
+By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
 \begin{align}
@@
 
+\newcommand{\calM}{M}
 In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
 \[
- C_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
+ \calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
 \]
@@
 
-We now state our main result.  The $m$th moment $C_n(m)$ of the lattice $A_n$ satisfies
+We now state our main result.  The $m$th moment $\calM_n(m)$ of the lattice $A_n$ satisfies
 \begin{equation}\label{eq:theCmformula}
-\frac{C_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)} 
+\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)} 
 \end{equation}
@@
 \]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $C_m(n)$ in terms of $n$.  The first five such formula are:
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
 \begin{align*}
-C_0(n) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
-C_1(n) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
-C_2(n) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
-C_3(n) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
-C_4(n) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
+\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
+\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
+\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
+\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
+\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
 \end{align*}
@@
 \label{eq:An_sub_Zn}
-  A_{n} = \ints^{n+1} \cap H = \big\{ \xbf \in \ints^{n+1} \mid \xbf\cdot\onebf = 0  \big\}.
+  A_{n} = \ints^{n+1} \cap H = \big\{ \xbf \in \ints^{n+1} \mid \dotprod{\xbf}{\onebf} = 0  \big\}.
 \end{equation}
@@
 % \]
-The volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$~\cite[p. 108]{SPLAG}.
+The lattice has $n(n+1)$ minimal vectors, each of squared Euclidean length $2$, so the packing radius is $\frac{1}{\sqrt{2}}$.
+The $n$-volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$~\cite[p.
+108]{SPLAG}.
 
@@
 \begin{proof}
-The $n$-volume of $\vor(A_n)$ is given $\sqrt{n + 1}$.  From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to 
+Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.
+The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.
+From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
 \[
@@
 We would like to be able to integrate functions over the Voronoi cell of $A_n$.
-The definition of $A_n$ above places it in the $n$ dimensional zero-sum plane, lying in $\reals^{n+1}$.
-The Voronoi cell is a subset of the zero-sum plane and has zero $n+1$-dimensional volume.
-By an appropriate change of variables it would be possible to write the Voronoi cell in an $n$-dimensional coordinate system, and then integrate.
+Consider a function $f : \reals^{n+1} \mapsto \reals$.
+The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.
+The Voronoi cell is a subset of the zero-sum plane that has zero $n+1$-dimensional volume.
+So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.
+This is not what we intend.
+By an appropriate change of variables it would be possible to write the Voronoi cell $\vor(A_n)$ in an $n$-dimensional coordinate system, and then integrate.
 However, we find the following approach simpler.
-Given a set $S$ let $S \times \onebf$ denote the set of elements that can be writen as $x + y$ where $x \in \vor(A_n)$ and $y = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
-Consider a function $f : \reals^{n+1} \mapsto \reals$.
-We would like to be able to solve integrals of the form
+Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $x + y$ where $x \in S$ and $y = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
+Now, the integral over the Voronoi cell can be written as
 \begin{equation}\label{eq:vorfint}
@@
 \[
-\int_{\vor(A_n)} f(\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
+\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
 \]
@@
 
-We now derive expressions for the $C_n(m)$.  Setting $f(\Qbf\xbf) = \left(\|\Qbf\wbf\|^{2}\right)^m$ in~\eqref{eq:intwewilluse} we obtain,
+We now derive expressions for the $\calM_n(m)$.  Setting $f(\Qbf\xbf) = \left(\|\Qbf\wbf\|^{2}\right)^m$ in~\eqref{eq:intwewilluse} we obtain,
 \[
-\frac{C_m(n)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
+\frac{\calM_n(m)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
 \]
@@
 \[
- \frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(C+D)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
+ \frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(C+D)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
 \]
@@
 \[
-\frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} C^{m-k} \int_0^t \cdots \int_0^t \, D^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
+\frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} C^{m-k} \int_0^t \cdots \int_0^t \, D^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
 \]
@@
 In Appendix~\ref{sec:mult-type-integr} we show that the integral of $A^{a} B^{2k - 2a - b}$ over $w_1,\dots w_{n-1}$ is
-\[
+\begin{equation} \label{eq:recurseint}
 \int_0^t \cdots \int_0^t A^{a} B^{2k-2a-b} dw_{n-1} \, \dots \, dw_1 = t^{n-1+2k - b} G(n-1,a,2k - 2a - b).
-\]
+\end{equation}
 where $G(n,c,d)$ satisfies the recursion given by~\eqref{eq:theGrecursion}.  So, let $P$ satisfy
@@
 \begin{align*}
-\frac{C_m(n)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy \\
+\frac{\calM_n(m)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy \\
 &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \beta(y,n+2m,0) dy \\
@@
 the Voronoi cell of $A_n$ scaled by a factor of $2$.
-The average power of these constellations can be approximated by four times the second moment $4C_1(m)$.
+The average power of these constellations can be approximated by four times the second moment $4\calM_n(1)$.
 This approximation is conservative as Voronoi constellations can have average power smaller than the second moment of the shaping region~\cite{Conway1983VoronoiCodes}.
@@
 \[
-\text{SNR} = \frac{4C_1(n)}{n \sigma^2} = \frac{n+3}{3\sqrt{n+1}\sigma^2}
+\text{SNR} = \frac{4\calM_n(1)}{n \sigma^2} = \frac{n+3}{3\sqrt{n+1}\sigma^2}
 \]
@@
 
-Given a lattice with inradius $\rho$ and Voronoi cell having volume $V$ we define the SNR as
-\[
-\frac{V_n}{\sigma^2(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
-\]
-where $V_n$ is the volume of the $n$ dimensional hypersphere and the logarithms are base $2$.
+%Given a lattice with inradius $\rho$ and Voronoi cell having volume $V$ we define the SNR as
+%\[
+%\frac{V_n}{\sigma^2(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
+%\]
+%where $V_n$ is the volume of the $n$ dimensional hypersphere and the logarithms are base $2$.
 
-Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $C_n(m)$ and~\eqref{eq:summomentproberror} (solid line).
+Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(1)$ and~\eqref{eq:summomentproberror} (solid line).
 The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.
@@
 		\includegraphics{plots/peplots-1.mps}
-		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$, $A_4$, and $A_5$ and $E_8$ and the Leech lattice $\Lambda_{24}$.}
+		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$ and $A_4$ and $E_8$ and the Leech lattice $\Lambda_{24}$.}
 		\label{fig:peplots}
@@
 Recursive formulae for the moments of the Voronoi cell of the lattice $A_n$ were found.
-These enable accurate prediction of the performance of codes constructed from $A_n$ when \emph{lattice decoding} is used.
+These enable accurate prediction of the performance of codes constructed from $A_n$.
 In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice and our results include codes constructed using these packings as a special case.
@@
 
-We need to evaluate integrals of the form
+In~(\ref{eq:recurseint}) we required to evaluate integrals of the form
 \[
@@
 The initial condition is $\calG(z,0,0) = \frac{z^{-1}}{1 - z^{-1}}$.
-By inverting this $z$-transform and using the resultant expressions in~\eqref{eq:theCmformula} we obtain formulae in $n$ for the moment $C_n(m)$.
+By inverting this $z$-transform and using the resultant expressions in~\eqref{eq:theCmformula} we obtain formulae in $n$ for the moment $\calM_n(m)$.
 This procedure was used to generate the formula described in Section~\ref{sec:main-result}.
+Mathematica 8.0 was used to perform these calculations.
 

### 2011-11-16 paper.tex
New plots and other additions.

@@
 We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3$ and $A_4$.
+For these plots the SNR is related to noise variance according to~\cite{Viterbo_diamond_cutting_1996}
+\[
+\text{SNR} = \frac{V^{2/n}}{4\sigma^2},
+\]
+where $V$ is the volume of the Voronoi cell and $n$ is the dimension of the lattice.
+Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(1)$ and~\eqref{eq:summomentproberror} (solid line).
+The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.
+At most 196 moments where needed for Figure~\ref{fig:peplots}.
+We also display the probability of error computed approximately by Monte-Carlo simulation (dots).
+%We have used the fast nearest point algorithm for $A_n$ described in~\cite{McKilliam2009CoxeterLattices} and~\cite[p.
+448]{SPLAG}.
+The simulations are iterated until 500 error events occur.
+
+The plot also displays approximations for the probability of error for the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.
+The approximations are made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
+The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.
+The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.
+By the union bound the probability of error satisfies
+\newcommand{\erfc}{\operatorname{erfc}}
+\newcommand{\erf}{\operatorname{erf}}
+\[
+P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
+\]
+where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
+%The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.
+
+For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  The union bound on the probability of error is
+\[
+P_E \leq 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
+\]
+% The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.
+This was computed numerically by Conway and Sloane~\cite[p.
+61]{SPLAG}.
+The relationship between SNR and $\sigma^2$ is $\text{SNR} = 4 \times 0.065771 \sigma^{-2}$.
+
+% We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3$ and $A_4$.
 We assume that codes are constructed from these lattices using a \emph{Voronoi constellation}~\cite{Conway1983VoronoiCodes} of size $2^n$.
@@
 For this reason we think that this approximation is valid.
-The SNR is then given by
-\[
-\text{SNR} = \frac{4\calM_n(1)}{n \sigma^2} = \frac{n+3}{3\sqrt{n+1}\sigma^2}
-\]
-where $\sigma^2$ is the noise variance.  %As we have assumed that 1 bit is transmitted per symbol the SNR at channel capacity is equal to one.
+% The SNR is then given by
+% \[
+% \text{SNR} = \frac{4\calM_n(1)}{n \sigma^2} = \frac{n+3}{3\sqrt{n+1}\sigma^2}
+% \]
+% where $\sigma^2$ is the noise variance.  %As we have assumed that 1 bit is transmitted per symbol the SNR at channel capacity is equal to one.
 
-%Given a lattice with inradius $\rho$ and Voronoi cell having volume $V$ we define the SNR as
-%\[
-%\frac{V_n}{\sigma^2(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
-%\]
-%where $V_n$ is the volume of the $n$ dimensional hypersphere and the logarithms are base $2$.
+% %Given a lattice with inradius $\rho$ and Voronoi cell having volume $V$ we define the SNR as
+% %\[
+% %\frac{V_n}{\sigma^2(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
+% %\]
+% %where $V_n$ is the volume of the $n$ dimensional hypersphere and the logarithms are base $2$.
 
-Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(1)$ and~\eqref{eq:summomentproberror} (solid line).
+% Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(1)$ and~\eqref{eq:summomentproberror} (solid line).
 The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.
@@
 448]{SPLAG}.
-The simulations are iterated until 500 error events occur.
+% The simulations are iterated until 500 error events occur.
 
-We have also plotted approximations for the probability of error for codes constructed from the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.
+% We have also plotted approximations for the probability of error for codes constructed from the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.
 The approximation are made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
@@
 By the union bound the probability of error satisfies
-\newcommand{\erfc}{\operatorname{erfc}}
-\newcommand{\erf}{\operatorname{erf}}
-\[
-P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
-\]
-where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
+% \newcommand{\erfc}{\operatorname{erfc}}
+% \newcommand{\erf}{\operatorname{erf}}
+% \[
+% P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
+% \]
+% where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
 The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.
 
-For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  The union bound on the probability of error is
-\[
-P_E \leq 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
-\]
-The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.
+% For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  The union bound on the probability of error is
+% \[
+% P_E \leq 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
+% \]
+% The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.
 This was computed numerically by Conway and Sloane~\cite[p.

### 2011-11-16 plots/znsim.scala

@@ -8,7 +8,7 @@ import pubsim.Util
 import pubsim.lattices.util.ProbabilityOfCodingError
 import pubsim.distributions.GaussianNoise
 
-/*
+
 {
 println
 println("Running Monte-Carlo trials with the Zn lattice")
@@ -27,7 +27,8 @@ println("SNRdB \t Prob Error")
 while(pe > tope){
   //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
   val snr = scala.math.pow(10.0,dbnow/10.0)
-  val v = 2.0/3.0/snr
+   val v = 1/snr/4
+  //val v = 2.0/3.0/snr
   pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),1000).probError
   println(dbnow + "\t" + pe)
   efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
@@ -35,7 +36,6 @@ while(pe > tope){
 }
 efile.close
 }
-*/
 
 
 {
@@ -57,7 +57,8 @@ println("SNRdB \t Prob Error")
 while(pe > tope){
   //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
   val r2sdev = new BigRational(scala.math.sqrt(varnow*8.0) ,25)
-  val snr = 2.0/3.0/varnow
+  //val snr = 2.0/3.0/varnow
+  val snr = 1/varnow/4
   val snrdb = 10*scala.math.log10(snr)
   pe = 1.0 - pubsim.Util.erf(r2sdev.reciprocal, tol)
   varnow = varnow*varstep

### 2011-11-16 plots/e8simsphere.scala
Added scripts.

@@ -1,0 +1,98 @@
+/**
+* Probability of error plots for the lattice Zn.
+*/
+import java.math.BigInteger
+import pubsim.BigRational
+import pubsim.lattices.E8
+import pubsim.Util
+import pubsim.lattices.util.ProbabilityOfCodingError
+import pubsim.distributions.GaussianNoise
+
+
+{
+println
+println("Running Monte-Carlo trials with the E8 lattice")
+
+val lattice = new E8
+val startdb = 5.5
+val stepdb = 0.5
+val tope = 1e-6
+
+val sm = lattice.secondMoment
+
+var pe = 1.0
+var dbnow = startdb 
+val efile = new java.io.FileWriter("pemontecarlo_E8")
+
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val snr = scala.math.pow(10.0,dbnow/10.0)
+  val v = 1/snr/4
+  pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),500).probError
+  println(dbnow + "\t" + pe)
+  efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+  dbnow = dbnow + stepdb
+}
+efile.close
+}
+
+/*
+{
+val varstart = 1.0/6.0
+val varstep = 9.5/10.0
+val tope = 1e-12
+
+var pe = 1.0
+var varnow = varstart 
+val mfile = new java.io.FileWriter("E8UnionBound")
+
+val tol = new BigRational(BigInteger.ONE, BigInteger.TEN.pow(25));
+
+println
+println("Caculating union lower bound on the error probabilty of E8")
+println
+println("SNRdB \t Prob Error")
+while(pe > tope){
+  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+  val r2sdev = new BigRational(scala.math.sqrt(varnow*4.0) ,20)
+  val snr = 1/varnow/4
+  val snrdb = 10*scala.math.log10(snr)
+  pe = 240*(1 - pubsim.Util.erf(r2sdev.reciprocal, tol))
+  varnow = varnow*varstep
+  println(snrdb + "\t" + pe)
[…]

### 2011-11-16 paper.tex
Added minor comments from Emanuele.
Fixed a couple of minor things: (1) Logical ordering of description in first paragraph of Section II. (2) argmax display in Eqn.(1)

@@
  where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
-Since the generator matrix has rank $n$, the lattice $\Lambda$ is termed $n$-dimensional.
+In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
+A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional.
 If the generator is square, i.e.
@@
 $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
-For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the sub-space spanned by the columns of the generator matrix.
-The set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
+For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix.
+It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional sub-space containing the lattice $\Lambda$.
 
@@
 \begin{equation}\label{eq:mldecoder}
-\hat{\xbf} = \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
+\hat{\xbf} = \underset{\xbf \in \calX}{\operatorname{argmin}}\;\| \ybf - \xbf \|^2.
+%\hat{\xbf} &=& \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
 \end{equation}

### 2011-11-16 paper.tex
Couple of fixes: (1) First paragraph of Section II - Re-did logical ordering of sentences for a clearer flow. (2) argmax notation for Eqn.(1) fixed.

@@
 For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix.
-It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional sub-space containing the lattice $\Lambda$.
+It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.
 
@@
 where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
+For the $E_8$ lattice this approximation is an upper bound because the relevant vectors of $E_8$ (those vectors that define the Voronoi cell) are precisely the 240 minimal vectors.
 %The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.
@@
 \[
-P_E \leq 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
+P_E \approx 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
 \]

### 2011-11-18 plots/e8simsphere.scala
minor additions.

@@ -31,7 +31,7 @@ while(pe > tope){
   val snr = scala.math.pow(10.0,dbnow/10.0)
   val v = 1/snr/4
   pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),500).probError
-  println(dbnow + "\t" + pe)
+  println(dbnow + "\t" + pe + "\t"  + v)
   efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
   dbnow = dbnow + stepdb
 }

### 2011-11-21 paper.tex
Added some plot data etc

--- 2011-11-16 paper.tex
@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal]{IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 %\documentclass[correspondence,9pt]{IEEEtran}
@@
 %opening
-\title{On the error performance of the $A_n$ lattices}
-\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson}
+\title{Coding with the Lattice $A_n$}
+\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo}
 
@@
 \begin{abstract}
-We consider the root lattice $A_n$ and derive explicit formulae for
-the moments of its Voronoi cell.  We then show that these formulae
-enable accurate prediction of the error probability of
-lattice codes constructed from $A_n$.
+Lattices can be used to construct block codes.
+The special family of root lattices has attracted particular attention for this purpose.
+We consider the root lattice $A_n$ and derive formulae for the moments of its Voronoi cell.
+These formulae enable accurate prediction of the performance of codes constructed from $A_n$.
+In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
+These are the densest known sphere packings in dimensions two and three.
+Our results automatically include codes constructed using these packings.
 \end{abstract}
@@
 
-The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
-The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
+We consider block codes for the Gaussian channel constructed from a subset of points from a Euclidean lattice using \emph{lattice decoding} (as opposed to maximum likelihood decoding)~\cite{Erex2004_lattice_decoding}.
+It is known that such codes can reach the capacity of the Gaussian channel~\cite{Buda1989_some_opt_codes_structure,Erez2005,Erex2004_lattice_decoding}.
+These capacity results make use of randomly generated lattices that do not have any obvious exploitable structure for efficient decoding and encoding~\cite{Jalden2005_sphere_decoding_complexity}.
+For this reason (and many others) the study of specific lattices with known structure is important.
+The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention for this purpose~\cite{SPLAG}.
 In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
 This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.
-Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices,~\cite{Conway1982VoronoiRegions}.
-In this paper we extend their technique to compute the moments of any order for $A_n$.
+Conway and Sloane suggested this type of procedure would be possible in~\cite{Conway1982VoronoiRegions}.
+This paper can be seen as a continuation of that work focusing particularly on the lattice $A_n$.
 
@@
 These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
-In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but does offer a coding gain in small dimensions.
-For these cases, we provide an error probability expression that can be computed to any degree of accuracy at any finite signal-to-noise ratio.
-%We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
+Unfortunately, but not surprisingly, the lattice $A_n$ does not produce good codes in large dimensions.
+We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.
 In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.
@@
 In  Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
-We also plot the results of Monte-Carlo simulations that support our analytical results.
+We also plot the results of Monte-Carlo simulations.
+These simulations corroborate with our analytical results.
 %An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.
@@
 
-A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
+A \term{lattice}\index{lattice}, $\Lambda$, is a countable subset of points in $\reals^m$ such that
 \[
@@
  \]
- where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+ for a certain rank-$n$ $m \times n$ matrix $\Bbf$ of real elements, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
 Since the generator matrix has rank $n$, the lattice $\Lambda$ is termed $n$-dimensional.
@@
 $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
-For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the sub-space spanned by the columns of the generator matrix.
+For any lattice $\Lambda$ with an $m \times n$ generator matrix, we also define $\mathcal{S}_{\Lambda}$ to be the sub-space spanned by the columns of the generator matrix.
 The set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
@@
 
-The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, minimal vectors, center density, thickness, and the normalized second moment (or quantizing constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
+The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, center density, thickness, and the normalized second moment (or quantising constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
 The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.
@@
 %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder.
-A good overview of these techniques is given by Agrell~et.~al.~\cite{Agrell2002}.
-Fast nearest point algorithms are known for specific lattices~\cite{Conway1982FastQuantDec, McKilliam2008,McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
+A good overview of these techniques is given by~\cite{Agrell2002}.
+Fast nearest point algorithms are known for specific lattices~\citep{Conway1982FastQuantDec, McKilliam2008, McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
 For example, the root lattices $D_n$ and $A_n$ and their dual lattices $D_n^*$ and $A_n^*$ can be decoded in linear-time, i.e.
@@
 \[
-P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ),
+\sum_{\xbf \in \calX} \frac{\prob( \hat{\xbf} = \xbf )}{\abs{\calX}}.
 \]
-and the probability of error is
-\[
-P_E = 1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob(
-\hat{\xbf} = \xbf ).
-\]
+The probability of error is $1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf )$.
 %The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.
@@
 \begin{align}
-P_C &\leq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
- \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
-\vor(\Lambda) + \xbf )  \nonumber \\
+P_C &= \frac{1}{\abs{\calX}} \sum_{\cbf\in \calX} \prob( \xbf + \wbf \in \vor(\Lambda) + \xbf )  \\
 &= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
-&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
-e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
+&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)} e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
 \end{align}
-The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).
-By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
+The probability of error is $P_E = 1 - P_C$.  By expanding the exponential in terms of its McLaurin's series, we obtain
+%Removed $e^x = 1  + x + \frac{x^2}{2} + \dots$ and simply say "McLaurin's series"
 \begin{align}
-P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
-- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
-4\sigma^42!} - \dots d\xbf \nonumber \\
-&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
-\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
-d\xbf.  \label{eq:summomentproberror}
+P_C &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{ 4\sigma^42!} - \dots d\xbf \nonumber \\
+&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty \frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf.  \label{eq:summomentproberror}
 \end{align}
-So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
-The number of terms required will increase as the noise variance gets smaller.
-This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
+Hence it is enough to know the values of $\int_{\vor(\Lambda)} \left(\|\xbf\|^2\right)^m d\xbf$ for $m=1,2\dots,M$ for some sufficiently large $M$, in order to obtain arbitrarily accurate approximations to the probability of error.
 
-\newcommand{\calM}{M}
-In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
+In this paper we focus on lattice codes constructed from the family of lattices called $A_n$. We will discover expressions for
 \[
- \calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
+ C_n(m) = \int_{\vor(A_n)} \left(\|\xbf\|^2\right)^m d\xbf.
 \]
-These can be summed in (\ref{eq:summomentproberror}) to give arbitrarily accurate approximations for the probability of error.
+These can be summed to give arbitrarily accurate approximations for the probability of error.  In this paper we call $C_n(m)$ the $m^{\text{th}}$ \emph{moment} of the lattice $A_n$.
 
@@
 
-We now state our main result.  The $m$th moment $\calM_n(m)$ of the lattice $A_n$ satisfies
+We now state our main result.  The $m$th moment $C_n(m)$ satisfies
 \begin{equation}\label{eq:theCmformula}
-\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)} 
+\frac{C_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)}
 \end{equation}
@@
 \]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $C_m(n)$ in terms of $n$.  The first four such formula are:
 \begin{align*}
-\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
-\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
-\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
-\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
-\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
+C_0(n) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
+C_1(n) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
+C_2(n) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
+C_3(n) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
+%C_4(n) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
 \end{align*}
-We have explicitly tabulated these formula for $m=0$ to $40$.
-For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.
-We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.
+We have tabulated these formula for $m=0$ to $40$.  We derive these results in Section~\ref{sec:powers-eucl-norm}.  First we need some properties of the lattice $A_n$.
 
@@
 \label{eq:An_sub_Zn}
-  A_{n} = \ints^{n+1} \cap H = \big\{ \xbf \in \ints^{n+1} \mid \dotprod{\xbf}{\onebf} = 0  \big\}.
+  A_{n} = \ints^{n+1} \cap H = \big\{ \xbf \in \ints^{n+1} \mid \xbf\cdot\onebf = 0  \big\}.
 \end{equation}
@@
 % \]
-The lattice has $n(n+1)$ minimal vectors, each of squared Euclidean length $2$, so the packing radius is $\frac{1}{\sqrt{2}}$.
-The $n$-volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$~\cite[p.
+The lattice has determinant $n+1$ so the volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$.
+The lattice has $n(n+1)$ minimal vectors, each of norm $2$, so the packing radius is $\frac{1}{\sqrt{2}}$~\cite[p.
 108]{SPLAG}.
@@
 \[
-\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{\onebf^\prime \onebf} = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
+\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
 \]
 the projection matrix orthogonal to $\onebf$ (i.e.
-into the zero-sum plane) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.
+into $H$) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.
 Given a set $S$ of vectors from $\reals^{n+1}$ we write $\Qbf S$  to denote the set with elements $\Qbf \sbf$ for all $\sbf \in S$, i.e.
@@
 \begin{lemma} \label{lem:QVorZnsubsetVorAn}
-The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is a subset of $\vor(A_{n})$.  That is,
+The projection of $\vor(\ints^{n+1})$ into H is a subset of $\vor(A_{n}) \cap H$.  That is,
 \[
@@
 By definition \eqref{eq:An_sub_Zn} $\xbf'\onebf = 0$ so $\|\xbf - \ybf\|^2 < \|\ybf\|^2$.
-This violates that $\ybf \in \vor(\ints^{n+1})$ and hence $\Qbf\ybf \in \vor(A_n)$.\footnote{This proof can be generalised to show that for any lattice $L$ and hyperplane $P$ such that $P\cap L$ is also a lattice it is true that $p\vor(L) \subseteq \vor(L \cap P)$ where $p$ indicates the orthogonal projection into $P$~\cite[Lemma~2.1]{McKilliam2010thesis}.}
+This violates that $\ybf \in \vor(\ints^{n+1})$ and hence $\Qbf\ybf \in \vor(A_n)$.\footnote{This proof can be generalised to show that for any lattice $L$ and hyperplane $P$ such that $P\cap L$ is also a lattice it is true that $p\vor(L) \subseteq \vor(L \cap P) \cap P$ where $p$ indicates the orthogonal projection into $P$~\cite[Lemma~2.1]{McKilliam2010thesis}.}
 \end{IEEEproof}
@@
 \begin{theorem}  \label{thm:VorAn=QVorZn1}
-The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is equal to $\vor(A_{n})$. That is,
+The projection of $\vor(\ints^{n+1})$ into H is equal to $H \cap \vor(A_{n})$. That is,
 \[
-\vor(A_n) = \Qbf\vor(\ints^{n+1}).
+H \cap \vor(A_n) = \Qbf\vor(\ints^{n+1}).
 \]
@@
 \begin{proof}
-Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.
-The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.
-From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
+The $n$-volume of $H \cap \vor(A_n)$ is given by the square root of the determinant of $A_n$, that is
 \[
-\sum_{i = 1}^{n+1} \frac{\dotprod{\onebf}{\ebf_i}}{\|\onebf\|} =  \sum_{i = 1}^{n+1} \frac{1}{\sqrt{n+1}} = \sqrt{n + 1}
+\sqrt{\det{A_n}} = \sqrt{n + 1}.
 \]
-also.
-It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
+From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to $\sqrt{n + 1}$ also.
+It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq H \cap \vor(A_n)$, so, because the volumes are the same, and because $H \cap \vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $H \cap \vor(A_n) = \Qbf\vor(\ints^{n+1})$.
 \end{proof}
@@
 
-We would like to be able to integrate functions over the Voronoi cell of $A_n$.
-Consider a function $f : \reals^{n+1} \mapsto \reals$.
-The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.
-The Voronoi cell is a subset of the zero-sum plane that has zero $n+1$-dimensional volume.
-So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.
-This is not what we intend.
-By an appropriate change of variables it would be possible to write the Voronoi cell $\vor(A_n)$ in an $n$-dimensional coordinate system, and then integrate.
-However, we find the following approach simpler.
-Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $x + y$ where $x \in S$ and $y = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
-Now, the integral over the Voronoi cell can be written as
+Consider a function $f : \reals^{n+1} \mapsto \reals$.  We would like to be able to solve integrals of the form
 \begin{equation}\label{eq:vorfint}
-\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
+\int_{\vor(A_n)} f(\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1})} f(\xbf) d\xbf.
 \end{equation}
-It is not immediately clear how an integral over $\vor(A_n) \times \onebf$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
-\begin{equation}\label{eq:unormalised}
+It is not immediately clear how an integral over $\vor(A_n)$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
+\[
 \int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
-\end{equation}
+\]
 This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.
 To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.
-Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.
+Then $\Qbf\xbf = \Qbf(\xbf + k \onebf)$ for all $k \in \reals$ such that $\xbf +  k \onebf \in \vor(\ints^{n+1})$, i.e.
+for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.
 The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by
@@
 \]
-The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
+The integral~\eqref{eq:vorfint} can now be written as
 \[
-\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
+\int_{\vor(A_n)} f(\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
 \]
-The primary advantage of this integral is that the bounds are given by the hypercube $\vor(\ints^{n+1})$.
+The primary advantage of this integral is that the bounds of are given by the hypercube $\vor(\ints^{n+1})$.
 
@@
 \end{equation}
-Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector
-\[
-\wbf = [t, 0, w_1, w_2, \dots, w_{n-1}]^\prime.
-\]
+Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector $[t, 0, w_1, w_2, \dots, w_{n-1}]^\prime$.
 Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.
@@
 
-We now derive expressions for the $\calM_n(m)$.  Setting $f(\Qbf\xbf) = \left(\|\Qbf\wbf\|^{2}\right)^m$ in~\eqref{eq:intwewilluse} we obtain,
+We now derive expressions for the $C_n(m)$.  Setting $f(\Qbf\xbf) = \left(\|\Qbf\wbf\|^{2}\right)^m$ in~\eqref{eq:intwewilluse} we obtain,
 \[
-\frac{\calM_n(m)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
+\frac{C_m(n)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
 \]
@@
 &= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n+1}w_i + \left(\sum_{i=1}^{n+1}w_i \right)^2 \right)\\
-&= C + D,
+&= c + d,
 \end{align*}
@@
 \[
-C = \left(\frac{n}{n+1} \right) t^2 \qquad \text{and} \qquad D = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
+c = \left(\frac{n}{n+1} \right) t^2,\qquad \text{and} \qquad d = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
 \]
-and where,
+and where $A = \sum_{i=1}^{n-1}w_i^2$ and $B = \sum_{i=1}^{n-1}w_i$.  Now,
 \[
- A = \sum_{i=1}^{n-1}w_i^2 \qquad \text{and} \qquad B = \sum_{i=1}^{n-1}w_i.
+ \frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(c+d)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
 \]
-Now,
+and by expanding the binomial $(c+d)^m$ we get
 \[
- \frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(C+D)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
+\frac{C_m(n)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} c^{m-k} \int_0^t \cdots \int_0^t \, d^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
 \]
-and by expanding the binomial $(C+D)^m$ we get
-\[
-\frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} C^{m-k} \int_0^t \cdots \int_0^t \, D^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
-\]
-Expanding $D^k$ as a trinomial gives
+Expanding $d^k$ as a trinomial gives
 \begin{align*}
-D^k &= \sum_{k_1+k_2+k_3=k} \frac{k! A^{k_1} B^{2k_3 + k_2}}{k_1! k_2! k_3!} \left(\frac{-1}{n+1}\right)^{k_2+k_3}2^{k_2}t^{k_2}  \\
+d^k &= \sum_{k_1+k_2+k_3=k} \frac{k! A^{k_1} B^{2k_3 + k_2}}{k_1! k_2! k_3!} \left(\frac{-1}{n+1}\right)^{k_2+k_3}2^{k_2}t^{k_2}  \\
 &=  \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{k!  A^{a} B^{2k - 2a - b}}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}2^{b}t^{b}
@@
 In Appendix~\ref{sec:mult-type-integr} we show that the integral of $A^{a} B^{2k - 2a - b}$ over $w_1,\dots w_{n-1}$ is
-\begin{equation} \label{eq:recurseint}
+\[
 \int_0^t \cdots \int_0^t A^{a} B^{2k-2a-b} dw_{n-1} \, \dots \, dw_1 = t^{n-1+2k - b} G(n-1,a,2k - 2a - b).
-\end{equation}
-where $G(n,c,d)$ satisfies the recursion given by~\eqref{eq:theGrecursion}.  So, let $P$ satisfy
+\]
+where $G(n,k,m)$ satisfies the recursion given by~\eqref{eq:theGrecursion}.  So, let $D$ satisfy
 \begin{align*}
-P &= t^{1 - n - 2k}\int_0^t \cdots \int_0^t D^k dw_{n-1} \, \dots \, dw_1 \\
+D &= t^{1 - n - 2k}\int_0^t \cdots \int_0^t d^k dw_{n-1} \, \dots \, dw_1 \\
 &= \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{2^{b} k!G(n-1,a,2k - 2a - b)}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}.
 \end{align*}
-Now $C^{m-k} = \left(\frac{n}{n+1} \right)^{m-k} t^{2(m-k)}$ and
+Now $c^{m-k} = \left(\frac{n}{n+1} \right)^{m-k} t^{2(m-k)}$ and
 \begin{align*}
-\frac{\calM_n(m)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy \\
-&= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \beta(y,n+2m,0) dy \\
-&= \frac{1}{n+2m} \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P.
+\frac{C_m(n)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} D \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy. \\
+&= \frac{1}{n+2m} \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} D.
 \end{align*}
@@
 We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3$ and $A_4$.
-For these plots the SNR is related to noise variance according to~\cite{Viterbo_diamond_cutting_1996}
+We assume that codes are constructed from these lattices using a \emph{Voronoi constellation}~\cite{Conway1983VoronoiCodes} of size $2^n$.
+This corresponds to a rate of $1$ bit per dimension (or symbol).
+The shaping region of these constellations is $2\vor(A_n)$, i.e.
+the Voronoi cell of $A_n$ scaled by a factor of $2$.
+The average power of these constellations can be approximated by four times the second moment $4C_1(m)$.
+This approximation is conservative as Voronoi constellations can have average power smaller than the second moment of the shaping region~\cite{Conway1983VoronoiCodes}.
+The approximation becomes more accurate as the code rate and correspondingly the volume of the shaping region grows.
+%In proving that lattice codes using lattice decoding can achieve capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a common dither between transmitter and receiver.
+This is to ensure that the average power is \emph{exactly} the second moment of the shaping region.
+For this reason we think that this approximation is valid.
+The SNR is then given by
 \[
-\text{SNR} = \frac{V^{2/n}}{4\sigma^2},
+\text{SNR} = \frac{4C_1(n)}{n \sigma^2} = \frac{n+3}{3\sqrt{n+1}\sigma^2}
 \]
-where $V$ is the volume of the Voronoi cell and $n$ is the dimension of the lattice.
-Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(1)$ and~\eqref{eq:summomentproberror} (solid line).
+where $\sigma^2$ is the noise variance.  %As we have assumed that 1 bit is transmitted per symbol the SNR at channel capacity is equal to one.
+
+Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $C_n(m)$ and~\eqref{eq:summomentproberror} (solid line).
 The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.
@@
 
-The plot also displays approximations for the probability of error for the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.
-The approximations are made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
+We have also plotted approximations for the probability of error for codes constructed from the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.
+The approximation are made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
 The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.
@@
 where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
-%The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.
+The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.
 
@@
 \]
-% The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.
-This was computed numerically by Conway and Sloane~\cite[p.
-61]{SPLAG}.
-The relationship between SNR and $\sigma^2$ is $\text{SNR} = 4 \times 0.065771 \sigma^{-2}$.
-
-% We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3$ and $A_4$.
-We assume that codes are constructed from these lattices using a \emph{Voronoi constellation}~\cite{Conway1983VoronoiCodes} of size $2^n$.
-This corresponds to a rate of $1$ bit per dimension (or symbol).
-The shaping region of these constellations is $2\vor(A_n)$, i.e.
-the Voronoi cell of $A_n$ scaled by a factor of $2$.
-The average power of these constellations can be approximated by four times the second moment $4\calM_n(1)$.
-This approximation is conservative as Voronoi constellations can have average power smaller than the second moment of the shaping region~\cite{Conway1983VoronoiCodes}.
-The approximation becomes more accurate as the code rate and correspondingly the volume of the shaping region grows.
-%In proving that lattice codes using lattice decoding can achieve capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a common dither between transmitter and receiver.
-This is to ensure that the average power is \emph{exactly} the second moment of the shaping region.
-For this reason we think that this approximation is valid.
-% The SNR is then given by
-% \[
-% \text{SNR} = \frac{4\calM_n(1)}{n \sigma^2} = \frac{n+3}{3\sqrt{n+1}\sigma^2}
-% \]
-% where $\sigma^2$ is the noise variance.  %As we have assumed that 1 bit is transmitted per symbol the SNR at channel capacity is equal to one.
-
-% %Given a lattice with inradius $\rho$ and Voronoi cell having volume $V$ we define the SNR as
-% %\[
-% %\frac{V_n}{\sigma^2(n+2)(\log(V_n) - \log(V) + n\log(\rho))}
-% %\]
-% %where $V_n$ is the volume of the $n$ dimensional hypersphere and the logarithms are base $2$.
-
-% Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(1)$ and~\eqref{eq:summomentproberror} (solid line).
-The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.
-At most 196 moments where needed for Figure~\ref{fig:peplots}.
-We also display the probability of error computed approximately by Monte-Carlo simulation (dots).
-%We have used the fast nearest point algorithm for $A_n$ described in~\cite{McKilliam2009CoxeterLattices} and~\cite[p.
-448]{SPLAG}.
-% The simulations are iterated until 500 error events occur.
-
-% We have also plotted approximations for the probability of error for codes constructed from the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.
-The approximation are made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
-The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.
-The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.
-By the union bound the probability of error satisfies
-% \newcommand{\erfc}{\operatorname{erfc}}
-% \newcommand{\erf}{\operatorname{erf}}
-% \[
-% P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
-% \]
-% where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
-The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.
-
-% For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  The union bound on the probability of error is
-% \[
-% P_E \leq 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
-% \]
-% The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.
+The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.
 This was computed numerically by Conway and Sloane~\cite[p.
@@
 		\includegraphics{plots/peplots-1.mps}
-		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$ and $A_4$ and $E_8$ and the Leech lattice $\Lambda_{24}$.}
+		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$, $A_4$, $E_8$ and the Leech lattice $\Lambda_{24}$.}
 		\label{fig:peplots}
@@
 Recursive formulae for the moments of the Voronoi cell of the lattice $A_n$ were found.
-These enable accurate prediction of the performance of codes constructed from $A_n$.
+These enable accurate prediction of the performance of codes constructed from $A_n$ when \emph{lattice decoding} is used.
 In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice and our results include codes constructed using these packings as a special case.
-%The lattice $A_n$ does not produce good codes in large dimension, but we are hopeful that the techniques developed here can be applied to other lattices.
+The lattice $A_n$ does not produce good codes in large dimension, but we are hopeful that the techniques developed here can be applied to other lattices.
 In particular similar techniques might apply to the family Coxeter lattices, which contains the root lattice $E_8$.
@@
 
-In~(\ref{eq:recurseint}) we required to evaluate integrals of the form
+We need to evaluate integrals of the form
 \[
-F(n-1,a,2k - 2a - b) = \int^{t}_{0} \cdots \int^{t}_{0} A^a B^{2k - 2a - b} \,dw_{n-1}\cdots dw_{1},
+F(n,c,d) = \int^{t}_{0} \cdots \int^{t}_{0} \left(\sum_{j=1}^{n} x_j^2\right)^c \left( \sum_{i=1}^{n} x_i \right)^d \,dx_1\cdots dx_{n}.
  \]
-or equivalently, integrals of the form
-\[
-F(n,c,d) = \int^{t}_{0} \cdots \int^{t}_{0} \left(\sum_{j=1}^{n} x_j^2\right)^c \left( \sum_{i=1}^{n} x_i \right)^d \,dx_1\cdots dx_{n}
-\]
-where $n,c$ and $d$ are integers.  We shall find a recursion describing this integral.  Write
+We shall find a recursion describing this integral.  Write
 \[
@@
 The initial condition is $\calG(z,0,0) = \frac{z^{-1}}{1 - z^{-1}}$.
-By inverting this $z$-transform and using the resultant expressions in~\eqref{eq:theCmformula} we obtain formulae in $n$ for the moment $\calM_n(m)$.
+By inverting this $z$-transform and using the resultant expressions in~\eqref{eq:theCmformula} we obtain formulae in $n$ for the moment $C_n(m)$.
 This procedure was used to generate the formula described in Section~\ref{sec:main-result}.
-Mathematica 8.0 was used to perform these calculations.
 

### 2011-11-21 plots/proberrorsim.scala

@@ -6,11 +6,11 @@ import pubsim.lattices.An.AnFastSelect
 import pubsim.lattices.An.MomentComputer
 import pubsim.BigRational
 
-val ns = List(2,3,4,5)
+val ns = List(2,3,4,5,6,7,8)
 //val lattice = new AnFastSelect(n)
 val varstart = new BigRational(1,2)
-val varstep = new BigRational(9,10)
-val tope = 1e-12
+val varstep = new BigRational(95,100)
+val tope = 5e-12
 val petol = 1e-17
 
 print("Reading existing moment data, this might take a moment....")

### 2011-11-21 paper.tex
minor edits to proberrorsim.

--- 2011-11-16 paper.tex
@@
 However, we find the following approach simpler.
-Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $x + y$ where $x \in S$ and $y = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
+Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
 Now, the integral over the Voronoi cell can be written as
@@
 The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.
-By the union bound the probability of error satisfies
+Applying the union bound on the  probability of error satisfies
 \newcommand{\erfc}{\operatorname{erfc}}
@@
 
-For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  The union bound on the probability of error is
+For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  An approximation for the probability of error is
 \[

### 2011-11-21 plots/proberrorsim.scala

@@ -6,12 +6,12 @@ import pubsim.lattices.An.AnFastSelect
 import pubsim.lattices.An.MomentComputer
 import pubsim.BigRational
 
-val ns = List(2,3,4,5,6,7,8)
+val ns = List(2,3,4,5,6)
 //val lattice = new AnFastSelect(n)
 val varstart = new BigRational(1,2)
-val varstep = new BigRational(95,100)
-val tope = 5e-12
-val petol = 1e-17
+val varstep = new BigRational(93,100)
+val tope = 5e-11
+val petol = 1e-16
 
 print("Reading existing moment data, this might take a moment....")
 val mc = new MomentComputer("momentstore")

### 2011-11-21 plots/proberrorsimsphere.scala
Added probsimsphere that was missing from the repository.

@@ -1,0 +1,45 @@
+/**
+* Probability of error plots for the lattice An.
+*/
+import pubsim.lattices.util.PowerOfEuclideanNorm
+import pubsim.lattices.An.AnFastSelect
+import pubsim.lattices.An.MomentComputer
+import pubsim.BigRational
+import pubsim.Util
+
+val ns = List(3,4,5)
+//val lattice = new AnFastSelect(n)
+val varstart = new BigRational(1,2)
+val varstep = new BigRational(9,10)
+val tope = 1e-12
+val petol = 1e-17
+
+print("Reading existing moment data, this might take a moment....")
+val mc = new MomentComputer("momentstore")
+println("finished")
+
+for(n <- ns){
+  var pe = 1.0
+  var varnow = varstart 
+  val mfile = new java.io.FileWriter("moment" + n)
+
+  val V = scala.math.sqrt(n+1.0)
+
+  println
+  println("Now running n = " + n)
+  println("SNRdB \t Prob Error")
+  while(pe > tope){
+    //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
+    val snr = scala.math.pow(V,2.0/n)/varnow.doubleValue/4
+    val snrdb = 10*scala.math.log10(snr)
+    pe = mc.ProbError(n,varnow,petol)
+    varnow = varnow.multiply(varstep)
+    println(snrdb + "\t" + pe + "\t" + mc.numberMomentsLastUsed)
+    mfile.write(snrdb.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
+  }
+
+  mfile.close
+
+}
+
+mc.save

### 2011-11-26 plots/proberrorsimsphere.scala
Added new data for plots.

@@ -7,10 +7,10 @@ import pubsim.lattices.An.MomentComputer
 import pubsim.BigRational
 import pubsim.Util
 
-val ns = List(3,4,5)
+val ns = List(2,3,4,5)
 //val lattice = new AnFastSelect(n)
 val varstart = new BigRational(1,2)
-val varstep = new BigRational(9,10)
+val varstep = new BigRational(93,100)
 val tope = 1e-12
 val petol = 1e-17
 

### 2011-11-26 paper.tex
minor edits.  Removed Leech lattice from plot.

@@
 A \emph{lattice code} is a signal space code with codewords taken from a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
-There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$ and
+There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$.
+The codewords are given by those lattice points inside the shaping region, that is,
 \[
@@
 \]
-Possible choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
+Common choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes, or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
 A consequence of the finiteness of a lattice code is that the maximum likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to lattice decoding.
@@
 \begin{align*}
-C_0 &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
+M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
  &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy. \\
@@
 
-We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3$ and $A_4$.
+We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3, A_4$ and $A_5$.
 For these plots the SNR is related to noise variance according to~\cite{Viterbo_diamond_cutting_1996}
@@
 where $V$ is the volume of the Voronoi cell and $n$ is the dimension of the lattice.
-Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(1)$ and~\eqref{eq:summomentproberror} (solid line).
+Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(m)$ and~\eqref{eq:summomentproberror} (solid line).
 The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.
-At most 196 moments where needed for Figure~\ref{fig:peplots}.
+At most 252 moments where needed for Figure~\ref{fig:peplots}.
 We also display the probability of error computed approximately by Monte-Carlo simulation (dots).
@@
 
-The plot also displays approximations for the probability of error for the 8-dimensional $E_8$ lattice and the 24-dimensional Leech lattice $\Lambda_{24}$.
-The approximations are made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
+The plot also display an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
+The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
 The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.
 The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.
-Applying the union bound on the  probability of error satisfies
+Applying the union bound the probability of error satisfies
 \newcommand{\erfc}{\operatorname{erfc}}
@@
 
-For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  An approximation for the probability of error is
-\[
-P_E \approx 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
-\]
+%For $\Lambda_{24}$ there are $196560$ minimal vectors each of length $2$.  The packing radius is $\rho = 1$.  An approximation for the probability of error is
+%\[
+%P_E \approx 196560\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 196560\erfc\left(\frac{1}{\sqrt{2}\sigma}\right).
+%\]
 % The second moment of this representation of $\Lambda_{24}$ is approximately $24 \times 0.065771$.
@@
 		\includegraphics{plots/peplots-1.mps}
-		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3$ and $A_4$ and $E_8$ and the Leech lattice $\Lambda_{24}$.}
+		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5$ and $E_8$.}
 		\label{fig:peplots}
@@
 These enable accurate prediction of the performance of codes constructed from $A_n$.
-In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice and our results include codes constructed using these packings as a special case.
+In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
+Our results include codes constructed using these packings as a special case.
 %The lattice $A_n$ does not produce good codes in large dimension, but we are hopeful that the techniques developed here can be applied to other lattices.

### 2011-11-26 plots/montecarlo.scala

@@ -5,7 +5,7 @@ import pubsim.lattices.util.ProbabilityOfCodingError
 import pubsim.lattices.An.AnFastSelect
 import pubsim.distributions.GaussianNoise
 
-val ns = List(2,3,4)
+val ns = List(1,2,3,4,5)
 
 val startdb = 5.5
 val stepdb = 0.5
@@ -27,7 +27,7 @@ for(n <- ns){
     val snr = scala.math.pow(10.0,dbnow/10.0)
     //val v = (3.0+n)/3.0/scala.math.sqrt(n+1)/snr
     val v  = scala.math.pow(V,2.0/n)/snr/4
-    pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),500).probError
+    pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),1000).probError
     println(dbnow + "\t" + pe)
     mfile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
     dbnow = dbnow + stepdb

### 2011-12-04 paper.tex
removed unnesecary stuff.  Added some instructions to README.

@@
 
-We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3, A_4$ and $A_5$.
+We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3, A_4, A_5$ and $A_8$.
 For these plots the SNR is related to noise variance according to~\cite{Viterbo_diamond_cutting_1996}
@@
 The number of moments needed to ensure a certain number of decimal places accuracy depends on $n$ and also on the noise variance $\sigma^2$.
-At most 252 moments where needed for Figure~\ref{fig:peplots}.
+At most 321 moments where needed for Figure~\ref{fig:peplots}.
 We also display the probability of error computed approximately by Monte-Carlo simulation (dots).
@@
 		\includegraphics{plots/peplots-1.mps}
-		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5$ and $E_8$.}
+		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5, A_8$ and $E_8$.}
 		\label{fig:peplots}

### 2011-12-04 plots/montecarlo.scala

@@ -5,7 +5,7 @@ import pubsim.lattices.util.ProbabilityOfCodingError
 import pubsim.lattices.An.AnFastSelect
 import pubsim.distributions.GaussianNoise
 
-val ns = List(1,2,3,4,5)
+val ns = List(8)
 
 val startdb = 5.5
 val stepdb = 0.5
@@ -27,7 +27,7 @@ for(n <- ns){
     val snr = scala.math.pow(10.0,dbnow/10.0)
     //val v = (3.0+n)/3.0/scala.math.sqrt(n+1)/snr
     val v  = scala.math.pow(V,2.0/n)/snr/4
-    pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),1000).probError
+    pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),500).probError
     println(dbnow + "\t" + pe)
     mfile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
     dbnow = dbnow + stepdb

### 2012-01-23 paper.tex
minor fixes to paper.

@@
 The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
-The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
+The highly symmetric structure of these lattices provides the grounds for extremely efficient encoding and decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
 In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
 This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.
-Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices,~\cite{Conway1982VoronoiRegions}.
+Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices~\cite{Conway1982VoronoiRegions}.
 In this paper we extend their technique to compute the moments of any order for $A_n$.
@@
 In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
-These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
+These are the densest sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
 In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but does offer a coding gain in small dimensions.
@@
 We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.
-Section~\ref{sec:main-result} states the main result (without proof), which gives recursive formula to compute the moments of the Voronoi cell of $A_n$.
+Section~\ref{sec:main-result} states the main result, describing recursive formula to compute the moments of the Voronoi cell of $A_n$.
 Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.
-An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n+1$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $(n+1)$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
-In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $n$-dimensional hypercube.
+In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube.
 These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.
-In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.
+In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$, $A_5$ and $A_8$.
 We also plot the results of Monte-Carlo simulations that support our analytical results.
@@
 For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix.
-It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.
+%It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.
 
 The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\mathcal{S}_{\Lambda}$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point.
-It can be shown that the Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
+The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
 It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.
-That is, the Voronoi cell tessellates space when translated by points in $\Lambda$.
+That is, the Voronoi cell tessellates when translated by points in $\Lambda$.
 To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed.
@@
 \begin{align}
-P_C &\leq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
+P_C &\geq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
  \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
@@
 \begin{equation}\label{eq:theCmformula}
-\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)}
+\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{H(n,m,k,a,b)},
 \end{equation}
@@
 \[
-H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{2^{b} n^{m-k}},
+H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{(-1)^{k-a}2^{b} n^{m-k}},
 \]
@@
 \begin{equation}\label{eq:theGrecursion}
-G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
+G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1},
 \end{equation}
@@
 \calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
-\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
+\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, \\
 \calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
@@
 
-The Voronoi cell of $A_n$ is closely related to the $n+1$ dimensional hypercube $\vor(\ints^{n+1})$ as the next theorem will show.
+The Voronoi cell of $A_n$ is closely related to the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$ as the next theorem will show.
 This result has appeared previously~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}, but we repeat it here so that this paper is self contained.
@@
 The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.
-The Voronoi cell is a subset of the zero-sum plane that has zero $n+1$-dimensional volume.
+The Voronoi cell is a subset of the zero-sum plane that has zero $(n+1)$-dimensional volume.
 So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.
@@
 \]
-The primary advantage of this integral is that the bounds are given by the hypercube $\vor(\ints^{n+1})$.
+The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.
 
@@
 
-We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables gives
+We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables,
 \[
@@
 Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.
-Interestingly $\wbf$ does not contain $y$ so the term inside the integral does not depend on $y$.
+Interestingly $\wbf$ does not contain $y$ so the term inside the integral~\eqref{eq:intwewilluse} does not depend on $y$.
 This is the integral we will use to compute the moments of $A_n$.
@@
 
-The plot also display an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
+The plot also displays an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
 The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
@@
 \]
-where $n,c$ and $d$ are integers.  We shall find a recursion describing this integral.  Write
+where $n,c$ and $d$ are integers.  We will find a recursion describing this integral.  Write
 \[
@@
 \[
-F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'}}{2c'+d'+1}F(n-1,c-c',d-d')
+F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1}F(n-1,c-c',d-d')
 \]

### 2012-01-23 plots/montecarlo.scala
minor edits to the montecarlo script.

@@ -5,7 +5,7 @@ import pubsim.lattices.util.ProbabilityOfCodingError
 import pubsim.lattices.An.AnFastSelect
 import pubsim.distributions.GaussianNoise
 
-val ns = List(8)
+val ns = List(2,3,4,5,8)
 
 val startdb = 5.5
 val stepdb = 0.5

### 2012-01-23 plots/montecarlo.sh

@@ -1,1 +1,2 @@
 scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar montecarlo.scala
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar znsim.scala

### 2012-01-23 build.sh
did some cleanup

@@ -2,7 +2,9 @@ cd plots
 mpost -interaction=nonstopmode peplots.mp
 cd ..
 
-pdflatex paper.tex
+latex paper.tex
 bibtex paper
-pdflatex paper.tex
-pdflatex paper.tex
+latex paper.tex
+latex paper.tex
+dvips paper.dvi
+ps2pdf paper.ps

### 2012-01-23 plots/e8sim.scala

@@ -8,7 +8,7 @@ import pubsim.Util
 import pubsim.lattices.util.ProbabilityOfCodingError
 import pubsim.distributions.GaussianNoise
 
-/*
+
 {
 println
 println("Running Monte-Carlo trials with the E8 lattice")
@@ -29,15 +29,14 @@ println("SNRdB \t Prob Error")
 while(pe > tope){
   //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
   val snr = scala.math.pow(10.0,dbnow/10.0)
-  val v = 4.0*sm/snr/8
-  pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),200).probError
-  println(dbnow + "\t" + pe)
+  val v = 1/snr/4
+  pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),500).probError
+  println(dbnow + "\t" + pe + "\t"  + v)
   efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
   dbnow = dbnow + stepdb
 }
 efile.close
 }
-*/
 
 
 {
@@ -58,7 +57,7 @@ println("SNRdB \t Prob Error")
 while(pe > tope){
   //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
   val r2sdev = new BigRational(scala.math.sqrt(varnow*4.0) ,20)
-  val snr = 929.0/3240.0/varnow
+  val snr = 1/varnow/4
   val snrdb = 10*scala.math.log10(snr)
   pe = 240*(1 - pubsim.Util.erf(r2sdev.reciprocal, tol))
   varnow = varnow*varstep
@@ -67,33 +66,3 @@ while(pe > tope){
 }
 mfile.close
 }
-
-
-{
-val varstart = 1.0/6.0
-val varstep = 9.5/10.0
-val tope = 1e-12
-
-var pe = 1.0
-var varnow = varstart 
-val mfile = new java.io.FileWriter("LeechUnionBound")
-
-val tol = new BigRational(BigInteger.ONE, BigInteger.TEN.pow(25));
-
-println
-println("Caculating union lower bound on the error probabilty of Leech lattice")
-println
-println("SNRdB \t Prob Error")
-while(pe > tope){
-  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
[…]

### 2012-01-23 plots/e8simsphere.scala

@@ -1,98 +1,0 @@
-/**
-* Probability of error plots for the lattice Zn.
-*/
-import java.math.BigInteger
-import pubsim.BigRational
-import pubsim.lattices.E8
-import pubsim.Util
-import pubsim.lattices.util.ProbabilityOfCodingError
-import pubsim.distributions.GaussianNoise
-
-
-{
-println
-println("Running Monte-Carlo trials with the E8 lattice")
-
-val lattice = new E8
-val startdb = 5.5
-val stepdb = 0.5
-val tope = 1e-6
-
-val sm = lattice.secondMoment
-
-var pe = 1.0
-var dbnow = startdb 
-val efile = new java.io.FileWriter("pemontecarlo_E8")
-
-println
-println("SNRdB \t Prob Error")
-while(pe > tope){
-  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
-  val snr = scala.math.pow(10.0,dbnow/10.0)
-  val v = 1/snr/4
-  pe = new ProbabilityOfCodingError(lattice,new GaussianNoise(0,v),500).probError
-  println(dbnow + "\t" + pe + "\t"  + v)
-  efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
-  dbnow = dbnow + stepdb
-}
-efile.close
-}
-
-/*
-{
-val varstart = 1.0/6.0
-val varstep = 9.5/10.0
-val tope = 1e-12
-
-var pe = 1.0
-var varnow = varstart 
-val mfile = new java.io.FileWriter("E8UnionBound")
-
-val tol = new BigRational(BigInteger.ONE, BigInteger.TEN.pow(25));
-
-println
-println("Caculating union lower bound on the error probabilty of E8")
-println
-println("SNRdB \t Prob Error")
-while(pe > tope){
-  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
-  val r2sdev = new BigRational(scala.math.sqrt(varnow*4.0) ,20)
-  val snr = 1/varnow/4
-  val snrdb = 10*scala.math.log10(snr)
-  pe = 240*(1 - pubsim.Util.erf(r2sdev.reciprocal, tol))
-  varnow = varnow*varstep
-  println(snrdb + "\t" + pe)
[…]

### 2012-01-23 plots/leechsim.scala

@@ -1,74 +1,0 @@
-/**
-* Probability of error plots for the lattice Zn.
-*/
-import java.math.BigInteger
-import pubsim.BigRational
-import pubsim.lattices.leech.Leech
-import pubsim.Util
-import pubsim.lattices.util.ProbabilityOfCodingError
-import pubsim.distributions.GaussianNoise
-import pubsim.lattices.GeneralLatticeAndNearestPointAlgorithm
-
-{
-println
-println("Running Monte-Carlo trials with the Eleech lattice")
-
-val lattice = new Leech
-val startdb = 5.5
-val stepdb = 0.5
-val tope = 1e-4
-
-val sm = lattice.secondMoment
-val decoder = new GeneralLatticeAndNearestPointAlgorithm(lattice.getGeneratorMatrix)
-
-var pe = 1.0
-var dbnow = startdb 
-val efile = new java.io.FileWriter("pemontecarlo_Leech_new")
-
-println
-println("SNRdB \t Prob Error")
-while(pe > tope){
-  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
-  val snr = scala.math.pow(10.0,dbnow/10.0)
-  val v = 4.0*sm/snr/24
-  pe = new ProbabilityOfCodingError(decoder,new GaussianNoise(0,v),100).probError
-  println(dbnow + "\t" + pe)
-  efile.write(dbnow.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
-  dbnow = dbnow + stepdb
-}
-efile.close
-}
-
-
-/*
-{
-val varstart = 1.0/2.0
-val varstep = 9.0/10.0
-val tope = 5e-11
-val petol = 1e-16
-
-var pe = 1.0
-var varnow = varstart 
-val mfile = new java.io.FileWriter("moment1")
-
-val tol = new BigRational(BigInteger.ONE, BigInteger.TEN.pow(15));
-
-println
-println("Caculating 'exact' error probabilty using the error function")
-println
-println("SNRdB \t Prob Error")
-while(pe > tope){
-  //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
-  val r2sdev = new BigRational(scala.math.sqrt(varnow*8.0) ,25)
-  val snr = 2.0/3.0/varnow
-  val snrdb = 10*scala.math.log10(snr)
[…]

### 2012-01-23 plots/proberrorsim.scala

@@ -5,13 +5,14 @@ import pubsim.lattices.util.PowerOfEuclideanNorm
 import pubsim.lattices.An.AnFastSelect
 import pubsim.lattices.An.MomentComputer
 import pubsim.BigRational
+import pubsim.Util
 
-val ns = List(2,3,4,5,6)
+val ns = List(2,3,4,5)
 //val lattice = new AnFastSelect(n)
 val varstart = new BigRational(1,2)
 val varstep = new BigRational(93,100)
-val tope = 5e-11
-val petol = 1e-16
+val tope = 1e-12
+val petol = 1e-17
 
 print("Reading existing moment data, this might take a moment....")
 val mc = new MomentComputer("momentstore")
@@ -22,12 +23,14 @@ for(n <- ns){
   var varnow = varstart 
   val mfile = new java.io.FileWriter("moment" + n)
 
+  val V = scala.math.sqrt(n+1.0)
+
   println
   println("Now running n = " + n)
   println("SNRdB \t Prob Error")
   while(pe > tope){
     //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
-    val snr = (3.0+n)/3.0/scala.math.sqrt(n+1)/varnow.doubleValue
+    val snr = scala.math.pow(V,2.0/n)/varnow.doubleValue/4
     val snrdb = 10*scala.math.log10(snr)
     pe = mc.ProbError(n,varnow,petol)
     varnow = varnow.multiply(varstep)

### 2012-01-23 plots/proberrorsimsphere.scala

@@ -1,45 +1,0 @@
-/**
-* Probability of error plots for the lattice An.
-*/
-import pubsim.lattices.util.PowerOfEuclideanNorm
-import pubsim.lattices.An.AnFastSelect
-import pubsim.lattices.An.MomentComputer
-import pubsim.BigRational
-import pubsim.Util
-
-val ns = List(2,3,4,5)
-//val lattice = new AnFastSelect(n)
-val varstart = new BigRational(1,2)
-val varstep = new BigRational(93,100)
-val tope = 1e-12
-val petol = 1e-17
-
-print("Reading existing moment data, this might take a moment....")
-val mc = new MomentComputer("momentstore")
-println("finished")
-
-for(n <- ns){
-  var pe = 1.0
-  var varnow = varstart 
-  val mfile = new java.io.FileWriter("moment" + n)
-
-  val V = scala.math.sqrt(n+1.0)
-
-  println
-  println("Now running n = " + n)
-  println("SNRdB \t Prob Error")
-  while(pe > tope){
-    //val snrdb = lattice.noiseVarianceToSNRdB(varnow.doubleValue)
-    val snr = scala.math.pow(V,2.0/n)/varnow.doubleValue/4
-    val snrdb = 10*scala.math.log10(snr)
-    pe = mc.ProbError(n,varnow,petol)
-    varnow = varnow.multiply(varstep)
-    println(snrdb + "\t" + pe + "\t" + mc.numberMomentsLastUsed)
-    mfile.write(snrdb.toString.replace('E', 'e') + "\t" + pe.toString.replace('E', 'e') + "\n")
-  }
-
-  mfile.close
-
-}
-
-mc.save

### 2012-01-23 plots/montecarlo.sh
more cleanup

@@ -1,2 +1,3 @@
 scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar montecarlo.scala
 scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar znsim.scala
+scala -nocompdaemon -cp PubSim.jar:Jama-1.0.2.jar:flanagan.jar:colt.jar:RngPack.jar e8sim.scala

### 2012-01-23 plots/e8sim.scala
minor edits to montecarlo simulations.

@@ -14,7 +14,7 @@ println
 println("Running Monte-Carlo trials with the E8 lattice")
 
 val lattice = new E8
-val startdb = 5.5
+val startdb = 10.0
 val stepdb = 0.5
 val tope = 1e-6
 

### 2012-01-23 plots/montecarlo.scala

@@ -7,7 +7,7 @@ import pubsim.distributions.GaussianNoise
 
 val ns = List(2,3,4,5,8)
 
-val startdb = 5.5
+val startdb = 10.0
 val stepdb = 0.5
 val tope = 1e-6
 

### 2012-01-23 plots/znsim.scala

@@ -14,7 +14,7 @@ println
 println("Running Monte-Carlo trials with the Zn lattice")
 
 val lattice = new Zn(1)
-val startdb = 0.0
+val startdb = 10.0
 val stepdb = 0.5
 val tope = 1e-6
 

### 2012-02-20 paper.tex
fixed silly notational error in section 2

@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal, onecolumn, 10pt]{IEEEtran}
 %\documentclass[journal]{IEEEtran}
@@
 \title{On the error performance of the $A_n$ lattices}
-\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson}
+\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson
+ \thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Emanuele Viterbo is with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.}
+}
 
@@
 \newcommand{\calX}{\mathcal X}
+\newcommand{\calC}{\mathcal C}
 Lattices can be used to construct \emph{signal space codes}.
-A signal space code $\calX$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.
-Each vector in $\calX$ is called a \emph{codeword} and represents a particular signal.
-The number of codewords is denoted by $\abs{\calX}$.
-If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calX}$ bits per codeword.
-The average power of the code is given by $P = \frac{1}{\abs{\calX}}\sum_{\xbf \in \calX}\|\xbf\|^2$.
+A signal space code $\calC$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.
+Each vector in $\calC$ is called a \emph{codeword} and represents a particular signal.
+The number of codewords is denoted by $\abs{\calC}$.
+If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calC}$ bits per codeword.
+The average power of the code is given by $P = \frac{1}{\abs{\calC}}\sum_{\cbf \in \calC}\|\cbf\|^2$.
 In the AWGN channel the received signal takes the form
 \[
-\ybf = \xbf + \wbf
+\ybf = \cbf + \wbf
 \]
-where $\ybf \in \reals^n$, $\xbf \in \calX$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
-If the receiver employs maximum likelihood decoding then the estimator of $\xbf$ given $\ybf$ at the receiver is
+where $\ybf \in \reals^n$, $\cbf \in \calC$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
+If the receiver employs maximum likelihood decoding then the estimator of $\cbf$ given $\ybf$ at the receiver is
 \begin{equation}\label{eq:mldecoder}
-\hat{\xbf} = \underset{\xbf \in \calX}{\operatorname{argmin}}\;\| \ybf - \xbf \|^2.
-%\hat{\xbf} &=& \mbox{arg~}\min_{\xbf \in \calX} \| \ybf - \xbf \|^2.
+\hat{\cbf} = \underset{\cbf \in \calC}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2.
+%\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
 \end{equation}
-That is, the receiver computes the codeword in $\calX$ nearest in Euclidean distance to the received signal $\ybf$.
+That is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
 Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
 \[
-P_C =\frac{1}{\abs{\calX}}\sum_{\xbf \in \calX} \prob( \hat{\xbf} = \xbf ),
+P_C =\frac{1}{\abs{\calC}}\sum_{\cbf \in \calC} \prob( \hat{\cbf} = \cbf ),
 \]
@@
 \[
-P_E = 1 - \frac{1}{\abs{\calX}} \sum_{\xbf \in \calX} \prob(
-\hat{\xbf} = \xbf ).
+P_E = 1 - \frac{1}{\abs{\calC}} \sum_{\cbf \in \calC} \prob(
+\hat{\cbf} = \cbf ).
 \]
@@
 %A problem with randomly generated codes and maximum likelihood decoding is that there are no computationally efficient methods for encoding and decoding.
-For encoding the transmitter must store a list of size $\abs{\calX} \approx 2^{nR/B}$ and for decoding the receiver must compute $\hat{\xbf}$ requiring a search over all $\abs{\calX} \approx 2^{nR/B}$ codewords.
+For encoding the transmitter must store a list of size $\abs{\calC} \approx 2^{nR/B}$ and for decoding the receiver must compute $\hat{\cbf}$ requiring a search over all $\abs{\calC} \approx 2^{nR/B}$ codewords.
 As $n$ gets large this becomes infeasible.
@@
 \[
-\calX = S \cap \Lambda.
+\calC = S \cap \Lambda.
 \]
@@
 
-Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.
-Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
+Let $\calC$ be a lattice code constructed from a lattice $\Lambda$.
+Assuming that each codeword $\cbf \in \calC$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
 %\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
@@
 \begin{align}
-P_C &\geq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
- \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
-\vor(\Lambda) + \xbf )  \nonumber \\
-&= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
+P_C &\geq P_C(\Lambda) = \lim_{\abs{\calC}\rightarrow \infty}
+ \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in
+\vor(\Lambda) + \cbf )  \nonumber \\
+&= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
 &=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}

### 2012-02-20 threepager.tex
Added to threepager.

--- 2012-02-20 paper.tex
@@
 %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal, onecolumn, 10pt]{IEEEtran}
-%\documentclass[journal]{IEEEtran}
+%\documentclass[journal, onecolumn, 10pt]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 %\documentclass[correspondence,9pt]{IEEEtran}
@@
 \begin{abstract}
-We consider the root lattice $A_n$ and derive explicit formulae for
-the moments of its Voronoi cell.  We then show that these formulae
-enable accurate prediction of the error probability of
-lattice codes constructed from $A_n$.
+We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.
+We then show that these formulae enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
 \end{abstract}
@@
 
-This paper is organised as follows.
+%This paper is organised as follows.
 In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.
@@
  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
-In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube.
+%In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube.
 These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.
@@
 $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
-For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix.
-%It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.
 
-The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\mathcal{S}_{\Lambda}$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point.
+The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\reals^m$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point.
 The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
-It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.
+If we translate the Voronoi cell by a lattice point $\xbf$, we obtain the reals of $\reals^m$ that is closer to $\xbf$ than any other lattice point.
+We write $\vor(\Lambda) + \xbf$ to denote the Voronoi cell translated about the lattice point $\xbf$.
+%It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.
 That is, the Voronoi cell tessellates when translated by points in $\Lambda$.
@@
 We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.
-
 The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, minimal vectors, center density, thickness, and the normalized second moment (or quantizing constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
@@
 
-The Voronoi cell is linked with the problem of \emph{lattice decoding}.
+%The Voronoi cell is linked with the problem of \emph{lattice decoding}.
 Given some point $\ybf \in \reals^n$ a \emph{lattice decoder} (or \emph{nearest lattice point algorithm}) returns the lattice point in $\Lambda$ that is nearest to $\ybf$~\cite{Agrell2002}.
@@
  %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder.
-A good overview of these techniques is given by Agrell~et.~al.~\cite{Agrell2002}.
+%A good overview of these techniques is given by Agrell~et.~al.~\cite{Agrell2002}.
 Fast nearest point algorithms are known for specific lattices~\cite{Conway1982FastQuantDec, McKilliam2008,McKilliam2009CoxeterLattices,Vardy1993_leech_lattice_MLD}.
@@
 \newcommand{\calC}{\mathcal C}
-Lattices can be used to construct \emph{signal space codes}.
-A signal space code $\calC$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.
-Each vector in $\calC$ is called a \emph{codeword} and represents a particular signal.
-The number of codewords is denoted by $\abs{\calC}$.
-If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calC}$ bits per codeword.
-The average power of the code is given by $P = \frac{1}{\abs{\calC}}\sum_{\cbf \in \calC}\|\cbf\|^2$.
-In the AWGN channel the received signal takes the form
+Lattices can be used to construct \emph{lattice codes}.
+A lattice code $\calC$ of dimension $n$ is a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
+There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$.
+The codewords are given by those lattice points inside the shaping region, that is, $\calC = S \cap \Lambda.$  Common choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes, or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
+
+Each vector in $\calC$ is called a \emph{codeword} and represents a particular signal.  The number of codewords is denoted by $\abs{\calC}$.  In the AWGN channel the received signal takes the form
 \[
@@
 where $\ybf \in \reals^n$, $\cbf \in \calC$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
-If the receiver employs maximum likelihood decoding then the estimator of $\cbf$ given $\ybf$ at the receiver is
+If the receiver employs \emph{lattice decoding} then the estimator of $\cbf$ given $\ybf$ at the receiver is
 \begin{equation}\label{eq:mldecoder}
-\hat{\cbf} = \underset{\cbf \in \calC}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2.
+\hat{\cbf} = \underset{\cbf \in \Lambda}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2.
 %\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
 \end{equation}
-That is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
-Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
-\[
-P_C =\frac{1}{\abs{\calC}}\sum_{\cbf \in \calC} \prob( \hat{\cbf} = \cbf ),
-\]
-and the probability of error is
-\[
-P_E = 1 - \frac{1}{\abs{\calC}} \sum_{\cbf \in \calC} \prob(
-\hat{\cbf} = \cbf ).
-\]
-%The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.
-If the codewords are chosen uniformly randomly on a sphere of radius $\sqrt{P}$ then as $n$ gets large the rate of the code can be made arbitrarily close to the channel capacity while simultaneously the probability of error can be made arbitrarily small.
-Another way of stating this is that we can choose close to $2^{nR/B}$ codewords uniformly randomly on a sphere of radius $\sqrt{P}$ and as $n$ gets large the probability of error decreases to zero.
-
-%A problem with randomly generated codes and maximum likelihood decoding is that there are no computationally efficient methods for encoding and decoding.
-For encoding the transmitter must store a list of size $\abs{\calC} \approx 2^{nR/B}$ and for decoding the receiver must compute $\hat{\cbf}$ requiring a search over all $\abs{\calC} \approx 2^{nR/B}$ codewords.
-As $n$ gets large this becomes infeasible.
-To avoid this computational complexity it is preferable for a code to have some type of `structure' that is exploitable for efficient encoding and decoding.
-One way of enforcing structure is to choose codewords from a lattice.
-The idea being that encoding and decoding can now be performed efficiently by \emph{lattice decoding}, i.e.
-by finding a nearest point in a lattice.
-
-A \emph{lattice code} is a signal space code with codewords taken from a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
-There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$.
-The codewords are given by those lattice points inside the shaping region, that is,
-\[
-\calC = S \cap \Lambda.
-\]
-Common choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes, or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
-A consequence of the finiteness of a lattice code is that the maximum likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to lattice decoding.
-Computing a nearest lattice point will not in general return a lattice point from the code (the decoded lattice point might lie outside the shaping region).
- %A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.
-This question was answered in the affirmative by Erez and Zamir~\cite{Erex2004_lattice_decoding}.
-
-Let $\calC$ be a lattice code constructed from a lattice $\Lambda$.
-Assuming that each codeword $\cbf \in \calC$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
-%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
-A result of this is that that observed noise distribution is no longer Gaussian.
-However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.
-In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
+That is, the receiver computes the lattice point in $\Lambda$ nearest in Euclidean distance to the received signal $\ybf$.
+Assuming that each codeword $\cbf$ is transmitted with probability $p(\cbf)$, then the probability of correct decoding is
 \begin{align}
-P_C &\geq P_C(\Lambda) = \lim_{\abs{\calC}\rightarrow \infty}
- \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in
-\vor(\Lambda) + \cbf )  \nonumber \\
+P_C &= \sum_{\cbf \in \calC} p(\cbf) \prob( \hat{\cbf} = \cbf ),
+&= \sum_{\cbf\in \calC} p(\cbf) \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
 &= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
-&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
-e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
+&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
 \end{align}
-The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).
-By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
+The probability of error is $P_E = 1 - P_C$.  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
 \begin{align}
@@
 \]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
-\begin{align*}
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.
+The first five such formula are displayed in~\eqref{eq:momentsAnformn} on~\pageref{eq:momentsAnformn}.
+We have explicitly tabulated these formula for $m=0$ to $40$.
+For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.
+%We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.
+
+\begin{figure*}[B]
+\begin{equation} \label{eq:momentsAnformn}
+\begin{split}
 \calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
@@
 \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
-\end{align*}
-We have explicitly tabulated these formula for $m=0$ to $40$.
-For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.
-We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.
-
-\section{The lattice $A_n$}\label{sec:lattice-a_n}
-Let $H$ be the hyperplane orthogonal to the all ones vector of length $n+1$, denoted by $\onebf$, that is
-\[
-\onebf = \left[ \begin{array}{cccc} 1 & 1 & \cdots & 1 \end{array} \right]^\prime,
-\]
-where superscript $^\prime$ indicates the transpose.
-Any vector in $H$ has the property that the sum (and therefore the mean) of its elements is zero and for this reason $H$ is often referred to as the \emph{zero-sum plane} or the \emph{zero-mean plane}.
-The lattice $A_n$ is the intersection of the integer lattice $\ints^{n+1}$ with the zero-sum plane, that is
-\begin{equation}
-\label{eq:An_sub_Zn}
-  A_{n} = \ints^{n+1} \cap H = \big\{ \xbf \in \ints^{n+1} \mid \dotprod{\xbf}{\onebf} = 0  \big\}.
+\end{split}
 \end{equation}
-Equivalently, $A_n$ consists of all of those points in $\ints^{n+1}$ with coordinate sum equal to zero.
-% A generator matrix for $A_n$ is the $(n+1)\times n$ matrix
-% \[
-% \left[ \begin{array}{rrrrrrr}
-% 1 & 0 & 0 &  \cdots & 0 & 0 \\
-% -1 & 1 & 0 &  \cdots & 0 & 0 \\
-% 0 & -1 & 1 &  \cdots & 0 & 0 \\
-% 0 & 0 & -1 & \cdots & 0 & 0 \\
-% \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
-% 0 & 0 & 0 & \cdots & -1 & 1 \\
-% 0 & 0 & 0 & \cdots & 0 & -1  \\
-% \end{array} \right].
-% \]
-The lattice has $n(n+1)$ minimal vectors, each of squared Euclidean length $2$, so the packing radius is $\frac{1}{\sqrt{2}}$.
-The $n$-volume of the Voronoi cell $\vor(A_n)$ is $\sqrt{n+1}$~\cite[p.
-108]{SPLAG}.
-
-The Voronoi cell of $A_n$ is closely related to the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$ as the next theorem will show.
-This result has appeared previously~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}, but we repeat it here so that this paper is self contained.
-We denote by
-\[
-\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{\onebf^\prime \onebf} = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
-\]
-the projection matrix orthogonal to $\onebf$ (i.e.
-into the zero-sum plane) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.
-Given a set $S$ of vectors from $\reals^{n+1}$ we write $\Qbf S$  to denote the set with elements $\Qbf \sbf$ for all $\sbf \in S$, i.e.
-the set containing the projection of the vectors from $S$.
-
-\begin{lemma} \label{lem:QVorZnsubsetVorAn}
-The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is a subset of $\vor(A_{n})$.  That is,
-\[
-\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_{n}).
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $\ybf \in \vor(\ints^{n+1})$.
-Decompose $\ybf$ into orthogonal components so that $\ybf = \Qbf \ybf + t \onebf$ for some $t \in \reals$.
-Then $\Qbf\ybf \in \Qbf\vor(\ints^{n+1})$.
-Assume that $\Qbf\ybf \notin \vor(A_n)$.
-Then there exists some $\xbf \in A_n$ such that
-\begin{align*}
-\|\xbf - \Qbf\ybf\|^2 < \|\zerobf - \Qbf\ybf\|^2 & \Rightarrow \|\xbf - \ybf + t\onebf\|^2 < \|\ybf - t\onebf\|^2 \\
-%& \Rightarrow \|\xbf - \ybf\|^2 + 2t(\xbf - \ybf)'\onebf + t^2\|\onebf\|^2 < \|\ybf\|^2 - 2t\ybf'\onebf + t^2\|\onebf\|^2 \\
-& \Rightarrow \|\xbf - \ybf\|^2 + 2t\xbf'\onebf < \|\ybf\|^2.
-\end{align*}
-By definition \eqref{eq:An_sub_Zn} $\xbf'\onebf = 0$ so $\|\xbf - \ybf\|^2 < \|\ybf\|^2$.
-This violates that $\ybf \in \vor(\ints^{n+1})$ and hence $\Qbf\ybf \in \vor(A_n)$.\footnote{This proof can be generalised to show that for any lattice $L$ and hyperplane $P$ such that $P\cap L$ is also a lattice it is true that $p\vor(L) \subseteq \vor(L \cap P)$ where $p$ indicates the orthogonal projection into $P$~\cite[Lemma~2.1]{McKilliam2010thesis}.}
-\end{IEEEproof}
-
-\begin{theorem}  \label{thm:VorAn=QVorZn1}
-The projection of $\vor(\ints^{n+1})$ into the zero-sum plane is equal to $\vor(A_{n})$. That is,
-\[
-\vor(A_n) = \Qbf\vor(\ints^{n+1}).
-\]
-\end{theorem}
-\begin{proof}
-Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.
-The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.
-From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
-\[
-\sum_{i = 1}^{n+1} \frac{\dotprod{\onebf}{\ebf_i}}{\|\onebf\|} =  \sum_{i = 1}^{n+1} \frac{1}{\sqrt{n+1}} = \sqrt{n + 1}
-\]
-also.
-It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
-\end{proof}
-
-\section{Integrating a function over $\vor(A_n)$}\label{sec:integr-funct-over}
-
-We would like to be able to integrate functions over the Voronoi cell of $A_n$.
-Consider a function $f : \reals^{n+1} \mapsto \reals$.
-The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.
-The Voronoi cell is a subset of the zero-sum plane that has zero $(n+1)$-dimensional volume.
-So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.
-This is not what we intend.
-By an appropriate change of variables it would be possible to write the Voronoi cell $\vor(A_n)$ in an $n$-dimensional coordinate system, and then integrate.
-However, we find the following approach simpler.
-Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
-Now, the integral over the Voronoi cell can be written as
-\begin{equation}\label{eq:vorfint}
-\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
-\end{equation}
-It is not immediately clear how an integral over $\vor(A_n) \times \onebf$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
-\begin{equation}\label{eq:unormalised}
-\int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
-\end{equation}
-This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.
-To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.
-Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.
-The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by
-\[
-\|\onebf\|(1 - x_{\text{max}} + x_{\text{min}}) = \sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}}).
-\]
-The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
-\[
-\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
-\]
-The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.
-
-Let us now restrict $f(\xbf)$ so that it depends only on the magnitude $\|\xbf\|$, for example $f(\xbf) = \|\xbf\|^{2m}$ could be a power of the Euclidean norm of $\xbf$.
-Now $f(\xbf)$ is invariant to permutation of $\xbf$.
-Let $\xbf$ be such that $x_1$ is the maximum element and $x_2$ is the minimum element.
-Our integral is now equal to
-\[
-\frac{n (n+1)}{\sqrt{n+1}} \int^{1/2}_{-1/2} \int^{x_1}_{-1/2}  \int^{x_1}_{x_2} \cdots \int^{x_1}_{x_2} \frac{f(\Qbf\xbf)}{1 - x_{1} + x_{2}}  dx_{n+1} \, \dots \, dx_2 \, dx_1.
-\]
-The factor $n(n+1)$ arises because there are $n(n+1)$ ways to place two elements (i.e. $x_1$ and $x_2$) into $n+1$ positions.
-
-We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables,
-\[
-n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} \frac{f(\Qbf\xbf )}{1 - t}  dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
-\]
-and letting $w_{i-2} = x_i - y + 1/2 + t$ for $i = 3,\dots,n+1$ gives
-\begin{equation}\label{eq:intwewilluse}
-n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{t}_{0} \cdots \int^{t}_{0} \frac{f(\Qbf\xbf )}{1 - t}  dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
-\end{equation}
-Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector
-\[
-\wbf = [t, 0, w_1, w_2, \dots, w_{n-1}]^\prime.
-\]
-Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.
-Interestingly $\wbf$ does not contain $y$ so the term inside the integral~\eqref{eq:intwewilluse} does not depend on $y$.
-This is the integral we will use to compute the moments of $A_n$.
-
-\begin{example}\emph{\textbf{(The volume of the Voronoi cell)}
-In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting $f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1$ we obtain,
-\begin{align*}
-M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
- &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy. \\
-&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1}.
-\end{align*}
-as required.
-Here $\beta(x,a,b) =  \int_{0}^{x}  t^{a-1}(1 - t)^{b-1} \,dt$ is the incomplete beta function~\cite{Pearson_tables_of_beta_functions} and we have used the identity $\int^{1}_{0} \beta(y, n, 0) \, dy = \frac{1}{n}$.}
-\end{example}
-
-% \subsection{The second moment of the Voronoi cell}
-
-% Setting
-% \begin{align*}
-% f(\Qbf\wbf) = \|\Qbf\wbf\|^2  &= \|\wbf\|^2 - \frac{(\wbf^\prime \onebf)^2}{n+1} \\
-% &= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t + \sum_{i=1}^{n+1}w_i \right)^2 \\
-% &= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n+1}w_i + \left(\sum_{i=1}^{n+1}w_i \right)^2 \right)\\
-% &= c + d
-% \end{align*}
-% where
-% \[
-% c = \frac{n}{n+1} t^2,\qquad \text{and} \qquad d = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
-% \]
-% where $A = \sum_{i=1}^{n-1}w_i^2$ and $B = \sum_{i=1}^{n-1}w_i$.  Now the second moment is given by
-% \[
-% C_1 = n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{c + d}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy.
-% \]
-% Integrating $A$, $B$ and $B^2$ over the $w_1,\dots,w_{n-1}$ gives,
-% \begin{align*}
-% &\int^{t}_{0} \cdots \int^{t}_{0} A \,dw_{n-1}\dots dw_1 = \frac{n-1}{3}t^{n+1},\\
-% &\int^{t}_{0} \cdots \int^{t}_{0} B \,dw_{n-1}\dots dw_1 = \frac{n-1}{2}t^n, \,\,\,\, \text{and} \\
-% &\int^{t}_{0} \cdots \int^{t}_{0} B^2 \,dw_{n-1}\dots dw_1 = \frac{(n-1)(3n - 2)}{12} t^{n+1}
-% \end{align*}
-% So integrating $d$ over the $w_1,\dots,w_{n-1}$ gives,
-% \[
-% \int^{t}_{0} \cdots \int^{t}_{0} d \,dw_{n-1}\dots dw_1 =  \frac{7(n-1)(n-2)}{12(n+1)}t^{n+1}.
-% \]
-% Also,
-% \[
-% \int^{t}_{0} \cdots \int^{t}_{0} c \,dw_{n-1}\dots dw_1 =  \frac{n}{n+1}t^{n+1}.
-% \]
-% So,
-% \begin{align*}
-% C_1 &=  \sqrt{n+1}\frac{n(12n + 7(n-1)(n-2))}{12(n+1)} \int^{1}_{0} \int^{y}_{0}\frac{t^{n+1}}{1 - t}  \, dt  \, dy \\
-% &= \sqrt{n+1}\frac{12n + 7(n-1)(n-2)}{12(n+1)}
-% \end{align*}
-
-%\newcommand{\calV}{\mathcal{V}}
-\section{The moments of $A_n$}\label{sec:powers-eucl-norm}
-
-We now derive expressions for the $\calM_n(m)$.  Setting $f(\Qbf\xbf) = \left(\|\Qbf\wbf\|^{2}\right)^m$ in~\eqref{eq:intwewilluse} we obtain,
-\[
-\frac{\calM_n(m)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
-\]
-Now $\|\Qbf\wbf\|^2 = \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2$ and recalling that $\wbf = [t,0,w_1,\dots,w_{n-1}]'$ we can write
-\begin{align*}
-\|\Qbf\wbf\|^2 &= \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2 \\
-&= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t + \sum_{i=1}^{n+1}w_i \right)^2 \\
-&= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n+1}w_i + \left(\sum_{i=1}^{n+1}w_i \right)^2 \right)\\
-&= C + D,
-\end{align*}
-say, where
-\[
-C = \left(\frac{n}{n+1} \right) t^2 \qquad \text{and} \qquad D = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
-\]
-and where,
-\[
- A = \sum_{i=1}^{n-1}w_i^2 \qquad \text{and} \qquad B = \sum_{i=1}^{n-1}w_i.
-\]
-Now,
-\[
- \frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(C+D)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
-\]
-and by expanding the binomial $(C+D)^m$ we get
-\[
-\frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} C^{m-k} \int_0^t \cdots \int_0^t \, D^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
-\]
-Expanding $D^k$ as a trinomial gives
-\begin{align*}
-D^k &= \sum_{k_1+k_2+k_3=k} \frac{k! A^{k_1} B^{2k_3 + k_2}}{k_1! k_2! k_3!} \left(\frac{-1}{n+1}\right)^{k_2+k_3}2^{k_2}t^{k_2}  \\
-&=  \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{k!  A^{a} B^{2k - 2a - b}}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}2^{b}t^{b}
-\end{align*}
-where the second line follows by setting $k_1 = a$, $k_2 = b$ and $k_3 = k - a - b$.
-In Appendix~\ref{sec:mult-type-integr} we show that the integral of $A^{a} B^{2k - 2a - b}$ over $w_1,\dots w_{n-1}$ is
-\begin{equation} \label{eq:recurseint}
-\int_0^t \cdots \int_0^t A^{a} B^{2k-2a-b} dw_{n-1} \, \dots \, dw_1 = t^{n-1+2k - b} G(n-1,a,2k - 2a - b).
-\end{equation}
-where $G(n,c,d)$ satisfies the recursion given by~\eqref{eq:theGrecursion}.  So, let $P$ satisfy
-\begin{align*}
-P &= t^{1 - n - 2k}\int_0^t \cdots \int_0^t D^k dw_{n-1} \, \dots \, dw_1 \\
-&= \sum_{a=0}^k\sum_{b=0}^{k-a} \frac{2^{b} k!G(n-1,a,2k - 2a - b)}{a! b! (k-a-b)!} \left(\frac{-1}{n+1}\right)^{k - a}.
-\end{align*}
-Now $C^{m-k} = \left(\frac{n}{n+1} \right)^{m-k} t^{2(m-k)}$ and
-\begin{align*}
-\frac{\calM_n(m)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy \\
-&= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \beta(y,n+2m,0) dy \\
-&= \frac{1}{n+2m} \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P.
-\end{align*}
-This expression is equivalent to that from~(\ref{eq:theCmformula}).
+\end{figure*}
 
@@
 
-\appendix
-
-\subsection{A multinomial type integral over a hypercube}\label{sec:mult-type-integr}
-
-In~(\ref{eq:recurseint}) we required to evaluate integrals of the form
-\[
-F(n-1,a,2k - 2a - b) = \int^{t}_{0} \cdots \int^{t}_{0} A^a B^{2k - 2a - b} \,dw_{n-1}\cdots dw_{1},
-\]
-or equivalently, integrals of the form
-\[
-F(n,c,d) = \int^{t}_{0} \cdots \int^{t}_{0} \left(\sum_{j=1}^{n} x_j^2\right)^c \left( \sum_{i=1}^{n} x_i \right)^d \,dx_1\cdots dx_{n}
-\]
-where $n,c$ and $d$ are integers.  We will find a recursion describing this integral.  Write
-\[
-F(n,c,d) = \int^t_0 \cdots \int^t_0 \left(x_n^2 + \sum_{j=1}^{n-1} x_j^2\right)^c \left( x_n + \sum_{i=1}^{n-1} x_i \right)^d \,dx_1\cdots dx_n.
-\]
-Expanding the two binomials gives
-\begin{align*}
-F(n,c,d) &=  \int^t_0 \cdots \int^t_0 \sum_{c'=0}^{c} \binom{c}{c'} x_n^{2c'} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'}  \sum_{d'=0}^{d}\binom{d}{d'} x_n^{d'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n} \\
-&=  \sum_{c'=0}^{c} \sum_{d'=0}^d \int^t_0 \cdots \int^t_0 \binom{c}{c'} \binom{d}{d'} x_n^{2c'+d'} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n}.
-\end{align*}
-Integrating the $x_n$ term gives
-\[
-F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1} \int^t_0 \cdots \int^t_0 \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
-\]
-Note that
-\[
-F(n-1,c-c',d-d')  = \int^t_0 \cdots \int^t_0 \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
-\]
-So $F(n,c,d)$ satisfies the recursion
-\[
-F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1}F(n-1,c-c',d-d')
-\]
-with the initial conditions
-\[
-F(1,c,d) = \frac{t^{2c+d+1}}{2c+d+1} \qquad \text{and} \qquad F(n,0,0) = t^n.
-\]
-The $F(n,c,d)$ can be written as $t^{n+2c+d}G(n,c,d)$ where $G(n,c,d)$ is rational.  To see this write
-\begin{align*}
-F(n,c,d) &=  \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1}t^{n-1+2(c-c')+d-d'}G(n-1,c-c',d-d') \\
-&=  t^{n+2c+d}\sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1} \\
-&=  t^{n+2c+d}G(n,c,d).
-\end{align*}
-Now $G(n,c,d)$ is the rational number satisfying the recursion
-\[
-G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
-\]
-with the initial conditions
-\[
-G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
-\]
-
-\subsection{Solving this recursion for fixed $d$ and $c$}\label{sec:solv-this-recurs}
-\newcommand{\calG}{\mathcal G}
-
-For fixed $d$ and $c$ this recursion can be solved explicitly.  Write
-\[
-G(n,c,d) = G(n-1,c,d) + \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
-\]
-where the sum $\sum_{ (c',d') \neq (0,0)}$ is over all $0 \leq c' \leq c$ and $0 \leq d' \leq d$ except when both $d$ and $c$ are zero.
-Denote by $\calG(z,c,d)$ the $z$-transform of $G(n,c,d)$.
-Taking the $z$-transform of both sides in the equation above gives
-\[
-\calG(z,c,d) = \frac{z^{-1}}{1-z^{-1}} \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{\calG(z,c-c',d-d')}{2c'+d'+1}.
-\]
-So the $z$-transform $\calG(z,c,d)$ satisfies this recursive equation.
-The initial condition is $\calG(z,0,0) = \frac{z^{-1}}{1 - z^{-1}}$.
-By inverting this $z$-transform and using the resultant expressions in~\eqref{eq:theCmformula} we obtain formulae in $n$ for the moment $\calM_n(m)$.
-This procedure was used to generate the formula described in Section~\ref{sec:main-result}.
-Mathematica 8.0 was used to perform these calculations.
-
-% \section{Conclusion}
-% \begin{itemize}
-% \item We can use these techniques to compute other things about the Voronoi cell, i.e. the average $L_p$ norms.
-% \item Not restricted to Gaussian noise.  Any noise pdf that can be described by the moments will work.
-% \item Can other lattices be done?
-% \end{itemize}
-
 \small

### 2012-02-20 threepager.tex
modified some srcipts

@@
 Emanuele Viterbo is with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
-Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.}
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+A preliminary full version of this paper is available at http://arxiv.org/abs/1111.6334.}
 }
@@
 
-The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\reals^m$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point.
+The \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\reals^m$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point.
 The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
-If we translate the Voronoi cell by a lattice point $\xbf$, we obtain the reals of $\reals^m$ that is closer to $\xbf$ than any other lattice point.
+If we translate the Voronoi cell by a lattice point $\xbf$, we obtain the subset of $\reals^m$ that is closer to $\xbf$ than any other lattice point.
 We write $\vor(\Lambda) + \xbf$ to denote the Voronoi cell translated about the lattice point $\xbf$.
@@
 A lattice code $\calC$ of dimension $n$ is a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
+Each point in $\calC$ is called a \emph{codeword} and represents a particular signal.
 There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$.
@@
 
-Each vector in $\calC$ is called a \emph{codeword} and represents a particular signal.  The number of codewords is denoted by $\abs{\calC}$.  In the AWGN channel the received signal takes the form
+In the AWGN channel the received signal takes the form
 \[
@@
 where $\ybf \in \reals^n$, $\cbf \in \calC$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
-If the receiver employs \emph{lattice decoding} then the estimator of $\cbf$ given $\ybf$ at the receiver is
+If the receiver employs maximum likelihood decoding then the estimator of $\cbf$ given $\ybf$ at the receiver is
 \begin{equation}\label{eq:mldecoder}
-\hat{\cbf} = \underset{\cbf \in \Lambda}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2.
+\hat{\cbf}_{\text{ML}} = \underset{\cbf \in C}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2.
+%\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
+\end{equation}
+That is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
+Typically maximum likelihood decoding is computationally complex and it is preferable to use \emph{lattice deoding}~\cite{Erex2004_lattice_decoding}.
+The estimator of $\cbf$ is then,
+\begin{equation}\label{eq:latticedecoder}
+\hat{\cbf} = \underset{\cbf \in \Lambda}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
 %\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
@@
 That is, the receiver computes the lattice point in $\Lambda$ nearest in Euclidean distance to the received signal $\ybf$.
+Equivalently, $\hat{\cbf}$ is the lattice point such that $\ybf \in \vor(\Lambda) + \hat{\cbf}$.
+The codeword is correctly decoded when $\hat{\cbf} = \cbf$, or equivalently when $\ybf \in \vor(\Lambda) + \cbf$ implying that $\wbf \in \vor(\Lambda)$.
 Assuming that each codeword $\cbf$ is transmitted with probability $p(\cbf)$, then the probability of correct decoding is
 \begin{align}
-P_C &= \sum_{\cbf \in \calC} p(\cbf) \prob( \hat{\cbf} = \cbf ),
+P_C &= \sum_{\cbf \in \calC} p(\cbf) \prob( \hat{\cbf} = \cbf ) \\
 &= \sum_{\cbf\in \calC} p(\cbf) \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
@@
 \begin{align}
-P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
+P_C  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
@@
 
+\begin{figure*}[htbp]
+	\centering
+		\includegraphics{plots/threepagerpeplot-1.mps}
+		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5, A_8$ and $E_8$.}
+		\label{fig:peplots}
+\end{figure*}
+
+\begin{figure*}[htbp]
+\begin{equation*} \label{eq:momentsAnformn}
+\begin{split}
+\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),
+\\ \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, 
+\\ \calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, 
+\\ \calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, 
+\\ \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}},
+%\\ \calM_n(5) &= \frac{3577728 n - 1825648 n^2 + 2410804 n^3 + 1569392 n^4 + 1644423 n^5 + 906105 n^6 + 341550 n^7 + 75526 n^8 + 8855 n^9 + 385 n^{10}}{95800320 (1 + n)^{9/2}}.
+\end{split}
+\end{equation*}
+\end{figure*}
+
 \section{The main result}\label{sec:main-result}
@@
 For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.
-The first five such formula are displayed in~\eqref{eq:momentsAnformn} on~\pageref{eq:momentsAnformn}.
+The first five such formula are displayed on page~\pageref{eq:momentsAnformn}.
 We have explicitly tabulated these formula for $m=0$ to $40$.
@@
 %We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.
-
-\begin{figure*}[B]
-\begin{equation} \label{eq:momentsAnformn}
-\begin{split}
-\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
-\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, \\
-\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
-\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
-\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
-\end{split}
-\end{equation}
-\end{figure*}
 
@@
 
-\begin{figure*}[tp]
-	\centering
-		\includegraphics{plots/peplots-1.mps}
-		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5, A_8$ and $E_8$.}
-		\label{fig:peplots}
-\end{figure*}
-
 \section{Conclusion}

### 2012-02-20 build.sh

@@ -1,10 +1,10 @@
-cd plots
-mpost -interaction=nonstopmode peplots.mp
-cd ..
+cd plots   
+mpost -interaction=nonstopmode peplots.mp   
+cd ..   
 
-latex paper.tex
-bibtex paper
-latex paper.tex
-latex paper.tex
-dvips paper.dvi
-ps2pdf paper.ps
+latex paper.tex  
+bibtex paper  
+latex paper.tex  
+latex paper.tex  
+dvips paper.dvi  
+ps2pdf paper.ps  

### 2012-02-20 build.sh
Added to build and threepager

@@ -1,10 +1,13 @@
-cd plots   
-mpost -interaction=nonstopmode peplots.mp   
-cd ..   
+cd plots
+mpost -interaction=nonstopmode peplots.mp
+cd ..
 
-latex paper.tex  
-bibtex paper  
-latex paper.tex  
-latex paper.tex  
-dvips paper.dvi  
-ps2pdf paper.ps  
+pdflatex paper.tex
+bibtex paper
+pdflatex paper.tex
+pdflatex paper.tex
+
+pdflatex threepager.tex
+bibtex threepager
+pdflatex threepager.tex
+pdflatex threepager.tex

### 2012-02-20 threepager.tex
Added to build and threepager
minor threep edits.

@@
 We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.
-We then show that these formulae enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
+These formulae enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
 \end{abstract}
@@
 In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
-A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional.
+%A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional.
 If the generator is square, i.e.
@@
 If we translate the Voronoi cell by a lattice point $\xbf$, we obtain the subset of $\reals^m$ that is closer to $\xbf$ than any other lattice point.
-We write $\vor(\Lambda) + \xbf$ to denote the Voronoi cell translated about the lattice point $\xbf$.
+We write $\vor(\Lambda) + \xbf$ to denote the Voronoi cell translated by $\xbf$.
 %It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.
@@
 \begin{equation}\label{eq:mldecoder}
-\hat{\cbf}_{\text{ML}} = \underset{\cbf \in C}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2.
+\hat{\cbf}_{\text{ML}} = \underset{\cbf \in C}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
 %\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
 \end{equation}
-That is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
+that is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
 Typically maximum likelihood decoding is computationally complex and it is preferable to use \emph{lattice deoding}~\cite{Erex2004_lattice_decoding}.
@@
 \end{equation}
-That is, the receiver computes the lattice point in $\Lambda$ nearest in Euclidean distance to the received signal $\ybf$.
+that is, the receiver computes the lattice point in $\Lambda$ nearest in Euclidean distance to the received signal $\ybf$.
 Equivalently, $\hat{\cbf}$ is the lattice point such that $\ybf \in \vor(\Lambda) + \hat{\cbf}$.
-The codeword is correctly decoded when $\hat{\cbf} = \cbf$, or equivalently when $\ybf \in \vor(\Lambda) + \cbf$ implying that $\wbf \in \vor(\Lambda)$.
-Assuming that each codeword $\cbf$ is transmitted with probability $p(\cbf)$, then the probability of correct decoding is
+Note that with lattice decoding the decoded lattice point $\hat{\cbf}$ is not guaranteed to be inside $\calC$.
+
+Correct decoding occurs when $\hat{\cbf} = \cbf$, or equivalently when $\ybf \in \vor(\Lambda) + \cbf$, or equivalently when $\wbf \in \vor(\Lambda)$, i.e.
+when the noise $\wbf$ is inside the Voronoi cell of the lattice.
+Assuming that each codeword $\cbf$ is transmitted with probability $p(\cbf)$, then the probability of correct lattice decoding is
 \begin{align}
@@
 So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
-The number of terms required will increase as the noise variance gets smaller.
-This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
+The number of terms required increases as the noise variance gets smaller.
+%This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
 
@@
 	\centering
-		\includegraphics{plots/threepagerpeplot-1.mps}
+		\includegraphics{plots/peplots-1.mps}
 		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5, A_8$ and $E_8$.}
@@
 \\ \calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, 
-\\ \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}},
+\\ \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
 %\\ \calM_n(5) &= \frac{3577728 n - 1825648 n^2 + 2410804 n^3 + 1569392 n^4 + 1644423 n^5 + 906105 n^6 + 341550 n^7 + 75526 n^8 + 8855 n^9 + 385 n^{10}}{95800320 (1 + n)^{9/2}}.

### 2012-02-21 other/threepager/threepager.tex
added 3page summary of this paper.

--- 2012-02-20 threepager.tex
@@
 %\documentclass[journal, onecolumn, 10pt]{IEEEtran}
-\documentclass[journal]{IEEEtran}
+\documentclass[journal]{../../IEEEtran}
 %\documentclass[correspondence,9pt]{IEEEtran}
 
-\usepackage{mathbf-abbrevs}
+\usepackage{../../mathbf-abbrevs}
 \usepackage{amsmath}
@@
 The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
-The highly symmetric structure of these lattices provides the grounds for extremely efficient encoding and decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
+The highly symmetric structure of these lattices provides the grounds for extremely efficient encoding and decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.
 In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
@@
 The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.
-There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Sikiric_vor_reduction_covering_2008,Valentin2003_coverings_tilings_low_dimension}.
+There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Valentin2003_coverings_tilings_low_dimension}.
 In general these algorithms are only computationally feasible when the dimension is small (approximately $n \leq 9$).
@@
 that is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
-Typically maximum likelihood decoding is computationally complex and it is preferable to use \emph{lattice deoding}~\cite{Erex2004_lattice_decoding}.
+Typically maximum likelihood decoding is computationally complex and it is preferable to use \emph{lattice decoding}~\cite{Erex2004_lattice_decoding}.
 The estimator of $\cbf$ is then,
@@
 when the noise $\wbf$ is inside the Voronoi cell of the lattice.
-Assuming that each codeword $\cbf$ is transmitted with probability $p(\cbf)$, then the probability of correct lattice decoding is
+Assuming that the codeword $\cbf$ is transmitted with probability $p(\cbf)$, then the probability of correct lattice decoding is
 \begin{align}
@@
 So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
-The number of terms required increases as the noise variance gets smaller.
+%The number of terms required increases as the noise variance gets smaller.
 %This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
+We call these the \emph{moments} of the $\vor(\Lambda)$.
 
 \newcommand{\calM}{M}
-In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
+In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and we derive expressions for the moments
 \[
@@
 	\centering
-		\includegraphics{plots/peplots-1.mps}
+		\includegraphics{../../plots/thpplot-1.mps}
 		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5, A_8$ and $E_8$.}
@@
 
-\begin{figure*}[htbp]
-\begin{equation*} \label{eq:momentsAnformn}
-\begin{split}
-\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),
-\\ \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, 
-\\ \calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, 
-\\ \calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, 
-\\ \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
-%\\ \calM_n(5) &= \frac{3577728 n - 1825648 n^2 + 2410804 n^3 + 1569392 n^4 + 1644423 n^5 + 906105 n^6 + 341550 n^7 + 75526 n^8 + 8855 n^9 + 385 n^{10}}{95800320 (1 + n)^{9/2}}.
-\end{split}
-\end{equation*}
-\end{figure*}
+% \begin{figure*}[htbp]
+% \begin{equation*} \label{eq:momentsAnformn}
+% \begin{split}
+% \calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),
+% \\ \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, 
+% \\ \calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, 
+% \\ \calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, 
+% \\ \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
+% %\\ \calM_n(5) &= \frac{3577728 n - 1825648 n^2 + 2410804 n^3 + 1569392 n^4 + 1644423 n^5 + 906105 n^6 + 341550 n^7 + 75526 n^8 + 8855 n^9 + 385 n^{10}}{95800320 (1 + n)^{9/2}}.
+% \end{split}
+% \end{equation*}
+% \end{figure*}
+
+\section{The lattice $A_n$}\label{sec:lattice-a_n}
+Let $H$ be the hyperplane orthogonal to the all ones vector of length $n+1$, denoted by $\onebf$, that is
+\[
+\onebf = \left[ \begin{array}{cccc} 1 & 1 & \cdots & 1 \end{array} \right]^\prime,
+\]
+where superscript $^\prime$ indicates the transpose.
+Any vector in $H$ has the property that the sum of its elements is zero and for this reason $H$ is often referred to as the \emph{zero-sum plane}.
+The lattice $A_n$ is the intersection of the integer lattice $\ints^{n+1}$ with the zero-sum plane, that is
+\begin{equation}
+\label{eq:An_sub_Zn}
+  A_{n} = \ints^{n+1} \cap H = \big\{ \xbf \in \ints^{n+1} \mid \dotprod{\xbf}{\onebf} = 0  \big\}.
+\end{equation}
+Equivalently, $A_n$ consists of all of those points in $\ints^{n+1}$ with coordinate sum equal to zero.
+% A generator matrix for $A_n$ is the $(n+1)\times n$ matrix
+% \[
+% \left[ \begin{array}{rrrrrrr}
+% 1 & 0 & 0 &  \cdots & 0 & 0 \\
+% -1 & 1 & 0 &  \cdots & 0 & 0 \\
+% 0 & -1 & 1 &  \cdots & 0 & 0 \\
+% 0 & 0 & -1 & \cdots & 0 & 0 \\
+% \vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
+% 0 & 0 & 0 & \cdots & -1 & 1 \\
+% 0 & 0 & 0 & \cdots & 0 & -1  \\
+% \end{array} \right].
+% \]
+In two dimensions $A_2$ is the hexagonal lattice (Figure~\ref{fig:hexlatfig}), and in three dimension $A_3$ is the body centered cubic lattice~\cite[p.
+108]{SPLAG}.
+These are the densest known sphere packings in dimensions two and three.
+
+\begin{figure}[tbp]
+	\centering
+		\includegraphics[width=\linewidth]{hexlatfig-1.mps}
+		\caption{The hexagonal lattice $A_2$.  The Voronoi cell is a regular hexagon.}
+		\label{fig:hexlatfig}
+\end{figure}
 
@@
 
-We now state our main result.  The $m$th moment $\calM_n(m)$ of the lattice $A_n$ satisfies
+We now state our main result.  The moment $\calM_n(m)$ of the lattice $A_n$ satisfies
 \begin{equation}\label{eq:theCmformula}
@@
 \]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.
-The first five such formula are displayed on page~\pageref{eq:momentsAnformn}.
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first three such formula are
+\begin{align*}
+\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume},
+\\ \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment~\cite[p. 462]{SPLAG}}, 
+\\ \calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}.
+%\\ \calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
+%\\ \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
+%\\ \calM_n(5) &= \frac{3577728 n - 1825648 n^2 + 2410804 n^3 + 1569392 n^4 + 1644423 n^5 + 906105 n^6 + 341550 n^7 + 75526 n^8 + 8855 n^9 + 385 n^{10}}{95800320 (1 + n)^{9/2}}.
+\end{align*}
 We have explicitly tabulated these formula for $m=0$ to $40$.
@@
 
-We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3, A_4, A_5$ and $A_8$.
-For these plots the SNR is related to noise variance according to~\cite{Viterbo_diamond_cutting_1996}
+We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3, A_4, A_5$ and $A_8$.  For these plots the SNR is related to noise variance according to
 \[
@@
 \]
-where $V$ is the volume of the Voronoi cell and $n$ is the dimension of the lattice.
+where $V$ is the volume of the Voronoi cell and $n$ is the dimension of the lattice~\cite[p.
+167]{Viterbo_diamond_cutting_1996}.
 Figure~\ref{fig:peplots} shows the `exact' probability of error (correct to 16 decimal places) computed using the moments $\calM_n(m)$ and~\eqref{eq:summomentproberror} (solid line).
@@
 \small
-\bibliography{bib}
+\bibliography{../../bib}
 

### 2012-02-21 paper.tex
modified some srcipts
added 3page summary of this paper.

@@
 %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal, onecolumn, 10pt]{IEEEtran}
+\documentclass[journal, onecolumn, 11pt]{IEEEtran}
 %\documentclass[journal]{IEEEtran}
@@
 \begin{abstract}
-We consider the root lattice $A_n$ and derive explicit formulae for
-the moments of its Voronoi cell.  We then show that these formulae
-enable accurate prediction of the error probability of
-lattice codes constructed from $A_n$.
+We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.  These formulae
+enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
 \end{abstract}

### 2012-02-21 build.sh

@@ -1,13 +1,8 @@
-cd plots
-mpost -interaction=nonstopmode peplots.mp
-cd ..
+cd plots 
+mpost -interaction=nonstopmode peplots.mp 
+cd .. 
 
 pdflatex paper.tex
 bibtex paper
 pdflatex paper.tex
 pdflatex paper.tex
-
-pdflatex threepager.tex
-bibtex threepager
-pdflatex threepager.tex
-pdflatex threepager.tex

### 2012-02-21 other/threepager/build.sh

@@ -1,0 +1,10 @@
+#build figures
+mpost -interaction=nonstopmode ../../plots/thpplot.mp 
+ruby1.9.1 hexlatfig.rb
+
+#build tex
+pdflatex threepager.tex
+bibtex threepager
+pdflatex threepager.tex
+pdflatex threepager.tex
+

### 2012-03-07 paper.tex
Added beginning of reviewer comments.

@@
 %\documentclass[a4paper,10pt]{article}
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal, onecolumn, 11pt]{IEEEtran}
+\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+%\documentclass[journal, onecolumn, 11pt]{IEEEtran}
 %\documentclass[journal]{IEEEtran}
@@
 \begin{abstract}
-We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.  These formulae
+We consider the root lattice $A_n$ and derive explicit recursive formulae for the moments of its Voronoi cell.  These formulae
 enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
@@
 Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.
-An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $(n+1)$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of a $(n+1)$-dimensional hypercube onto a hyperplace orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
@@
  \]
- where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
-In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
+ where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called a \term{generator matrix} or \term{basis matrix} or simply generator or basis.
+In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as a generator) and we call this the~\emph{integer lattice}.
 A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional.
@@
 Equivalently it returns the lattice point $\xbf$ such that the translated Voronoi cell $\vor(\Lambda) + \xbf$ contains $\ybf$.
-Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001, Jalden2005_sphere_decoding_complexity}.
+Computationally lattice decoding is known to be NP-hard under certain conditions when the lattice itself, or rather a basis thereof, is considered as an additional input parameter~\cite{micciancio_hardness_2001}.
 Nevertheless, algorithms exist that can compute the nearest lattice point in reasonable time if the dimension is small (approximately $n \leq 60$).
-One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002}.
+One such algorithm is the \term{sphere decoder}~\cite{Viterbo_sphere_decoder_1999,Pohst_sphere_decoder_1981,Agrell2002,Jalden2005_sphere_decoding_complexity}.
  %Kannan~\cite{Kannan1987_fast_general_np} suggested a different approach that is known to be asymptotically faster than the sphere decoder.
@@
 \begin{align}
-P_C &\geq P_C(\Lambda) = \lim_{\abs{\calC}\rightarrow \infty}
- \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in
-\vor(\Lambda) + \cbf )  \nonumber \\
+%P_C &\geq P_C(\Lambda) = \lim_{\abs{\calC}\rightarrow \infty} \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
+P_C &\geq P_C(\Lambda) = \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
 &= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
@@
 \newcommand{\calM}{M}
-In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the $m$th \emph{moment}s
+In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the \emph{moment}s
 \[
@@
 The $n$-volume of $\vor(A_n)$ is $\sqrt{n + 1}$.
-From Berger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
+From Burger~et.~al.~\cite[Theorem 1.1]{Burger1996} we find that the $n$-volume of the projected hypercube $\Qbf\vor(\ints^{n+1})$ is equal to
 \[
@@
 The Voronoi cell is a subset of the zero-sum plane that has zero $(n+1)$-dimensional volume.
-So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.
+So, the volume integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.
 This is not what we intend.
@@
 Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
+If $S$ has $n$-volume $V$ then the $(n+1)$-volume of $S\times\onebf$ is $V \sqrt{n+1}$.
 Now, the integral over the Voronoi cell can be written as
 \begin{equation}\label{eq:vorfint}
-\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
+\frac{1}{\sqrt{n+1}} \int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \frac{1}{\sqrt{n+1}} \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
 \end{equation}
@@
 Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.
-The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by
+The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of the set of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is
 \[
@@
 The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
-\[
-\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
-\]
+\begin{equation}\label{eq:vorfintafternorm}
+\frac{1}{\sqrt{n+1}} \int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
+\end{equation}
 The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.
@@
 
-We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables,
+We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables, gives
 \[
@@
 M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
- &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy. \\
-&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1}.
+ &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy \\
+&= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1},
 \end{align*}
@@
 \|\Qbf\wbf\|^2 &= \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2 \\
-&= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t + \sum_{i=1}^{n+1}w_i \right)^2 \\
-&= t^2 + \sum_{i=1}^{n+1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n+1}w_i + \left(\sum_{i=1}^{n+1}w_i \right)^2 \right)\\
+&= t^2 + \sum_{i=1}^{n-1}w_i^2 - \frac{1}{n+1}\left( t + \sum_{i=1}^{n-1}w_i \right)^2 \\
+&= t^2 + \sum_{i=1}^{n-1}w_i^2 - \frac{1}{n+1}\left( t^2 + 2 t \sum_{i=1}^{n-1}w_i + \left(\sum_{i=1}^{n-1}w_i \right)^2 \right)\\
 &= C + D,
@@
 448]{SPLAG}.
-The simulations are iterated until 500 error events occur.
+The simulations are iterated until 5000 error events occur.
 

### 2012-03-08 paper.tex
more additions of reviewer comments.

@@
 
-\begin{keywords}
+\begin{IEEEkeywords}
 Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
-\end{keywords}
+\end{IEEEkeywords}
 
@@
 Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.
-An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of a $(n+1)$-dimensional hypercube onto a hyperplace orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of an $(n+1)$-dimensional hypercube onto a hyperplace orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
@@
 If $\Bbf$ has more rows than columns, i.e.
-$m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
+$m > n$, then the lattice points lie in an $n$-dimensional subspace of $\reals^m$.
 For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix.
@@
 Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.
-We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.
+We will not specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.
 
@@
 \end{theorem}
-\begin{proof}
+\begin{IEEEproof}
 Let $\ebf_i$ denote a vector with $i$th element equal to one and the remaining elements zero.
@@
 It follows from Lemma~\ref{lem:QVorZnsubsetVorAn} that $\Qbf\vor(\ints^{n+1}) \subseteq \vor(A_n)$, so, because the volumes are the same, and because $\vor(A_n)$ and $\Qbf\vor(\ints^{n+1})$ are polytopes, we have $\vor(A_n) = \Qbf\vor(\ints^{n+1})$.
-\end{proof}
+\end{IEEEproof}
 
@@
 M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
- &= n\sqrt{n+1} \int^{y}_{0} \int_{0}^{1}  \frac{t^{n-1}}{1 - t} \, dt  \, dy \\
+ &= n\sqrt{n+1} \int_{0}^{1} \int^{y}_{0} \frac{t^{n-1}}{1 - t} \, dt  \, dy \\
 &= n\sqrt{n+1}  \int^{1}_{0} \beta(y, n, 0) \, dy = \sqrt{n+1},

### 2012-03-08 paper.tex
added cuberothex figure.

@@
 \end{IEEEproof}
+
+\begin{figure*}[tp]
+	\centering
+		\includegraphics{figs/cuberothex-1.mps}
+		\caption{Orthogonal projection of a cube as it is rotated about its center.
+The figure of the left is the view of a cube from side on, and the boundary is a square.
+When the cube is rotated the boundary becomes a hexagon.
+The hexagon is regular when the cube is viewed along one of its vertices (the rightmost figure).}
+		\label{fig:peplots}
+\end{figure*}
+
+This theorem asserts that the Voronoi cell of the lattice $A_n$ is the $n$-dimensional polytope that results from orthogonally projecting the $(n+1)$-dimensional into the zero sum plane.
+Results of this type have been studied for some time.
+A polytope that is the projection of a hypercube is called a \emph{zonotope}.
 

### 2012-03-08 figs/cuberothex.rb

@@ -1,0 +1,149 @@
+require 'rubygems'
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('cuberothex')
+
+def dround(x)
+  (2056*x).round/2056
+end
+
+scaler = 1.7
+
+#the vertices of a cube
+v1 = Vector[1,1,1]/2.0
+v2 = Vector[1,1,-1]/2.0
+v3 = Vector[1,-1,1]/2.0
+v4 = Vector[1,-1,-1]/2.0
+v5 = Vector[-1,1,1]/2.0
+v6 = Vector[-1,1,-1]/2.0
+v7 = Vector[-1,-1,1]/2.0
+v8 = Vector[-1,-1,-1]/2.0
+
+#put the vertices in a array
+vertices = [v1,v2,v3,v4,v5,v6,v7,v8]
+
+#collect hypercube graph edges into an array
+edges = Array.new
+(0..7).each do |i|
+  ((i+1)..7).each do |j|
+    if (vertices[i] - vertices[j]).r <= 1.01  
+      edges.push([vertices[i],vertices[j]]) 
+    end
+  end
+end
+
+picsideon = Picture.new
+edges.each do |e|
+  vi = e[0]; vj = e[1];
+  path = Path.new
+  path.add_pair(Pair.new(vi[0].cm,vi[1].cm))
+  path.add_pair(Pair.new(vj[0].cm,vj[1].cm))
+  picsideon.add_drawable(Draw.new(path))
+end
+
+picvertexon = Picture.new
+oc = Vector[-1/2.0,-1/2.0,-1/2.0]; #the occluded vertex
+r1 = Vector[1.0,1.0,1.0]; r1 = r1/r1.r #setup for a householder reflection
+r2 = Vector[0.0,0.0,1.0]; r2 = r2/r2.r
+rv = r1 + r2
+nrv = rv/rv.r;
+rm = Matrix[[nrv[0],nrv[1],nrv[2]]]
+R = Matrix.identity(3) - (rm.t*rm)*2 
+edges.each do |e|
+  vi = R*e[0]; vj = R*e[1];
+  path = Path.new
+  path.add_pair(Pair.new(vi[0].cm,vi[1].cm))
+  path.add_pair(Pair.new(vj[0].cm,vj[1].cm))
+  if  (e[0] + oc).r <= 0.01 || (e[1] + oc).r <= 0.01
+    print('true')
[…]

### 2012-03-08 paper.tex
Added more to review

@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal, onecolumn, 11pt]{IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal, onecolumn, 11pt]{IEEEtran}
 %\documentclass[journal]{IEEEtran}
@@
 	\centering
-		\includegraphics{figs/cuberothex-1.mps}
+		\includegraphics[scale=1.6]{figs/cuberothex-1.mps}
 		\caption{Orthogonal projection of a cube as it is rotated about its center.
@@
 When the cube is rotated the boundary becomes a hexagon.
-The hexagon is regular when the cube is viewed along one of its vertices (the rightmost figure).}
-		\label{fig:peplots}
+The hexagon is regular when the cube is viewed along one of its vertices (the rightmost figure).
+The regular hexagon is the Voronoi cell of the hexagonal lattice $A_2$.
+Every 2-dimensional zonotope is also unimodular so the boundary of each of the figures above can be used to tile Euclidean space~\cite{McMullen_space_tiling_zonotopes_1975}.}
+		\label{fig:cuberothex}
 \end{figure*}
@@
 This theorem asserts that the Voronoi cell of the lattice $A_n$ is the $n$-dimensional polytope that results from orthogonally projecting the $(n+1)$-dimensional into the zero sum plane.
-Results of this type have been studied for some time.
-A polytope that is the projection of a hypercube is called a \emph{zonotope}.
+Results of this type have been studied previously.
+A polytope that is the orthogonal projection of a hypercube is called a \emph{zonotope}~\cite[p.~313]{surveys_combin_zonotopes_1997}~\cite{McMullen_space_tiling_zonotopes_1975}.
+Figure~\ref{fig:cuberothex} depicts some 2-dimensional zonotopes.
+A zonotope is \emph{unimodular} if it can be used to tiles (or tesselate) Euclidean space.
+Such a tesselation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
+The lattice $A_n$ is a zonotopol lattice.
+The techniques in this paper can potentially be extended to other zonotopal lattices, but we do not consider such extentions here.
 

### 2012-03-08 figs/cuberothex.rb

@@ -109,29 +109,7 @@ edges.each do |e|
   end
 end
 
-#R = Matrix[[Math.cos(Math::PI/3), -Math.sin(Math::PI/3)],[Math.sin(Math::PI/3), Math.cos(Math::PI/3)]]
-#v = (M*Vector[1,0] + M*Vector[0,1])*(1/3)
-#vorpath = Path.new
-#6.times do
-#  vorpath.add_pair(Pair.new(dround(v[0]).cm, dround(v[1]).cm))
-#  v = R*v
-#end
-#vorpath.add_pair('cycle')
-
-#pic = Picture.new('picL')
-#circ = Picture.new('circ')
-#vorpic = Picture.new('vor')
-#inrad = 1.0
-#(-3..3).each do |x|
-#	(-3..3).each do |y|
-#		v = M*Vector[x,y]
-#		pic.add_drawable(Fill.new(Circle.new()).scale(0.1.cm).translate(v[0].cm,v[1].cm))
-#    circ.add_drawable(Draw.new(Circle.new()).scale((scaler*inrad).cm).translate(v[0].cm,v[1].cm))
-#    vorpic.add_drawable(Draw.new(vorpath).translate(v[0].cm,v[1].cm))
-#	end
-#end
-
-#figure 2 is the same but with packing circles added
+#figure 1 is 4 projected cubes, all unimodular zonotopes
 fig = Figure.new
 fig.add_drawable(Draw.new(picsideon))
 fig.add_drawable(Draw.new(picmiddle1).translate(2.0.cm,0.0.cm))
@@ -144,6 +122,31 @@ fig.add_drawable(Draw.new(picvertexon).translate(6.0.cm,0.0.cm))
 #fig.add_drawable(Draw.new(vorpic))
 #fig.add_drawable(Draw.new(pic))
 #fig.add_drawable(Draw.new(circ))
+file.add_figure(fig)
+
+
+picother = Picture.new
+#oc = Vector[1/2.0,1/2.0,-1/2.0]; #the occluded vertex
+# a random orthogonal matrix
+R = Matrix[[-0.1817, 0.7236, 0.6658],
+           [-0.6198, -0.6100, 0.4938],
+           [0.7634, -0.3230, 0.5594]] 
+edges.each do |e|
+  vi = R*e[0]; vj = R*e[1];
+  path = Path.new
+  path.add_pair(Pair.new(vi[0].cm,vi[1].cm))
+  path.add_pair(Pair.new(vj[0].cm,vj[1].cm))
[…]

### 2012-03-08 other/threepager/build.sh

@@ -1,10 +1,10 @@
 #build figures
 mpost -interaction=nonstopmode ../../plots/thpplot.mp 
-ruby1.9.1 hexlatfig.rb
+ruby1.9.1 hexlatfig.rb 
 
-#build tex
-pdflatex threepager.tex
-bibtex threepager
-pdflatex threepager.tex
-pdflatex threepager.tex
+#build tex 
+pdflatex threepager.tex 
+bibtex threepager 
+pdflatex threepager.tex 
+pdflatex threepager.tex 
 

### 2012-03-08 paper.tex
more additions

@@
 Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.
-An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of an $(n+1)$-dimensional hypercube onto a hyperplace orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
+An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the orthogonal projection of an $(n+1)$-dimensional hypercube onto a hyperplane orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
  %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
@@
 \newcommand{\calC}{\mathcal C}
-Lattices can be used to construct \emph{signal space codes}.
-A signal space code $\calC$ of dimension $n$ is a finite set of vectors (or points) in the Euclidean space $\reals^n$.
-Each vector in $\calC$ is called a \emph{codeword} and represents a particular signal.
+Lattices can be used to construct \emph{lattice codes}.
+A lattice code $\calC$ of dimension $n$ is a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
+Each point in $\calC$ is called a \emph{codeword} and represents a particular signal.
+There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$.
+The codewords are given by those lattice points inside the shaping region, that is, $\calC = S \cap \Lambda.$  Common choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes, or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
 The number of codewords is denoted by $\abs{\calC}$.
-If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2\abs{\calC}$ bits per codeword.
-The average power of the code is given by $P = \frac{1}{\abs{\calC}}\sum_{\cbf \in \calC}\|\cbf\|^2$.
+If each codeword is transmitted with equal probability then the rate of the code is $R = \frac{1}{n}\log_2{\abs{\calC}}$ bits per codeword.
+The average power of the code is $P = \frac{1}{\abs{\calC}}\sum_{\cbf \in \calC}\|\cbf\|^2$.
+
 In the AWGN channel the received signal takes the form
@@
 \begin{equation}\label{eq:mldecoder}
-\hat{\cbf} = \underset{\cbf \in \calC}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2.
+\hat{\cbf}_{\text{ML}} = \underset{\cbf \in C}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
 %\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
 \end{equation}
-That is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
-Assuming that each codeword is transmitted with equal probability then the probability of correct decoding is
+that is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
+Assuming that each codeword is transmitted with equal probability, then probability of correct maximum likelihood decoding is
 \[
-P_C =\frac{1}{\abs{\calC}}\sum_{\cbf \in \calC} \prob( \hat{\cbf} = \cbf ),
+\sum_{\cbf \in \calC}\prob( \hat{\cbf}_{\text{ML}} = \cbf ).
 \]
-and the probability of error is
-\[
-P_E = 1 - \frac{1}{\abs{\calC}} \sum_{\cbf \in \calC} \prob(
-\hat{\cbf} = \cbf ).
-\]
-%The maximum possible rate of reliable communications in this channel is given by the channel capacity $B\log_2\left(1 + \frac{P}{\sigma^2}\right)$.
-If the codewords are chosen uniformly randomly on a sphere of radius $\sqrt{P}$ then as $n$ gets large the rate of the code can be made arbitrarily close to the channel capacity while simultaneously the probability of error can be made arbitrarily small.
-Another way of stating this is that we can choose close to $2^{nR/B}$ codewords uniformly randomly on a sphere of radius $\sqrt{P}$ and as $n$ gets large the probability of error decreases to zero.
+Typically maximum likelihood decoding is computationally complex and it is preferable to use lattice decoding~\cite{Agrell2002,Erex2004_lattice_decoding}.  The estimator of $\cbf$ is then,
+\begin{equation}\label{eq:latticedecoder}
+\hat{\cbf} = \underset{\cbf \in \Lambda}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
+%\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
+\end{equation}
+that is, the receiver computes the lattice point in $\Lambda$ nearest in Euclidean distance to the received signal $\ybf$.
+Equivalently, $\hat{\cbf}$ is the lattice point such that $\ybf \in \vor(\Lambda) + \hat{\cbf}$.
+Note that with lattice decoding the decoded lattice point $\hat{\cbf}$ is not guaranteed to be inside the code $\calC$.
 
-%A problem with randomly generated codes and maximum likelihood decoding is that there are no computationally efficient methods for encoding and decoding.
-For encoding the transmitter must store a list of size $\abs{\calC} \approx 2^{nR/B}$ and for decoding the receiver must compute $\hat{\cbf}$ requiring a search over all $\abs{\calC} \approx 2^{nR/B}$ codewords.
-As $n$ gets large this becomes infeasible.
-To avoid this computational complexity it is preferable for a code to have some type of `structure' that is exploitable for efficient encoding and decoding.
-One way of enforcing structure is to choose codewords from a lattice.
-The idea being that encoding and decoding can now be performed efficiently by \emph{lattice decoding}, i.e.
-by finding a nearest point in a lattice.
+Correct lattice decoding occurs when $\hat{\cbf} = \cbf$, or equivalently when $\ybf \in \vor(\Lambda) + \cbf$, or equivalently when $\wbf \in \vor(\Lambda)$, i.e.
+when the noise $\wbf$ is inside the Voronoi cell of the lattice.
+Assuming that each codeword is transmitted with equal probability then the probability of correct lattice decoding is
+\begin{align}
+P_C &= \sum_{\cbf \in \calC} \prob( \hat{\cbf} = \cbf ) \\
+&= \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
+&= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
+&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
+\end{align}
+The probability of error is $P_E = 1 - P_C$.
+The probability of correct lattice decoding is smaller than the probability of correct maximum likelihood decoding, however, as the size of the code $\abs{\calC}$ increases, $P_C$ converges to the probability of correct maximum likelihood decoding (the proportion of codewords near the boundary of the shaping region become small).
 
-A \emph{lattice code} is a signal space code with codewords taken from a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
-There are infinitely many ways to choose a finite subset from a lattice, but common approaches make use of a bounded subset of $\reals^n$, called a \emph{shaping region} $S \subset \reals^n$.
-The codewords are given by those lattice points inside the shaping region, that is,
-\[
-\calC = S \cap \Lambda.
-\]
-Common choices of shaping region are $n$-dimensional spheres, spherical shells, hypercubes, or the Voronoi cell of a \emph{sublattice} of $\Lambda$~\cite{Buda1989_some_opt_codes_structure,Erex2004_lattice_decoding,Conway1983VoronoiCodes}.
-A consequence of the finiteness of a lattice code is that the maximum likelihood decoder~\eqref{eq:mldecoder} is \emph{not} equivalent to lattice decoding.
-Computing a nearest lattice point will not in general return a lattice point from the code (the decoded lattice point might lie outside the shaping region).
- %A question that remained open for some time was whether lattice codes employing lattice decoding could achieve channel capacity.
-This question was answered in the affirmative by Erez and Zamir~\cite{Erex2004_lattice_decoding}.
-
-Let $\calC$ be a lattice code constructed from a lattice $\Lambda$.
-Assuming that each codeword $\cbf \in \calC$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
-%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).
-A result of this is that that observed noise distribution is no longer Gaussian.
-However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.
-In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
-\begin{align}
-%P_C &\geq P_C(\Lambda) = \lim_{\abs{\calC}\rightarrow \infty} \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
-P_C &\geq P_C(\Lambda) = \frac{1}{\abs{\calC}} \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
-&= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
-&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
-e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
-\end{align}
-The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).
 By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
 \begin{align}
-P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
+P_C  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
 - \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
@@
 So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.
-The number of terms required will increase as the noise variance gets smaller.
+The number of terms required increases as the noise variance gets smaller.
 This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
+We call these the \emph{moments} of $\vor(\Lambda)$.
 
 \newcommand{\calM}{M}
-In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and  we will derive expressions for the \emph{moment}s
+In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and we derive expressions for the moments
 \[
@@
 The regular hexagon is the Voronoi cell of the hexagonal lattice $A_2$.
-Every 2-dimensional zonotope is also unimodular so the boundary of each of the figures above can be used to tile Euclidean space~\cite{McMullen_space_tiling_zonotopes_1975}.}
+Every 2-dimensional zonotope is also unimodular so the boundary of each of the figures above can be used to tile 2-dimensional Euclidean space~\cite{McMullen_space_tiling_zonotopes_1975}.}
 		\label{fig:cuberothex}

### 2012-03-08 build.sh

@@ -2,6 +2,10 @@ cd plots
 mpost -interaction=nonstopmode peplots.mp 
 cd .. 
 
+cd figs 
+mpost -interaction=nonstopmode cuberothex.mp 
+cd .. 
+
 pdflatex paper.tex
 bibtex paper
 pdflatex paper.tex

### 2012-03-08 figs/hexlatfig.rb
minor edits

@@ -1,0 +1,51 @@
+require 'rubygems'
+require 'rubypost'
+
+require 'matrix'
+require 'mathn'
+
+include RubyPost
+
+file = RubyPost::File.new('hexlatfig')
+
+def dround(x)
+  (2056*x).round/2056
+end
+
+scaler = 1.7
+
+M = scaler*Matrix[ [1,0.5], [0, 0.5*(3**(0.5))] ]
+R = Matrix[[Math.cos(Math::PI/3), -Math.sin(Math::PI/3)],[Math.sin(Math::PI/3), Math.cos(Math::PI/3)]]
+v = (M*Vector[1,0] + M*Vector[0,1])*(1/3)
+vorpath = Path.new
+6.times do
+  vorpath.add_pair(Pair.new(dround(v[0]).cm, dround(v[1]).cm))
+  v = R*v
+end
+vorpath.add_pair('cycle')
+
+pic = Picture.new('picL')
+circ = Picture.new('circ')
+vorpic = Picture.new('vor')
+inrad = 1.0
+(-3..3).each do |x|
+	(-3..3).each do |y|
+		v = M*Vector[x,y]
+		pic.add_drawable(Fill.new(Circle.new()).scale(0.1.cm).translate(v[0].cm,v[1].cm))
+    circ.add_drawable(Draw.new(Circle.new()).scale((scaler*inrad).cm).translate(v[0].cm,v[1].cm))
+    vorpic.add_drawable(Draw.new(vorpath).translate(v[0].cm,v[1].cm))
+	end
+end
+
+#figure 2 is the same but with packing circles added
+fig = Figure.new
+cliprect = "((3.5cm,2.5cm)--(-3.5cm,2.5cm)--(-3.5cm,-2.5cm)--(3.5cm,-2.5cm)--cycle)"
+fig.add_drawable(Clip.new(cliprect, pic))
+fig.add_drawable(Clip.new(cliprect, vorpic))
+fig.add_drawable(Clip.new(cliprect, circ))
+fig.add_drawable(Draw.new(vorpic))
+fig.add_drawable(Draw.new(pic))
+fig.add_drawable(Draw.new(circ))
+file.add_figure(fig)
+
+file.compile

### 2012-03-08 paper.tex
minor edits
Added to reviews.

@@
 The probability of error is $P_E = 1 - P_C$.
-The probability of correct lattice decoding is smaller than the probability of correct maximum likelihood decoding, however, as the size of the code $\abs{\calC}$ increases, $P_C$ converges to the probability of correct maximum likelihood decoding (the proportion of codewords near the boundary of the shaping region become small).
+The probability of correct lattice decoding is smaller than the probability of correct maximum likelihood decoding, however, as the size of the code $\abs{\calC}$ increases, $P_C$ converges to the probability of correct maximum likelihood decoding (the proportion of codewords near the boundary of the shaping region becomes small).
 
@@
 108]{SPLAG}.
+In two dimensions $A_2$ is the hexagonal lattice (Figure~\ref{fig:hexlatfig}), and in three dimension $A_3$ is the body centered cubic lattice~\cite[p.
+108]{SPLAG}.
+These are the densest known sphere packings in dimensions two and three.
+
+\begin{figure}[tbp]
+	\centering
+		\includegraphics{figs/hexlatfig-1.mps}
+		\caption{The hexagonal lattice $A_2$ with its sphere packing and Voronoi cell.  The Voronoi cell is a regular hexagon.}
+		\label{fig:hexlatfig}
+\end{figure}
 
@@
 
-This theorem asserts that the Voronoi cell of the lattice $A_n$ is the $n$-dimensional polytope that results from orthogonally projecting the $(n+1)$-dimensional into the zero sum plane.
+This theorem asserts that the Voronoi cell of the lattice $A_n$ is the $n$-dimensional polytope that results from orthogonally projecting the $(n+1)$-dimensional hypercube into the zero-sum plane.
 Results of this type have been studied previously.
@@
 Figure~\ref{fig:cuberothex} depicts some 2-dimensional zonotopes.
-A zonotope is \emph{unimodular} if it can be used to tiles (or tesselate) Euclidean space.
+A zonotope is \emph{unimodular} if it can be used to tile (or tesselate) Euclidean space.
 Such a tesselation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
@@
 Consider a function $f : \reals^{n+1} \mapsto \reals$.
-The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.
+The definition we have made for $A_n$ in Section~\ref{sec:lattice-a_n} places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.
 The Voronoi cell is a subset of the zero-sum plane that has zero $(n+1)$-dimensional volume.
@@
 However, we find the following approach simpler.
-Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
-If $S$ has $n$-volume $V$ then the $(n+1)$-volume of $S\times\onebf$ is $V \sqrt{n+1}$.
+Given a set $S$ of vectors from the zero-sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
+If $S$ has $n$-volume equal to $V$ then the $(n+1)$-volume of $S\times\onebf$ is $V\|\onebf\| = V \sqrt{n+1}$.
 Now, the integral over the Voronoi cell can be written as

### 2012-03-09 paper.tex
Added more to paper revision.  Added code to test the numerator/denominator conjection about the moments as n gets large.

@@
 
-The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
+The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the additive white Gaussian noise (AWGN) channel~\cite{SPLAG}.
 The highly symmetric structure of these lattices provides the grounds for extremely efficient encoding and decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.
@@
 \begin{equation}\label{eq:mldecoder}
-\hat{\cbf}_{\text{ML}} = \underset{\cbf \in C}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
+\hat{\cbf}_{\text{ML}} = \underset{\cbf \in \calC}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
 %\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
@@
 that is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
-Assuming that each codeword is transmitted with equal probability, then probability of correct maximum likelihood decoding is
+Assuming that each codeword is transmitted with equal probability, then the probability of correct maximum likelihood decoding is
 \[
-\sum_{\cbf \in \calC}\prob( \hat{\cbf}_{\text{ML}} = \cbf ).
+P_{\text{ML}} = \sum_{\cbf \in \calC}\frac{1}{\abs{\calC}}\prob( \hat{\cbf}_{\text{ML}} = \cbf ).
 \]
-Typically maximum likelihood decoding is computationally complex and it is preferable to use lattice decoding~\cite{Agrell2002,Erex2004_lattice_decoding}.  The estimator of $\cbf$ is then,
+
+Maximum likelihood decoding is typically computationally complex and it is preferable to use lattice decoding~\cite{Agrell2002,Erex2004_lattice_decoding}.  The estimator of $\cbf$ is then,
 \begin{equation}\label{eq:latticedecoder}
@@
 \begin{align}
-P_C &= \sum_{\cbf \in \calC} \prob( \hat{\cbf} = \cbf ) \\
-&= \sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
+P_C &= \frac{1}{\abs{\calC}}\sum_{\cbf \in \calC} \prob( \hat{\cbf} = \cbf ) \nonumber \\
+&= \frac{1}{\abs{\calC}}\sum_{\cbf\in \calC} \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
 &= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
@@
 The probability of error is $P_E = 1 - P_C$.
-The probability of correct lattice decoding is smaller than the probability of correct maximum likelihood decoding, however, as the size of the code $\abs{\calC}$ increases, $P_C$ converges to the probability of correct maximum likelihood decoding (the proportion of codewords near the boundary of the shaping region becomes small).
+The probability of correct lattice decoding is smaller than the probability of correct maximum likelihood decoding.
+However, as the size of the code $\abs{\calC}$ increases, the proportion of codewords near the boundary of the shaping region becomes small, and $P_C$ converges to $P_{\text{ML}}$.
 
@@
 This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
-We call these the \emph{moments} of $\vor(\Lambda)$.
+We call these terms the \emph{moments} of $\vor(\Lambda)$.
 
@@
 \]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$ (see Appendix~\ref{sec:solv-this-recurs}).  The first five such formula are:
 \begin{align*}
@@
 108]{SPLAG}.
-These are the densest known sphere packings in dimensions two and three.
+%These are the densest known sphere packings in dimensions two and three.
 
@@
 	\centering
-		\includegraphics{figs/hexlatfig-1.mps}
+		\includegraphics[scale=1.2]{figs/hexlatfig-1.mps}
 		\caption{The hexagonal lattice $A_2$ with its sphere packing and Voronoi cell.  The Voronoi cell is a regular hexagon.}
@@
 		\caption{Orthogonal projection of a cube as it is rotated about its center.
-The figure of the left is the view of a cube from side on, and the boundary is a square.
+The figure on the left is the view of a cube from side on, and the boundary is a square.
 When the cube is rotated the boundary becomes a hexagon.
 The hexagon is regular when the cube is viewed along one of its vertices (the rightmost figure).
-The regular hexagon is the Voronoi cell of the hexagonal lattice $A_2$.
+The regular hexagon is the Voronoi cell of the hexagonal lattice $A_2$ (see~Figure~\ref{fig:hexlatfig}).
 Every 2-dimensional zonotope is also unimodular so the boundary of each of the figures above can be used to tile 2-dimensional Euclidean space~\cite{McMullen_space_tiling_zonotopes_1975}.}
@@
 Figure~\ref{fig:cuberothex} depicts some 2-dimensional zonotopes.
-A zonotope is \emph{unimodular} if it can be used to tile (or tesselate) Euclidean space.
-Such a tesselation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
-The lattice $A_n$ is a zonotopol lattice.
-The techniques in this paper can potentially be extended to other zonotopal lattices, but we do not consider such extentions here.
+A zonotope is \emph{unimodular} if it can be used to tile (or tessellate) Euclidean space.
+Such a tessellation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
+The lattice $A_n$ is a zonotopal lattice.
+The techniques in this paper can potentially be extended to other zonotopal lattices, but we do not consider such extensions here.
 
@@
 Given a set $S$ of vectors from the zero-sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.
-If $S$ has $n$-volume equal to $V$ then the $(n+1)$-volume of $S\times\onebf$ is $V\|\onebf\| = V \sqrt{n+1}$.
+If $S$ has $n$-volume equal to $V$ then the $(n+1)$-volume of $S\times\onebf$ is equal to $V\|\onebf\| = V \sqrt{n+1}$.
 Now, the integral over the Voronoi cell can be written as
@@
 \end{equation}
-The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.
+The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.\footnote{A caveat applies when $x_{\text{max}} = \nicefrac{1}{2}$ and $x_{\text{min}} = -\nicefrac{1}{2}$ and the denominator in the integral in~\eqref{eq:vorfintafternorm} is equal to zero.
+In this case the interval $[-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$ is empty and we specify that these points do not contribute to the integral.}
 
@@
 These enable accurate prediction of the performance of codes constructed from $A_n$.
-In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
-Our results include codes constructed using these packings as a special case.
-%The lattice $A_n$ does not produce good codes in large dimension, but we are hopeful that the techniques developed here can be applied to other lattices.
-In particular similar techniques might apply to the family Coxeter lattices, which contains the root lattice $E_8$.
+The formulae were obtained by observing that the Voronoi cell of $A_n$ is a \emph{zonotope}, i.e.
+it can be described as an orthogonal projection of the $(n+1)$-dimensional hypercube.
+It is possible that the techniques developed here can be applied to other \emph{zonotopal lattices}~\cite{Vallentin_space_tiling_zonotope_2004}, i.e.
+those lattices with Voronoi cells that are zonotopes.
 
@@
 Mathematica 8.0 was used to perform these calculations.
+We have computed formula for $m = 0, 1, \dots, 40$ this way, but it becomes computationally infeasible for large $m$.
+
+We have observed the following property.  The formula for $M_n(m)$ appears to always be in the form
+\[
+\frac{p(n)}{(1+n)^{m-1/2} d}
+\] 
+where $p(n) = a_0 + a_1n + \dots a_{2m} n^{2m}$ is a polynomial of order $2m$ and $d \in \ints$ is the denominator.  The following appears to be true
+\[
+\frac{d}{a_{2m}} = 12^m,
+\]
+that is, the integer on the denominator $d$ is equal to the highest order coefficient $a_{2m}$ of the numerator polynomial $p(n)$ multiplied by $12^m$.
+This leads us to make the following conjecture about the moments of $A_n$ when the dimension $n$ gets large.
+
+\begin{conjecture}
+For any fixed $m$,
+\[
+\frac{M_n(m)}{n^{m+1/2}} \rightarrow \frac{1}{12^{m}} 
+\]
+as $n \rightarrow \infty$.
+\end{conjecture}
+
+It might be hoped that this conjecture leads to an asymptotic result about the probability of error of lattice codes constructed from $A_n$ when $n$ is large.
+However, this does not appear to be the case when the variance of the noise $\sigma^2$ is less than one.
+In this case, the term $\sigma^n$ on the denominator of~(\ref{eq:summomentproberror}) shrinks with $n$.
+So, when $\sigma < 1$, the number of moments required to produce an accurate estimate of the probability of error increases with $n$.
 

### 2012-03-09 build.sh

@@ -4,6 +4,7 @@ cd ..
 
 cd figs 
 mpost -interaction=nonstopmode cuberothex.mp 
+mpost -interaction=nonstopmode hexlatfig.mp 
 cd .. 
 
 pdflatex paper.tex

### 2012-03-12 paper.tex
minor edits.
minor edits

@@
 \[
-P_{\text{ML}} = \sum_{\cbf \in \calC}\frac{1}{\abs{\calC}}\prob( \hat{\cbf}_{\text{ML}} = \cbf ).
+P_{\text{ML}} = \frac{1}{\abs{\calC}}\sum_{\cbf \in \calC}\prob( \hat{\cbf}_{\text{ML}} = \cbf ).
 \]
@@
 Figure~\ref{fig:cuberothex} depicts some 2-dimensional zonotopes.
-A zonotope is \emph{unimodular} if it can be used to tile (or tessellate) Euclidean space.
-Such a tessellation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
-The lattice $A_n$ is a zonotopal lattice.
-The techniques in this paper can potentially be extended to other zonotopal lattices, but we do not consider such extensions here.
+A zonotope is \emph{unimodular} if it can be used to tile (or tesselate) Euclidean space.
+Such a tesselation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
+The lattice $A_n$ is a zonotopal lattice and the techniques in this paper can potentially be extended to other zonotopal lattices.
 
@@
 \] 
-where $p(n) = a_0 + a_1n + \dots a_{2m} n^{2m}$ is a polynomial of order $2m$ and $d \in \ints$ is the denominator.  The following appears to be true
+where $p(n) = a_0 + a_1n + \dots a_{2m} n^{2m}$ is a polynomial of degree $2m$ and $d$ is a positive integer.  The following appears to be true
 \[
@@
 \]
-that is, the integer on the denominator $d$ is equal to the highest order coefficient $a_{2m}$ of the numerator polynomial $p(n)$ multiplied by $12^m$.
+that is, the integer in the denominator $d$ is equal to the highest order coefficient $a_{2m}$ of the numerator polynomial $p(n)$ multiplied by $12^m$.
 This leads us to make the following conjecture about the moments of $A_n$ when the dimension $n$ gets large.
@@
 
-% \section{Conclusion}
-% \begin{itemize}
-% \item We can use these techniques to compute other things about the Voronoi cell, i.e. the average $L_p$ norms.
-% \item Not restricted to Gaussian noise.  Any noise pdf that can be described by the moments will work.
-% \item Can other lattices be done?
-% \end{itemize}
-
 \small
@@
 
-% \section{Notable references}
-% \begin{itemize}
-% \item Conway and Sloane compute the second moment of all the root lattices and there duals~\cite{Conway1982VoronoiRegions}.
-This is really \emph{the} paper on this topic.
-The approach is to break the Voronoi cells into simplicies and compute the second moment of the simplicies.
-
-% \begin{quotation}
-% In the second application, the same Euclidean code
-% $y_1,\dots ,y_n$ is used as a code for the Gaussian channel. Now
-% the Voronoi regions are the decoding regions: all points $x$
-% in the interior of $V(y_i)$ are decoded as $y_i$. If the codewords
-% are equally likely and all the Voronoi regions $V(y_i)$ are
-% congruent to a polytope P, the probability of correct decoding is proportional to
-% \[
-% \int_P e^{x^\prime x} dx.
-% \]
-% The description of the Voronoi regions given in Section III
-% thus makes it possible to calculate this probability exactly
-% for many Euclidean codes. These results will be described
-% elsewhere.
-% \end{quotation}
-
-% But it doesn't appear to be published elsewhere.  I'm not really sure how you would go about doing it either, because I think integrating $e^{x^\prime x}$ over an arbitray simplex is hard.
-
-% \item Baldoni \emph{et.
-al.}~\cite{Baldoni_how_to_integrate_poly_over_simplex_2008} claim to show that in general integrating a polynomial under a simplex is NP-hard.
-However, if you fix the degree of the polynomial then it can be done in polynomial time.
-
-% \end{itemize}
-
 \end{document}

### 2012-03-19 other/threepager/threepager.tex
added modified stuff for the threepager.tex

@@
 %\documentclass[journal, onecolumn, 10pt]{IEEEtran}
-\documentclass[journal]{../../IEEEtran}
+%\documentclass[journal]{../../IEEEtran}
+
+\documentclass[final]{proceedings}
 %\documentclass[correspondence,9pt]{IEEEtran}
 
+\usepackage{crop}
+\crop
+%\makeindex
+
 \usepackage{../../mathbf-abbrevs}
-\usepackage{amsmath}
-\DeclareMathOperator{\argmin}{\arg\!\min}
+%\usepackage{amsmath}
+\newcommand{\argmin}{\arg\!\min}
 
@@
 
+%\title{On the error performance of the $A_n$ lattices}
+%\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson
+ %\thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Emanuele Viterbo is %with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
+Vaughan~Clarkson is with the School of Information %Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+A preliminary full version of this paper is available at http://arxiv.org/abs/1111.6334.}
+%}
+
+\begin{document}
+
 %opening
-\title{On the error performance of the $A_n$ lattices}
-\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson
- \thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+\chapter[On the error performance of the $A_n$ lattices]{On the error performance of the $A_n$ lattices}
+
+\begin{authorline}
+Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo and I.
+Vaughan L.
+Clarkson\thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
 Emanuele Viterbo is with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
@@
 A preliminary full version of this paper is available at http://arxiv.org/abs/1111.6334.}
-}
-
-\begin{document}
+\end{authorline}
 
@@
 \newcommand{\hist}{\operatorname{hist}}
-
-\maketitle
-
-\begin{abstract}
-We consider the root lattice $A_n$ and derive explicit formulae for the moments of its Voronoi cell.
-These formulae enable accurate prediction of the error probability of lattice codes constructed from $A_n$.
-\end{abstract}
-
-\begin{keywords}
-Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
-\end{keywords}
-
-\section{Introduction}\label{sec:introduction}
 
@@
 % \]
-In two dimensions $A_2$ is the hexagonal lattice (Figure~\ref{fig:hexlatfig}), and in three dimension $A_3$ is the body centered cubic lattice~\cite[p.
+In two dimensions $A_2$ is the hexagonal lattice, and in three dimension $A_3$ is the body centered cubic lattice~\cite[p.
 108]{SPLAG}.
@@
 
-\begin{figure}[tbp]
-	\centering
-		\includegraphics[width=\linewidth]{hexlatfig-1.mps}
-		\caption{The hexagonal lattice $A_2$.  The Voronoi cell is a regular hexagon.}
-		\label{fig:hexlatfig}
-\end{figure}
+%\begin{figure}[tbp]
+%	\centering
+%		\includegraphics[width=\linewidth]{hexlatfig-1.mps}
+%		\caption{The hexagonal lattice $A_2$.  The Voronoi cell is a regular hexagon.}
+%		\label{fig:hexlatfig}
+%\end{figure}
 
@@
 
-\section{Conclusion}
+%\section{Conclusion}
 
-Recursive formulae for the moments of the Voronoi cell of the lattice $A_n$ were found.
+%Recursive formulae for the moments of the Voronoi cell of the lattice $A_n$ were found.
 These enable accurate prediction of the performance of codes constructed from $A_n$.

### 2012-03-19 other/threepager/build.sh

@@ -3,8 +3,8 @@ mpost -interaction=nonstopmode ../../plots/thpplot.mp
 ruby1.9.1 hexlatfig.rb 
 
 #build tex 
-pdflatex threepager.tex 
+latex threepager.tex 
 bibtex threepager 
-pdflatex threepager.tex 
-pdflatex threepager.tex 
+latex threepager.tex 
+latex threepager.tex 
 

### 2012-03-27 other/threepager/threepager.tex
modified the metapost files so that they (should) work with newer versions of metapost

@@
 
-The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}.
-The highly symmetric structure of these lattices provides the grounds for extremely efficient encoding and decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices}.
-In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.
+The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the  additive white Gaussian noise (AWGN) channel.  %~\cite{SPLAG}. 
+The highly symmetric structure of these lattices provides the grounds for extremely efficient encoding and decoding algorithms.
+In this paper we derive formulae for accurately predicting the performance of lattice codes constructed from $A_n$.
 This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.
-Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices~\cite{Conway1982VoronoiRegions}.
-In this paper we extend their technique to compute the moments of any order for $A_n$.
+%Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices~\cite{Conway1982VoronoiRegions}.
+Here, we extend their technique to compute the moments of any order for $A_n$.
 
-In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
+%In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.
 These are the densest sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.
@@
  where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis.
-In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
+%In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}.
 %A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional.
@@
 $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$.
-
 The \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\reals^m$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point.
 The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.
-If we translate the Voronoi cell by a lattice point $\xbf$, we obtain the subset of $\reals^m$ that is closer to $\xbf$ than any other lattice point.
+%If we translate the Voronoi cell by a lattice point $\xbf$, we obtain the subset of $\reals^m$ that is closer to $\xbf$ than any other lattice point.
 We write $\vor(\Lambda) + \xbf$ to denote the Voronoi cell translated by $\xbf$.
@@
 We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.
-The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, minimal vectors, center density, thickness, and the normalized second moment (or quantizing constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
+%The Voronoi cell encodes many interesting lattice properties such as the packing radius, covering radius, kissing number, minimal vectors, center density, thickness, and the normalized second moment (or quantizing constant)~\cite{Viterbo_diamond_cutting_1996, SPLAG}.
 The error probability of a lattice code can also be evaluated from the Voronoi cell as we will see.
-There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Valentin2003_coverings_tilings_low_dimension}.
+%There exist algorithms to completely enumerate the Voronoi cell of an arbitrary lattice~\cite{Viterbo_diamond_cutting_1996,Sikiric_complex_algs_vor_cells_2009,Valentin2003_coverings_tilings_low_dimension}.
 In general these algorithms are only computationally feasible when the dimension is small (approximately $n \leq 9$).
@@
 \newcommand{\calC}{\mathcal C}
-Lattices can be used to construct \emph{lattice codes}.
+%Lattices can be used to construct \emph{lattice codes}.
 A lattice code $\calC$ of dimension $n$ is a finite subset of points of some lattice $\Lambda$ in $\reals^n$.
@@
 
-In the AWGN channel the received signal takes the form
-\[
-\ybf = \cbf + \wbf
-\]
-where $\ybf \in \reals^n$, $\cbf \in \calC$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
+% In the AWGN channel the received signal takes the form
+% \[
+% \ybf = \cbf + \wbf
+% \]
+% where $\ybf \in \reals^n$, $\cbf \in \calC$ and $\wbf$ is a vector of independent and identically distributed Gaussian random variables with variance $\sigma^2$.
 If the receiver employs maximum likelihood decoding then the estimator of $\cbf$ given $\ybf$ at the receiver is
-\begin{equation}\label{eq:mldecoder}
-\hat{\cbf}_{\text{ML}} = \underset{\cbf \in C}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
-%\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
-\end{equation}
-that is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
+% \begin{equation}\label{eq:mldecoder}
+% \hat{\cbf}_{\text{ML}} = \underset{\cbf \in C}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
+% %\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
+% \end{equation}
+% that is, the receiver computes the codeword in $\calC$ nearest in Euclidean distance to the received signal $\ybf$.
 Typically maximum likelihood decoding is computationally complex and it is preferable to use \emph{lattice decoding}~\cite{Erex2004_lattice_decoding}.
 The estimator of $\cbf$ is then,
-\begin{equation}\label{eq:latticedecoder}
-\hat{\cbf} = \underset{\cbf \in \Lambda}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
-%\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
-\end{equation}
-that is, the receiver computes the lattice point in $\Lambda$ nearest in Euclidean distance to the received signal $\ybf$.
+% \begin{equation}\label{eq:latticedecoder}
+% \hat{\cbf} = \underset{\cbf \in \Lambda}{\operatorname{argmin}}\;\| \ybf - \cbf \|^2,
+% %\hat{\cbf} &=& \mbox{arg~}\min_{\cbf \in \calC} \| \ybf - \cbf \|^2.
+% \end{equation}
+% that is, the receiver computes the lattice point in $\Lambda$ nearest in Euclidean distance to the received signal $\ybf$.
 Equivalently, $\hat{\cbf}$ is the lattice point such that $\ybf \in \vor(\Lambda) + \hat{\cbf}$.
@@
 
-Correct decoding occurs when $\hat{\cbf} = \cbf$, or equivalently when $\ybf \in \vor(\Lambda) + \cbf$, or equivalently when $\wbf \in \vor(\Lambda)$, i.e.
+% Correct decoding occurs when $\hat{\cbf} = \cbf$, or equivalently when $\ybf \in \vor(\Lambda) + \cbf$, or equivalently when $\wbf \in \vor(\Lambda)$, i.e.
 when the noise $\wbf$ is inside the Voronoi cell of the lattice.
 Assuming that the codeword $\cbf$ is transmitted with probability $p(\cbf)$, then the probability of correct lattice decoding is
+
+Lattices can be used to construct \emph{lattice codes}.
+If a lattice code is constructed from a lattice $\Lambda$, and is used in the AWGN channel with noise variance $\sigma^2$, then the probability of correct decoding can be described using the Voronoi cell,
 \begin{align}
-P_C &= \sum_{\cbf \in \calC} p(\cbf) \prob( \hat{\cbf} = \cbf ) \\
-&= \sum_{\cbf\in \calC} p(\cbf) \prob( \cbf + \wbf \in \vor(\Lambda) + \cbf )  \nonumber \\
-&= \prob(\wbf \in \vor(\Lambda) ) \nonumber \\
-&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
+P_C =  \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
 \end{align}
@@
 %This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.
-We call these the \emph{moments} of the $\vor(\Lambda)$.
-
-\newcommand{\calM}{M}
-In this paper we focus on $n$-dimensional lattice codes constructed from the family of lattices called $A_n$ and we derive expressions for the moments
-\[
- \calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf.
-\]
-These can be summed in (\ref{eq:summomentproberror}) to give arbitrarily accurate approximations for the probability of error.
+We call these the \emph{moments} of $\vor(\Lambda)$.
 
@@
 
-\begin{figure*}[htbp]
-	\centering
-		\includegraphics{../../plots/thpplot-1.mps}
-		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5, A_8$ and $E_8$.}
-		\label{fig:peplots}
-\end{figure*}
-
 % \begin{figure*}[htbp]
@@
 \section{The lattice $A_n$}\label{sec:lattice-a_n}
-Let $H$ be the hyperplane orthogonal to the all ones vector of length $n+1$, denoted by $\onebf$, that is
-\[
-\onebf = \left[ \begin{array}{cccc} 1 & 1 & \cdots & 1 \end{array} \right]^\prime,
-\]
-where superscript $^\prime$ indicates the transpose.
-Any vector in $H$ has the property that the sum of its elements is zero and for this reason $H$ is often referred to as the \emph{zero-sum plane}.
-The lattice $A_n$ is the intersection of the integer lattice $\ints^{n+1}$ with the zero-sum plane, that is
+Let $H$ be the hyperplane orthogonal to the all ones vector of length $n+1$, denoted by $\onebf$, that is $\onebf = \left[ \begin{array}{cccc} 1 & 1 & \cdots & 1 \end{array} \right]^\prime$, where superscript $^\prime$ indicates the transpose.
+The lattice $A_n$ is the intersection of $\ints^{n+1}$ with $H$, that is
 \begin{equation}
@@
 \end{equation}
-Equivalently, $A_n$ consists of all of those points in $\ints^{n+1}$ with coordinate sum equal to zero.
+Equivalently, $A_n$ is those points in $\ints^{n+1}$ with coordinate sum equal to zero.
 % A generator matrix for $A_n$ is the $(n+1)\times n$ matrix
@@
 % \]
-In two dimensions $A_2$ is the hexagonal lattice, and in three dimension $A_3$ is the body centered cubic lattice~\cite[p.
+%In two dimensions $A_2$ is the hexagonal lattice, and in three dimension $A_3$ is the body centered cubic lattice~\cite[p.
 108]{SPLAG}.
@@
 %\end{figure}
-
-\section{The main result}\label{sec:main-result}
-
-We now state our main result.  The moment $\calM_n(m)$ of the lattice $A_n$ satisfies
+\newcommand{\calM}{M}
+Let $\calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf$ be a moment of the Voronoi cell of $A_n$.  Our main results is as follows.  The moment $\calM_n(m)$ satisfies
 \begin{equation}\label{eq:theCmformula}
@@
 \end{equation}
-with the initial conditions
-\[
-G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
-\]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first three such formula are
+with the initial conditions $G(1,c,d) = \frac{1}{2c+d+1}$ and $G(n,0,0) = 1$.
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.
+The first four such formula are
 \begin{align*}
 \calM_n(0) &= \sqrt{n+1} \qquad \text{the volume},
-\\ \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment~\cite[p. 462]{SPLAG}}, 
-\\ \calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}.
-%\\ \calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
+\\ \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment}, 
+\\ \calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}},
+\\ \calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}.
 %\\ \calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
@@
 
+\begin{figure*}[t]
+	\centering
+		\includegraphics{../../plots/thpplot-1.mps}
+		\caption{The probability of error versus SNR for $A_1 \simeq \ints$, $A_2$, $A_3\simeq D_3, A_4, A_5$ and  $A_8$.}
+		\label{fig:peplots}
+\end{figure*}
+
 \section{Results and simulations}\label{sec:results-simulations}
 
-We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3, A_4, A_5$ and $A_8$.  For these plots the SNR is related to noise variance according to
-\[
-\text{SNR} = \frac{V^{2/n}}{4\sigma^2},
-\]
-where $V$ is the volume of the Voronoi cell and $n$ is the dimension of the lattice~\cite[p.
+We now plot the probability of coding error versus signal to noise ratio (SNR) for the lattices $A_1, A_2, A_3, A_4, A_5$ and $A_8$.
+For these plots the SNR is related to noise variance according to $\text{SNR} = \frac{V^{2/n}}{4\sigma^2}$, where $V$ is the volume of the Voronoi cell and $n$ is the dimension of the lattice.
+%~\cite[p.
 167]{Viterbo_diamond_cutting_1996}.
@@
 448]{SPLAG}.
-The simulations are iterated until 500 error events occur.
+The simulations are iterated until 5000 error events occur.
 
-The plot also displays an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
+%The plot also displays an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
 The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
@@
 Applying the union bound the probability of error satisfies
-\newcommand{\erfc}{\operatorname{erfc}}
-\newcommand{\erf}{\operatorname{erf}}
-\[
-P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
-\]
-where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
+%\newcommand{\erfc}{\operatorname{erfc}}
+%\newcommand{\erf}{\operatorname{erf}}
+%\[
+%P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
+%\]
+%where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
 For the $E_8$ lattice this approximation is an upper bound because the relevant vectors of $E_8$ (those vectors that define the Voronoi cell) are precisely the 240 minimal vectors.
@@
 
-\small
-\bibliography{../../bib}
+%\bibliography{../../bib}
 

### 2012-03-27 other/threepager/threepager.tex
Added stuff for threepager

@@
 
-\documentclass[final]{proceedings}
+\documentclass[final]{siamltex}
 %\documentclass[correspondence,9pt]{IEEEtran}
 
-\usepackage{crop}
-\crop
+%\usepackage{crop}
+%\crop
 %\makeindex
@@
 
-%\title{On the error performance of the $A_n$ lattices}
-%\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson
- %\thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
-Emanuele Viterbo is %with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
-Vaughan~Clarkson is with the School of Information %Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
-A preliminary full version of this paper is available at http://arxiv.org/abs/1111.6334.}
-%}
-
-\begin{document}
-
-%opening
-\chapter[On the error performance of the $A_n$ lattices]{On the error performance of the $A_n$ lattices}
-
-\begin{authorline}
-Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo and I.
+\title{On the error performance of the $A_n$ lattices}
+\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I.
 Vaughan L.
@@
 A preliminary full version of this paper is available at http://arxiv.org/abs/1111.6334.}
-\end{authorline}
+}
+
+\begin{document}
+
+%opening
+%\chapter[On the error performance of the $A_n$ lattices]{On the error performance of the $A_n$ lattices}
+
+%\begin{authorline}
+%Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo and I.
+Vaughan L.
+Clarkson\thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Emanuele Viterbo is with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+A preliminary full version of this paper is available at http://arxiv.org/abs/1111.6334.}
+%\end{authorline}
+
+\maketitle
+
+\begin{keywords} 
+Lattices, lattice decoding, root lattice, probability of error, Voronoi cell.
+\end{keywords}
+
+%\begin{AMS}
+%15A15, 15A09, 15A23
+%\end{AMS}
 
@@
 
+\section*{EXTENDED ABSTRACT}
 The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the  additive white Gaussian noise (AWGN) channel.  %~\cite{SPLAG}. 
@@
 This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.
+
 %Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices~\cite{Conway1982VoronoiRegions}.
@@
 
-\section{Lattices, lattice codes, and lattice decoding} \label{sec:latt-latt-codes}
+\section*{Lattices, lattice codes, and lattice decoding} \label{sec:latt-latt-codes}
 
@@
 
-\section{The lattice $A_n$}\label{sec:lattice-a_n}
+\section*{The lattice $A_n$}\label{sec:lattice-a_n}
 Let $H$ be the hyperplane orthogonal to the all ones vector of length $n+1$, denoted by $\onebf$, that is $\onebf = \left[ \begin{array}{cccc} 1 & 1 & \cdots & 1 \end{array} \right]^\prime$, where superscript $^\prime$ indicates the transpose.
@@
 
-\section{Results and simulations}\label{sec:results-simulations}
+\section*{Results and simulations}\label{sec:results-simulations}
 
@@
 
-%The plot also displays an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
-The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.
-The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.
-The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.
-Applying the union bound the probability of error satisfies
-%\newcommand{\erfc}{\operatorname{erfc}}
-%\newcommand{\erf}{\operatorname{erf}}
-%\[
-%P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
-%\]
-%where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
+The plot also displays an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
+The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice.%~\cite[p.~71]{SPLAG}.
+The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.  The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.  Applying the union bound the probability of error satisfies
+\newcommand{\erfc}{\operatorname{erfc}}
+\newcommand{\erf}{\operatorname{erf}}
+\[
+P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
+\]
+where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.
 For the $E_8$ lattice this approximation is an upper bound because the relevant vectors of $E_8$ (those vectors that define the Voronoi cell) are precisely the 240 minimal vectors.

### 2012-04-22 other/threepager/threepager.tex
Added new siam format to the three pager.
Added the final MTNS 3 page version

@@
 
-%\begin{AMS}
-%15A15, 15A09, 15A23
-%\end{AMS}
+\begin{AMS}
+94B70
+\end{AMS}
 
@@
 \newcommand{\calM}{M}
-Let $\calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf$ be a moment of the Voronoi cell of $A_n$.  Our main results is as follows.  The moment $\calM_n(m)$ satisfies
+Let $\calM_n(m) = \int_{\vor(A_n)} \|\xbf\|^{2m} d\xbf$ be a moment of the Voronoi cell of $A_n$.  Our main result is as follows.  The moment $\calM_n(m)$ satisfies
 \begin{equation}\label{eq:theCmformula}
@@
 The plot also displays an approximation for the probability of error for the 8-dimensional $E_8$ lattice.
-The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice.%~\cite[p.~71]{SPLAG}.
-The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.  The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.  Applying the union bound the probability of error satisfies
+The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice.
+The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.
+The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.
+Applying the union bound the probability of error satisfies
 \newcommand{\erfc}{\operatorname{erfc}}

### 2012-04-22 other/threepager/build.sh

@@ -3,8 +3,8 @@ mpost -interaction=nonstopmode ../../plots/thpplot.mp
 ruby1.9.1 hexlatfig.rb 
 
 #build tex 
-latex threepager.tex 
+pdflatex threepager.tex 
 bibtex threepager 
-latex threepager.tex 
-latex threepager.tex 
+pdflatex threepager.tex 
+pdflatex threepager.tex 
 

### 2012-04-27 double.tex
Added start of the double column version

--- 2012-03-12 paper.tex
@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
 %\documentclass[journal, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 %\documentclass[correspondence,9pt]{IEEEtran}
@@
 \title{On the error performance of the $A_n$ lattices}
-\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson
+\author{Copyright (c) 2012 IEEE.
+Personal use of this material is permitted.
+However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.
+Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I.
+Vaughan L.
+Clarkson
  \thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
@@
 \]
-For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$ (see Appendix~\ref{sec:solv-this-recurs}).  The first five such formula are:
+For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$ (see Appendix~\ref{sec:solv-this-recurs}).  The first three such formula are:
 \begin{align*}
-\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
-\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, \\
-\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
-\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
-\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
+\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of~}\vor(A_n),\\
+\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the 2nd moment~ \cite[p. 462]{SPLAG}}, \\
+\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}.
+%\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
+%\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
 \end{align*}

### 2012-04-27 double.tex
Added to double version.  Everything fits on one line now.

@@
 \title{On the error performance of the $A_n$ lattices}
-\author{Copyright (c) 2012 IEEE.
+\author{Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I. Vaughan L. Clarkson
+ \thanks{Copyright (c) 2012 IEEE.
 Personal use of this material is permitted.
 However, permission to use this material for any other purposes must be obtained from the IEEE by sending a request to pubs-permissions@ieee.org.
-Robby McKilliam, Ramanan Subramanian, Emanuele Viterbo, I.
-Vaughan L.
-Clarkson
- \thanks{Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
 Emanuele Viterbo is with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
@@
 \begin{equation}\label{eq:vorfint}
-\frac{1}{\sqrt{n+1}} \int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \frac{1}{\sqrt{n+1}} \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
+\begin{split}
+\frac{1}{\sqrt{n+1}} &\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf \\
+&= \frac{1}{\sqrt{n+1}} \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
+\end{split}
 \end{equation}
@@
 \begin{equation}\label{eq:vorfintafternorm}
-\frac{1}{\sqrt{n+1}} \int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
+\begin{split}
+\frac{1}{\sqrt{n+1}} &\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf \\
+&=  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
+\end{split}
 \end{equation}
@@
 Our integral is now equal to
-\[
-\frac{n (n+1)}{\sqrt{n+1}} \int^{1/2}_{-1/2} \int^{x_1}_{-1/2}  \int^{x_1}_{x_2} \cdots \int^{x_1}_{x_2} \frac{f(\Qbf\xbf)}{1 - x_{1} + x_{2}}  dx_{n+1} \, \dots \, dx_2 \, dx_1.
-\]
+\begin{align*}
+\frac{n (n+1)}{\sqrt{n+1}} \int^{1/2}_{-1/2} \int^{x_1}_{-1/2}  \int^{x_1}_{x_2} \cdots \int^{x_1}_{x_2} &\frac{f(\Qbf\xbf)}{1 - x_{1} + x_{2}}  \\
+&\hspace{0.5cm} dx_{n+1} \, \dots \, dx_2 \, dx_1.
+\end{align*}
 The factor $n(n+1)$ arises because there are $n(n+1)$ ways to place two elements (i.e. $x_1$ and $x_2$) into $n+1$ positions.
@@
 We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables, gives
-\[
-n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} \frac{f(\Qbf\xbf )}{1 - t}  dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
-\]
+\begin{align*}
+n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} &\frac{f(\Qbf\xbf )}{1 - t}  \\
+&dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
+\end{align*}
 and letting $w_{i-2} = x_i - y + 1/2 + t$ for $i = 3,\dots,n+1$ gives
@@
 \begin{example}\emph{\textbf{(The volume of the Voronoi cell)}
-In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting $f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1$ we obtain,
+In order to demonstrate this approach we will derive the $0$th moment (i.e. the volume) of the Voronoi cell using~(\ref{eq:intwewilluse}).  Setting 
+\[
+f(\Qbf\wbf) = \|\Qbf\wbf\|^0 = 1
+\] 
+we obtain,
 \begin{align*}
-M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{1}{1 - t} \,dw_{n-1}\dots dw_1 \, dt  \, dy\\
+M_n(0) &= n\sqrt{n+1} \int^{1}_{0} \int^{y}_{0} \int^{t}_{0} \cdots \int^{t}_{0} \frac{dw_{n-1}\dots dw_1 \, dt  \, dy}{1 - t}\\
  &= n\sqrt{n+1} \int_{0}^{1} \int^{y}_{0} \frac{t^{n-1}}{1 - t} \, dt  \, dy \\
@@
 We now derive expressions for the $\calM_n(m)$.  Setting $f(\Qbf\xbf) = \left(\|\Qbf\wbf\|^{2}\right)^m$ in~\eqref{eq:intwewilluse} we obtain,
-\[
-\frac{\calM_n(m)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
-\]
+\begin{align*}
+\frac{\calM_n(m)}{n\sqrt{n+1}} = \int_0^1 \int_0^y \int_0^t \cdots \int_0^t &\frac{\left(\|\Qbf\wbf\|^2\right)^m}{1 - t} \\
+&dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
+\end{align*}
 Now $\|\Qbf\wbf\|^2 = \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2$ and recalling that $\wbf = [t,0,w_1,\dots,w_{n-1}]'$ we can write
 \begin{align*}
-\|\Qbf\wbf\|^2 &= \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2 \\
+&\|\Qbf\wbf\|^2 \\
+&= \|\wbf\|^2 - \frac{1}{n+1}(\wbf^\prime \onebf)^2 \\
 &= t^2 + \sum_{i=1}^{n-1}w_i^2 - \frac{1}{n+1}\left( t + \sum_{i=1}^{n-1}w_i \right)^2 \\
@@
 \[
-C = \left(\frac{n}{n+1} \right) t^2 \qquad \text{and} \qquad D = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
+C = \left(\frac{n}{n+1} \right) t^2 \;\;\;\; \text{and} \;\;\;\; D = A - \frac{2t}{n+1}B - \frac{1}{n+1}B^2,
 \]
@@
 Now,
-\[
- \frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t \frac{(C+D)^m}{1 - t} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
-\]
+\begin{align*}
+ \frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \int_0^t \cdots \int_0^t &\frac{(C+D)^m}{1 - t} \\
+&dw_{n-1} \, \dots \, dw_1 \, dt \, dy,
+\end{align*}
 and by expanding the binomial $(C+D)^m$ we get
-\[
-\frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} C^{m-k} \int_0^t \cdots \int_0^t \, D^{k} \,dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
-\]
+\begin{align*}
+\frac{\calM_n(m)}{n\sqrt{n+1}} =  \int_0^1 \int_0^y \frac{1}{1 - t}\sum_{k=0}^{m}\binom{m}{k} &C^{m-k} \int_0^t \cdots \int_0^t \, D^{k} \\
+&dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
+\end{align*}
 Expanding $D^k$ as a trinomial gives
@@
 \begin{equation} \label{eq:recurseint}
-\int_0^t \cdots \int_0^t A^{a} B^{2k-2a-b} dw_{n-1} \, \dots \, dw_1 = t^{n-1+2k - b} G(n-1,a,2k - 2a - b).
+\begin{split}
+  \int_0^t \cdots \int_0^t &A^{a} B^{2k-2a-b} dw_{n-1} \, \dots \, dw_1 \\
+&= t^{n-1+2k - b} G(n-1,a,2k - 2a - b).
+\end{split}
 \end{equation}
@@
 \begin{align*}
-\frac{\calM_n(m)}{n\sqrt{n+1}} &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy \\
+&\frac{\calM_n(m)}{n\sqrt{n+1}} \\
+&= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \int_0^y \frac{t^{n - 1 + 2m}}{1 - t} \, dt \, dy \\
 &= \sum_{k=0}^{m} \binom{m}{k} \left(\frac{n}{n+1} \right)^{m-k} P \int_0^1 \beta(y,n+2m,0) dy \\
@@
 In~(\ref{eq:recurseint}) we required to evaluate integrals of the form
-\[
-F(n-1,a,2k - 2a - b) = \int^{t}_{0} \cdots \int^{t}_{0} A^a B^{2k - 2a - b} \,dw_{n-1}\cdots dw_{1},
-\]
+\begin{align*}
+F(n-1,a,&2k - 2a - b) \\
+&= \int^{t}_{0} \cdots \int^{t}_{0} A^a B^{2k - 2a - b} \,dw_{n-1}\cdots dw_{1},
+\end{align*}
 or equivalently, integrals of the form
@@
 \]
-where $n,c$ and $d$ are integers.  We will find a recursion describing this integral.  Write
-\[
-F(n,c,d) = \int^t_0 \cdots \int^t_0 \left(x_n^2 + \sum_{j=1}^{n-1} x_j^2\right)^c \left( x_n + \sum_{i=1}^{n-1} x_i \right)^d \,dx_1\cdots dx_n.
-\]
+where $n, c$ and $d$ are integers.  We will find a recursion describing this integral.  Write
+\begin{align*}
+&F(n,c,d) \\
+&= \int^t_0 \cdots \int^t_0 \left(x_n^2 + \sum_{j=1}^{n-1} x_j^2\right)^c \left( x_n + \sum_{i=1}^{n-1} x_i \right)^d \,dx_1\cdots dx_n.
+\end{align*}
 Expanding the two binomials gives
 \begin{align*}
-F(n,c,d) &=  \int^t_0 \cdots \int^t_0 \sum_{c'=0}^{c} \binom{c}{c'} x_n^{2c'} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'}  \sum_{d'=0}^{d}\binom{d}{d'} x_n^{d'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n} \\
-&=  \sum_{c'=0}^{c} \sum_{d'=0}^d \int^t_0 \cdots \int^t_0 \binom{c}{c'} \binom{d}{d'} x_n^{2c'+d'} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n}.
+&F(n,c,d) \\
+&=  \int^t_0 \cdots \int^t_0  \sum_{c'=0}^{c} \binom{c}{c'} x_n^{2c'} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'}  \\
+& \hspace{2.6cm}  \sum_{d'=0}^{d}\binom{d}{d'} x_n^{d'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n}  \\
+&= \sum_{c'=0}^{c} \sum_{d'=0}^d \int^t_0 \cdots \int^t_0 \binom{c}{c'} \binom{d}{d'} x_n^{2c'+d'} \\ 
+& \hspace{2.6cm} \left(\sum_{j=1}^{n-1} x_j^2\right)^{c-c'} \left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n}.
 \end{align*}
 Integrating the $x_n$ term gives
-\[
-F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1} \int^t_0 \cdots \int^t_0 \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
-\]
+\begin{align*}
+F(&n,c,d) \\
+&= \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1} \int^t_0 \cdots \int^t_0 \\
+& \hspace{1cm} \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
+\end{align*}
 Note that
-\[
-F(n-1,c-c',d-d')  = \int^t_0 \cdots \int^t_0 \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
-\]
+\begin{align*}
+&F(n-1,c-c',d-d')  \\
+&= \int^t_0 \cdots \int^t_0 \left(\sum_{j=1}^{n-1}x_j^2\right)^{c-c'}\left(\sum_{i=1}^{n-1} x_i \right)^{d-d'} \,dx_1\cdots dx_{n-1}.
+\end{align*}
 So $F(n,c,d)$ satisfies the recursion
-\[
-F(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1}F(n-1,c-c',d-d')
-\]
+\begin{align*}
+&F(n,c,d) \\
+&= \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1}F(n-1,c-c',d-d')
+\end{align*}
 with the initial conditions
@@
 \begin{align*}
-F(n,c,d) &=  \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{t^{2c'+d'+1}}{2c'+d'+1}t^{n-1+2(c-c')+d-d'}G(n-1,c-c',d-d') \\
+&F(n,c,d) \\
+&=  \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d') t^{2c'+d'+1}}{(2c'+d'+1) t^{1-n-2(c-c')-d+d'}} \\
 &=  t^{n+2c+d}\sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1} \\
@@
 For fixed $d$ and $c$ this recursion can be solved explicitly.  Write
-\[
-G(n,c,d) = G(n-1,c,d) + \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
-\]
+\begin{align*}
+G(n,c,d) &= G(n-1,c,d) \\
+&\hspace{0.2cm}+ \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
+\end{align*}
 where the sum $\sum_{ (c',d') \neq (0,0)}$ is over all $0 \leq c' \leq c$ and $0 \leq d' \leq d$ except when both $d$ and $c$ are zero.
@@
 Taking the $z$-transform of both sides in the equation above gives
-\[
-\calG(z,c,d) = \frac{z^{-1}}{1-z^{-1}} \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{\calG(z,c-c',d-d')}{2c'+d'+1}.
-\]
+\begin{align*}
+&\calG(z,c,d) \\
+&= \frac{z^{-1}}{1-z^{-1}} \sum_{ (c',d') \neq (0,0)} \binom{c}{c'}\binom{d}{d'} \frac{\calG(z,c-c',d-d')}{2c'+d'+1}.
+\end{align*}
 So the $z$-transform $\calG(z,c,d)$ satisfies this recursive equation.

### 2012-04-27 double.tex
Added biographies to double

@@
 
+\begin{biography}{Robby G. McKilliam}
+was born in Brisbane, Queensland, Australia in 1983.
+He received the B.Sc.
+degree in mathematics and the B.E.
+degree (Hons.
+I) in computer systems engineering from the University of Queensland, Brisbane, Australia in 2006.
+In December 2010 he completed his PhD at the University of Queensland, Brisbane, Australia.
+He is now at the Institute for Telecommunications Research at the University of South Australia, Adelaide.
+His fields of interest are lattice theory, number theory, estimation theory and signal processing and communications.
+\end{biography}
+
+\begin{biography}{Ramanan Subramanian}
+\end{biography}
+
+\begin{biography}{Emanuele Viterbo}
+received his degree (Laurea) in Electrical Engineering
+in 1989 and his Ph.D. in 1995 in Electrical Engineering, both from the
+Politecnico di Torino, Torino, Italy.  From 1990 to 1992 he was with the
+European Patent Office, The Hague, The Netherlands, as a patent examiner
+in the field of dynamic recording and error-control coding. Between 1995
+and 1997 he held a post-doctoral position in the Dipartimento di
+Elettronica of the Politecnico di Torino. In 1997-98 he was a
+post-doctoral research fellow in the Information Sciences Research
+Center of AT\&T Research, Florham Park, NJ, USA. He became first
+Assistant Professor (1998) then Associate Professor (2005) in
+Dipartimento di Elettronica at Politecnico di Torino. In 2006 he became
+Full Professor in DEIS at University of Calabria, Italy. From 2010 he is
+Full Professor in the ECSE Department and Associate Dean for Research
+Training in the Faculty of Engineering at Monash University, Melbourne,
+Australia.
+Prof. Emanuele Viterbo is a 2011 Fellow of the IEEE,  a ISI Highly Cited
+Researcher and Member of the Board of Governors of the IEEE Information
+Theory Society (2011-2013). He is Associate Editor of IEEE Transactions
+on Information Theory, European Transactions on Telecommunications and
+Journal of Communications and Networks, and Guest Editor for IEEE
+Journal of Selected Topics in Signal Processing: Special Issue Managing
+Complexity in Multiuser MIMO Systems.
+In 1993 he was visiting researcher in the Communications Department of
+DLR, Oberpfaffenhofen, Germany. In 1994 and 1995 he was visiting the
+cole Nationale Suprieure des Telcommunications (E.N.S.T.), Paris. In
+2003 he was visiting researcher at the Maths Department of EPFL,
+Lausanne, Switzerland. In 2004 he was visiting researcher at the
+Telecommunications Department of UNICAMP, Campinas, Brazil. In 2005,
+2006 and 2009 he was visiting researcher at the ITR of UniSA, Adelaide,
+Australia. In 2007 he was visiting fellow at the Nokia Research Center,
+Helsinki, Finland.
+Dr. Emanuele Viterbo was awarded a NATO Advanced Fellowship in 1997 from
+the Italian National Research Council. His main research interests are
+in lattice codes for the Gaussian and fading channels, algebraic coding
+theory, algebraic space-time coding, digital terrestrial television
+broadcasting, and digital magnetic recording.
+\end{biography}
+
+\begin{biography}{I. Vaughan L. Clarkson}
+was born in Brisbane,
+Queensland, Australia, in 1968. He received the
+B.Sc. degree in mathematics and the B.E. degree
+(Hons. I) in computer systems engineering from
+The University of Queensland, Brisbane, in 1989
+and 1990, respectively, and the Ph.D. degree in
+systems engineering from The Australian National
+University, Canberra, in 1997.
+In 1988, he was with the Defence Science and
+Technology Organisation, Adelaide, Australia, ﬁrst
+as a Cadet, later as a Professional Ofﬁcer, and ﬁnally
+as a Research Scientist. From 1998 to 2000, he was a Lecturer at The University
+of Melbourne, Melbourne, Australia. From 2000 to 2008, he was a Senior
+Lecturer in the School of Information Technology and Electrical Engineering
+at The University of Queensland. In 2008, he was promoted to Reader. In 2005,
+he was a Visiting Professor in the Department of Electrical and Computer
+Engineering, The University of British Columbia, Vancouver, Canada. His
+research interests include statistical signal processing for communications and
+defense, image processing, information theory, and lattice theory.
+\end{biography}
+
 \end{document}

### 2012-04-29 double.tex
minor edits?
Added to double

@@
 
+\begin{figure*}[bt]
+% ensure that we have normalsize text
+%\normalsize
+\small
+% Store the current equation number.
+%\setcounter{MYtempeqncnt}{\value{equation}}
+% Set the equation number to one less than the one
+% desired for the first equation here.
+% The value here will have to changed if equations
+% are added or removed prior to the place these
+% equations are referenced in the main text.
+%\setcounter{equation}{5}
+\begin{align*}
+\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}} \\
+\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}} \\
+\calM_n(5) &= \frac{3577728 n - 1825648 n^2 + 2410804 n^3 + 1569392 n^4 + 1644423 n^5 + 906105 n^6+341550 n^7+75526 n^8+8855 n^9+385 n^{10}}{95800320 (1 + n)^{9/2}}
+\end{align*}
+% Restore the current equation number.
+%\setcounter{equation}{\value{MYtempeqncnt}}
+% IEEE uses as a separator
+\hrulefill
+% The spacer can be tweaked to stop underfull vboxes.
+\vspace*{4pt}
+\end{figure*}
+
 \section{The main result}\label{sec:main-result}
@@
 \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the 2nd moment~ \cite[p. 462]{SPLAG}}, \\
-\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}.
+\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}
 %\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
@@
 \end{align*}
+and the next three formula are displayed at the top of this page.
 We have explicitly tabulated these formula for $m=0$ to $40$.
@@
 
-\begin{biography}{Robby G. McKilliam}
+\begin{IEEEbiographynophoto}{Robby G. McKilliam}
 was born in Brisbane, Queensland, Australia in 1983.
@@
 His fields of interest are lattice theory, number theory, estimation theory and signal processing and communications.
-\end{biography}
+\end{IEEEbiographynophoto}
 
-\begin{biography}{Ramanan Subramanian}
-\end{biography}
+\begin{IEEEbiographynophoto}{Ramanan Subramanian}
+received the Bachelor of Technology degree in Electrical Engineering from the Indian Institute of Technology, Madras in 2003, the Master of Science and PhD degrees in Electrical and Computer Engineering from the Georgia Institute of Technology, Atlanta, USA in 2006 and 2009 respectively.
+Since late 2009, he has been working as a research fellow at the Institute for Telecommunications Research (ITR) at the University of South Australia in Adelaide.
+His research interests include wireless networking, information theory, queuing theory, and statistics.
+\end{IEEEbiographynophoto}
 
-\begin{biography}{Emanuele Viterbo}
+\begin{IEEEbiographynophoto}{Emanuele Viterbo}
 received his degree (Laurea) in Electrical Engineering
@@
 broadcasting, and digital magnetic recording.
-\end{biography}
+\end{IEEEbiographynophoto}
 
-\begin{biography}{I. Vaughan L. Clarkson}
+\begin{IEEEbiographynophoto}{I. Vaughan L. Clarkson}
 was born in Brisbane,
@@
 defense, image processing, information theory, and lattice theory.
-\end{biography}
+\end{IEEEbiographynophoto}
 

### 2012-04-29 double.tex
added more to double column version

@@
 \calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the 2nd moment~ \cite[p. 462]{SPLAG}}, \\
-\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}
+\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}},
 %\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
@@
 Figure~\ref{fig:cuberothex} depicts some 2-dimensional zonotopes.
-A zonotope is \emph{unimodular} if it can be used to tile (or tesselate) Euclidean space.
-Such a tesselation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
+A zonotope is \emph{unimodular} if it can be used to tile (or tessellate) Euclidean space.
+Such a tessellation naturally gives rise to a lattice, a so called \emph{zonotopal lattice}~\cite{Vallentin_space_tiling_zonotope_2004}.
 The lattice $A_n$ is a zonotopal lattice and the techniques in this paper can potentially be extended to other zonotopal lattices.
@@
 Australia.
+
 Prof. Emanuele Viterbo is a 2011 Fellow of the IEEE,  a ISI Highly Cited
@@
 Complexity in Multiuser MIMO Systems.
+
 In 1993 he was visiting researcher in the Communications Department of
@@
 Helsinki, Finland.
+
 Dr. Emanuele Viterbo was awarded a NATO Advanced Fellowship in 1997 from
@@
 \begin{IEEEbiographynophoto}{I. Vaughan L. Clarkson}
-was born in Brisbane,
-Queensland, Australia, in 1968. He received the
+was born in Brisbane, Queensland, Australia, in 1968. He received the
 B.Sc. degree in mathematics and the B.E. degree
@@
 University, Canberra, in 1997.
-In 1988, he was with the Defence Science and
-Technology Organisation, Adelaide, Australia, ﬁrst
-as a Cadet, later as a Professional Ofﬁcer, and ﬁnally
-as a Research Scientist. From 1998 to 2000, he was a Lecturer at The University
-of Melbourne, Melbourne, Australia. From 2000 to 2008, he was a Senior
-Lecturer in the School of Information Technology and Electrical Engineering
+
+In 1988, he was with the Defence Science and Technology Organisation, Adelaide, Australia, first as a Cadet, later as a Professional Officer, and finally as a Research Scientist.
+From 1998 to 2000, he was a Lecturer at The University of Melbourne, Melbourne, Australia.
+From 2000 to 2008, he was a Senior Lecturer in the School of Information Technology and Electrical Engineering
 at The University of Queensland. In 2008, he was promoted to Reader. In 2005,

### 2012-04-29 double.tex
added Vaughan's biography edits

@@
 \begin{IEEEbiographynophoto}{I. Vaughan L. Clarkson}
-was born in Brisbane, Queensland, Australia, in 1968. He received the
-B.Sc. degree in mathematics and the B.E. degree
-(Hons. I) in computer systems engineering from
-The University of Queensland, Brisbane, in 1989
-and 1990, respectively, and the Ph.D. degree in
-systems engineering from The Australian National
-University, Canberra, in 1997.
+was born in Brisbane, Queensland, Australia, in 1968.
+He received a B.Sc.
+in Mathematics and a B.E.
+(Hons.
+I) in Computer Systems Engineering from The University of Queensland, Brisbane, Australia, in 1989 and 1990, respectively, and a Ph.D.
+in Systems Engineering from The Australian National University, Canberra, Australia, in 1997.
 
-In 1988, he was with the Defence Science and Technology Organisation, Adelaide, Australia, first as a Cadet, later as a Professional Officer, and finally as a Research Scientist.
+Starting in 1988, he was employed by the Defence Science and Technology Organisation, Adelaide, Australia, first as a Cadet, later as a Professional Officer, and finally as a Research Scientist.
 From 1998 to 2000, he was a Lecturer at The University of Melbourne, Melbourne, Australia.
-From 2000 to 2008, he was a Senior Lecturer in the School of Information Technology and Electrical Engineering
-at The University of Queensland. In 2008, he was promoted to Reader. In 2005,
-he was a Visiting Professor in the Department of Electrical and Computer
-Engineering, The University of British Columbia, Vancouver, Canada. His
-research interests include statistical signal processing for communications and
-defense, image processing, information theory, and lattice theory.
+From 2000 to 2008, he was a Senior Lecturer in the School of Information Technology and Electrical Engineering at The University of Queensland.
+In 2008, he was promoted to Reader.
+He was a Visiting Professor in the Department of Electrical and Computer Engineering at The University of British Columbia, Vancouver, Canada, in 2005 and at the Institute of Telecommunications, Technical University of Vienna, Austria, in 2012.
+His research interests include statistical signal processing for communications and defence, image processing, information theory and lattice theory.
 \end{IEEEbiographynophoto}

### 2012-05-01 double.tex
minor double edit

@@
 Robby~McKilliam and Ramanan Subramanian are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
-Emanuele Viterbo is with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia.
+Emanuele Viterbo is with the Department of Electrical and Computer Systems Engineering, Monash University, Melbourne, VIC, Australia, and is supported by the Monash Professional Fellowship within the Monash Software Defined Telecommunications Lab.
 Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.}
